### 2022-07-18 gtlc.tex
Hello, synthetic world!

\documentclass{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{tikz-cd}
\usepackage{mathpartir}
\usepackage{stmaryrd}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{nonnum-theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{lemma}{Lemma}[section]

\newcommand{\dyn}{\text{Dyn}}
\newcommand{\dynv}{\text{Dyn}^+}
\newcommand{\dync}{\text{Dyn}^-}
\newcommand{\good}[1]{\text{good}_{#1}}
\newcommand{\prop}{\mathbb P}
\newcommand{\with}{\mathrel{\&}}
\newcommand{\eqv}{\simeq}
\newcommand{\soln}{\text{Sol}}
\newcommand{\police}[1]{\text{police}_{#1}}
\newcommand{\later}{{\blacktriangleright}}
\newcommand{\errp}[1]{\mho_{#1}}
\newcommand{\kerrp}[2]{\mho^{#1}_{#2}}
\newcommand{\err}{\mho}
\newcommand{\errlift}{L_\err}
\newcommand{\ostar}{\circledast}

\newcommand{\ltdynp}[1]{\mathrel{\sqsubseteq_{#1}}}
\newcommand{\ltdynlift}[1]{\mathrel{\sqsubseteq_{\errlift#1}}}
\newcommand{\ltprecp}[1]{\mathrel{\sqsubseteq^{\prec}_{\errlift #1}}}
\newcommand{\ltsuccp}[1]{\mathrel{\sqsubseteq^{\succ}_{#1}}}

\newcommand{\kltp}[2]{\mathrel{\sqsubseteq^{#1}_{#2}}}

\newcommand{\ltgt}[1]{\mathbin{\langle\mspace{-1mu}#1\mspace{-1mu}\rangle}}
\newcommand{\fmap}{\ltgt{\$}}
\newcommand{\monto}{\mathrel{\to_{m}}}
\newcommand{\linto}{\multimap}

\begin{document}

\title{Semantics of Gradual Types in SGDT}
\author{Max S. New}

\maketitle

\section{Goals}

We want to define a denotational semantics for the core language of
GTLC and prove graduality.

Here's roughly the structure required
\begin{verbatim}
record GTLC(U)
  type : U
  tm : type -> U
  _<ty=_ : PreOrder type
  _<tm=_ : PreOrder (Sigma A (tm A))
  impl : (A, M) <tm= (B,N) -> A <ty= B

  dyn : type
  dynTop : {A} -> A <ty= dyn

  ans : type
  yes : tm ans
  no  : tm ans

  arr : type -> type -> type
  lam : (tm A -> tm B) -> tm (arr A B)
  app : tm (arr A B) -> tm A -> tm B
  -- beta, eta

  up : A <ty= B -> tm A -> tm B
  upR : ...

  dn : A <ty= B -> tm B -> tm A

\end{verbatim}

\section{Error predomains and domains}

First, we need to define our monad.
\begin{definition}
  Define the error-lift type as a solution to the guarded recursive
  domain equation:
  \[ \errlift X \cong (X + 1 + \later \errlift X) \]

  We label the three injections as follows:
  \begin{align*}
    \eta &: X \to \errlift X\\
    \err &: \errlift X\\
    \theta &: \later \errlift X \to \errlift X
  \end{align*}

  This carries the structure of a strong monad, with $\eta$ as the
  unit and the join defined by L\"ob induction as follows:
  \begin{align*}
    \mu(\eta x) &= x\\
    \mu(\mho) &= \mho\\
    \mu(\theta(u)) &= \theta (\mu^\later \ostar u)
  \end{align*}
  With map (implying strength) defined as
  \begin{align*}
    \errlift f(\eta x) &= f x\\
    \errlift f(\mho) &= \mho\\
    \errlift f(\theta(u)) &= \theta ((\errlift f)^\later \ostar u)
  \end{align*}
  And the equational laws of a monad can be proven using L\"ob
  induction.
\end{definition}

This can be seen as simply a composition of two monads using a
distributive law: the error monad and the recursive lift monad.
%
However, we present them as one combined monad because in gradual
typing we are interested in a specific ordering relation called the
\emph{error ordering} where $\err$ is the least element. This is
\emph{not} the composition of the lift and error monad, which would
instead have the error as an isolated, maximal element.

We can define this error ordering as a variation on the weak
simulation relations of prior work:
\begin{definition}
  Let $X$ be a type and $\sqsubseteq_X$ a binary relation. We define
  $\ltdynp{\errlift X} : \errlift X \to \errlift X \to \prop$ as follows:
  \begin{align*}
    \err \ltdynlift X d &= \top\\
    \eta x \ltdynlift X \eta y &= x \ltdynp X y\\
    \eta x \ltdynlift X \err &= \bot\\
    \eta x \ltdynlift X \theta t &= \exists{n:\mathbb N}.\exists{y: X}. \theta t = \delta^n (\eta y) \wedge x \ltdynp X y\\
    \theta s \ltdynlift X \eta y &= \exists{n:\mathbb N}.(\theta s = \delta^n \err)\vee(\exists{y: X}. \theta t = \delta^n (\eta y) \wedge x \ltdynp X y)\\
    \theta s \ltdynlift X \err &= \exists{n:\mathbb N}.\theta s = \delta^n \err\\
    \theta s \ltdynlift X \theta t &= \later[l \leftarrow s,r\leftarrow t] s \ltdynlift X t
  \end{align*}
\end{definition}

\begin{lemma}
  When $\ltdynp X$ is reflexive, so is $\ltdynlift X$.
\end{lemma}
\begin{proof}
  By L\"ob induction on the proposition $\forall d: \errlift X. d
  \ltdynlift X d$. Error and return cases are immediate.  We need to
  show
  \[ \theta t \ltdynlift X \theta t = \later[l \leftarrow t,r\leftarrow t] l \ltdynlift X r = \later[l \leftarrow t] l \ltdynlift X l \]
  The inductive hypothesis $h : \later (\forall d: \errlift X. d \ltdynlift X d)$
  We can then conclude
  \[ \text{next}[f \leftarrow h, l \leftarrow t]. h l \]
\end{proof}

Unfortunately, transitivity is not preserved.
%
The problematic case is of the form $\theta s \ltdynlift X \theta t
\ltdynlift X \eta z$. In this case we know $\theta t = \delta^n (\eta
y)$ but we cannot conclude that $\theta s$ must eventually terminate!
In fact this is false in the semantics because it might be that
$\theta s$ is only related to $\theta t$ because we ran out of fuel,
and actually $s$ is a diverging computation.

However, this should support transitive reasoning \emph{in the limit},
i.e., if the equivalence holds for \emph{all} levels of approximation
then eventually if $\theta t$ terminates, at high enough level of
approximation it will no longer be related to the diverging
computation.
%
This reasoning \emph{in the limit} is accomplished analytically by
quantification over all steps.
%
Here we accomplish it by using \emph{clocks} and \emph{clock
quantification}.
%
That is rather than a denotation of a syntactic type being a type it
will be given by a ``clocked'' type:

\begin{definition}
  A \emph{clocked} type $X$ is an element of $K \to U$.

  We call $\forall k:K. X k$, the \emph{limit} of $X$.
\end{definition}

Note that, similar to the interval in cubical type theory, $K$ is not
a type, but only a ``pseudo-type'' that we can quantify over in the
type $\Pi_{k:K} B$.
%
Next, we need a ``clocked'' error ordering associated to every type.
%
This will be assumed to be reflexive at each level of approximation,
but only transitive in the limit.

\begin{definition}
  A \emph{clocked} preorder on a clocked type $X : K \to U$ is a
  clocked relation $\kltp {} X : \forall k. X_k \to X_k \to \prop$ satisfying
  \begin{enumerate}
  \item Clocked Reflexivity $\forall k. \prod_{x:X_k} x \kltp k X x$
  \item Transitivity in the limit $\prod_{x,y,z:\forall k. X_k} x \kltp \omega X y \to y \kltp \omega X z \to x \kltp \omega X z$
  \end{enumerate}

  Where 
  \[ x \kltp \omega X y = \forall k. x_k \kltp k X y_k \]

  A \emph{monotone function} is a function $f : \forall k. X_0 k \to
  Y_0 k$ satisfying monotonicity:
  \[ \forall k.\prod_{x,x':X_0^k} x \kltp k X x' \to f_k x \kltp k Y f_k x' \]
\end{definition}

For any clocked preorder we can construct the delayed preorder by
\begin{definition}
  Define $\later : K \to (K \to U) \to (K \to U)$ by
  \[ (\later X)^k = \later^k X^k \]
  and on preorders by:
  \begin{align*}
    x \kltp k {\later X} y &= \later^k[l\leftarrow x, r\leftarrow y].~l \kltp k {X} r
  \end{align*}

\end{definition}

Then we can define our semantic notions
\begin{definition}
  An \emph{error predomain} $X$ consists of
  \begin{enumerate}
  \item A clocked type $X_0$
  \item A clocked preorder $\kltp {} X$ on $X_0$ 
  \end{enumerate}
  A morphism of error predomains is a monotone function.

  An \emph{error domain} is an error predomain $X$ with
  \begin{enumerate}
  \item A least element $\kerrp {} X : \forall k. X$, i.e., it
    satisfies $\forall k. \prod_{x:X_k} \kerrp k X \kltp k X x$.
  \item A monotone ``thunk'' function $\theta_X : \later X \to X$
  \end{enumerate}
  A morphism of error domains is a morphism of the underlying error
  predomains.

  A \emph{linear} morphism of error domains $S : X \multimap Y$ is a
  morphism of error domains additionally satisfying
  \begin{enumerate}
  \item Error preservation $S^k \err^k_X = \err^k_Y$
  \item Thunk preservation $S^k (\theta^k_X x) = \theta^k_Y (S^k \fmap x)$
  \end{enumerate}
\end{definition}

Finally we can return to our original goal of defining the monad for our semantics!
%

\begin{lemma}
  next and $\delta$ are monotone
\end{lemma}

\begin{definition}
  Let $X$ be an error predomain a binary relation. We define the error
  domain $\errlift X$ as

  TODO: put the ks in the right spots

  \begin{enumerate}
  \item $(\errlift X)_0^k = \mu L. X_0^k + 1 + \later^k L$
  \item Ordering defined by cases:
    \begin{align*}
      \err \kltp k {\errlift X} d &= \top\\
      \eta x \kltp k {\errlift X} \eta y &= x \ltdynp X y\\
      \eta x \kltp k {\errlift X} \err &= \bot\\
      \eta x \kltp k {\errlift X} \theta t &= \exists{n:\mathbb N}.\exists{y: X}. \theta t = \delta^n (\eta y) \wedge x \ltdynp X y\\
      \theta s \kltp k {\errlift X} \eta y &= \exists{n:\mathbb N}.(\theta s = \delta^n \err)\vee(\exists{y: X}. \theta t = \delta^n (\eta y) \wedge x \ltdynp X y)\\
      \theta s \kltp k {\errlift X} \err &= \exists{n:\mathbb N}.\theta s = \delta^n \err\\
      \theta s \kltp k {\errlift X} \theta t &= \later[l \leftarrow s,r\leftarrow t] s \kltp k {\errlift X} t
    \end{align*}
  \item reflexive: as above
  \item TODO: transitive in the limit
  \item TODO: strong monad
  \end{enumerate}
\end{definition}

\begin{lemma}
  $\errlift X$ is the free error domain on $X$, i.e., it is left
  adjoint to the forgetful functor from the category of error domains
  and linear maps to the category of predomains.
\end{lemma}

\begin{lemma}
  The adjunction is monadic: error domains are precisely algebras of
  the free error domain. Is this true?
\end{lemma}

\section{Lazy Semantics}

To define a semantics for lazy gradually typed language, we first need
to define our \emph{universal} error domain. Unless we add in features
like pattern matching on the dynamic type, there is some freedom in
this choice. We present two natural solutions:

The first is a classic-looking dynamic type: the lift of the sum of
the base type and the function type. A notable difference are the
guarded recursive uses.
\[ \dyn_+ \cong \errlift (2 + (\later \dyn_+ \monto \later \dyn_+)) \]
The second is slightly closer to untyped lambda calculus: we take a
product rather than a sum, and only need to lift the base type itself.
\[ \dyn_- \cong \errlift 2 \times (\later \dyn_- \monto \later \dyn_-) \]
The idea here is that any dynamically typed term can be safely applied
to any number of arguments and at some point ``projected'' to return a
boolean, which will actually cause computation to occur as indicated
by the $\errlift$.

To construct these domains we need to be able to solve recursive
domain equations \emph{of error domains}. To do this we simply need to
show that the universe of error predomains is a predomain.

\begin{definition}
  Define $\theta_{\predom} : \later \predom \to \predom$ as follows
  \begin{align*}
    \theta_{\predom}P &= (\later[(X,\ltdynp X) \leftarrow P] X, )
  \end{align*}
\end{definition}

\end{document}