### 2022-08-17 denotational-gradual-typing.tex
sketch of a naive view of gradual types

\documentclass{article}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{tikz-cd}
\usepackage{mathpartir}
\usepackage{stmaryrd}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{nonnum-theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{lemma}{Lemma}[section]

\newcommand{\dynv}{\text{Dyn}^+}
\newcommand{\dync}{\text{Dyn}^-}
\newcommand{\good}[1]{\text{good}_{#1}}
\newcommand{\prop}{\mathbb P}
\newcommand{\with}{\mathrel{\&}}
\newcommand{\eqv}{\simeq}
\newcommand{\soln}{\text{Sol}}
\newcommand{\police}[1]{\text{police}_{#1}}
\newcommand{\later}{{\blacktriangleright}}

\begin{document}

\title{Gradual Types as Subsets}
\author{Max S. New}

\maketitle

We fix a predomain $\dynv$ with poset structure and a domain $\dync$
with bounded below poset structure as solutions to the domain
equations:

\[ \dynv = 2 + \dync \]
\[ \dync = (\dynv \to \dync) \with L (\dynv_{\mho}) \]

Q: does this actually get us the right order?
Q: what domain-theoretic properties does the order have that matter?

There are several ways to define essentially the same interpretation
of CBV gradually typed lambda calculus with booleans into this
semantics:

\begin{enumerate}
\item Types as ((Co-)Kleisli) co-reflections
\item Types as interior operators
\end{enumerate}

I want to explore one that I previously dismissed as somewhat na\"ive:
value types are (well-behaved) subsets of the universal predomain and
computation types are (well-behaved) quotients of the universal
domain.

\begin{enumerate}
\item A value type $A$ is a \emph{policeable} predicate $\good A :
  \dynv \to \prop$
\item A computation type $B$ is a \emph{envelopable} equivalence
  relation $\eqv B : \text{Equivalence}(\dynv)$
\end{enumerate}

Let's explain what these mean.
%
First, define $\soln A$ to be the set of ``good'' elements of $\dynv$
according to $A$, i.e., $\{ d : \dynv \,|\, \good A d \}$.
%
$\good A$ is \emph{policeable} when there exists a function $\police A
: \dynv \to (\soln A)_\mho$ satisfying two properties:
\begin{enumerate}
\item (monotonicity) $\forall d \sqsubseteq d'. \police A d \sqsubseteq \police A d'$
\item (retraction) $\forall x \in \soln A. \police A x = \eta x$
\item (projection) $\forall d \in \dynv. \police d \sqsubseteq \eta d$
\end{enumerate}

If $\sqsubseteq$ is a poset and $\eta : A \to A_\mho$ is order-reflecting,
then $\police A$ is uniquely determined by these two properties, for
if there were another such function $(\police A)'$ we would have
\[ \]

\subsection{Synthetic Guarded Domain Theory}

We can solve the domain equation using synthetic guarded domain
theory.

\begin{align*}
  \dynv &= \mu (\lambda D^+. 2 + \mu (\lambda D^-. L (1 + D^+) \times (D^+ \to D^-) ))\\
  \dync &= \mu (\lambda D^-. L (1 + \later \dynv) \times (\later \dynv \to D^-) )
\end{align*}

unfolding this gives us

\begin{align*}
  \dynv &\cong 2 + \dync \\
  \dync &= L (1 + \later \dynv) \times (\later \dynv \to \later \dync)\\
  &\cong L (1 + \later \dynv) \times \later (\dynv \to \dync)
\end{align*}

\begin{lemma}{Universal Domain}
  $\dync$ is a domain with algebra
  \[ \theta_\dync = \lambda p: \later \dync. (\pi_1p, ) \]
\end{lemma}

Let's define orderings.
\begin{verbatim}
data _<=_ : Dyn+ -> Dyn+ -> Prop where
  bool : forall b : 2. inl b <= inl b

data _<=_ : Dyn- -> Dyn- -> Prop where

err-ord : BinRel a -> BinRel (a + 1)
err-ord _<=_ (inr *)       y = T
err-ord _<=_ (inl x) (inl y) = x <= y
err-ord _<=_ _ _ = _|_

laterr-ord< BinRel a -> BinRel (L (a + 1))
laterr-ord< _<=_ (done (inr *)) _ = T
laterr-ord< _<=_ (done (inl *)) _ = T

laterr-ord> BinRel a -> BinRel (L (a + 1))

err : Dyn-
err = mu (\e. ((inl (inl *)), \x. e))

\end{verbatim}

Let's define

\end{document}