### 2023-04-25 paper-new/graduality.tex
work on paper (changes to intro, work on description of syntax)

\section{Unary Canonicity}
Before discussing graduality, we seek to prove its ``unary'' analogue.
Namely, instead of considering inequality between terms, we start by considering equality.

\section{Graduality}\label{sec:graduality}
The main theorem we would like to prove is the following:

\begin{theorem}[Graduality]
  If $\cdot \vdash M \ltdyn N : \nat$, then
  \begin{enumerate}
    \item If $N = \mho$, then $M = \mho$
    \item If $N = `n$, then $M = \mho$ or $M = `n$
    \item If $M = V$, then $N = V$
  \end{enumerate}
\end{theorem}