### 2023-11-16 paper-new/graduality.tex
begin rearranging: discuss term syn/sem and then ordering syn/sem

\section{Graduality}\label{sec:graduality}
Now we move to the proof of graduality.
The main theorem we would like to prove is the following:

% TODO use the etmprec notation
\begin{theorem}[Graduality]
  If $\cdot \vdash M \ltdyn N : \nat$, then
  \begin{enumerate}
    \item $M \Downarrow$ iff $N \Downarrow$
    \item If $M \Downarrow v_?$ and $N \Downarrow v'_?$ then either $v_? = \mho$, or $v_? = v'_?$.
  \end{enumerate}
\end{theorem}

\subsection{Proof Overview}

We split the proof into three main steps. The first concerns the translation
of the surface syntax $\extlc$ to the intensional syntax $\intlc$.

\begin{lemma}[]
  If $\etmprec{\gamlt}{M}{N}{c}$, then there exist $M'$ and $N'$ such that

  \[ i(M) \synbisim M' \ltdyn_i N' \synbisim i(N). \]
\end{lemma}

Next we show that the notions of intensional syntactic term precision and
syntactic bisimilarity imply their corresponding semantic counterparts:

\begin{lemma}[]
  If $\itmprec{\gamlt}{M}{N}{c}$, then $\sem{M} \semlt {\sem{N}}$
\end{lemma}

\begin{lemma}
  If $M \synbisim N$, then $\sem{M} \bisim \sem{N}$.
\end{lemma}

Finally, we need an adequacy result stating that the semantic notions of error
ordering and bisimilarity for natural numbers imply that the values of the
lift monad that they relate have the expected behavior.

\begin{lemma}
  Let $l, l' : \li \Nat$.
  If $l \semlt l'$, then either (1) $l = \delta^n(\mho)$ for some $n$, or (2) 
  $l = l' = \delta^n(\eta m)$ for some $n$ and $m$, or (3) $l$ and $l'$ fail to terminate.
\end{lemma}

\begin{lemma}
  If $l \bisim l'$, then either (1) $l = \delta^n(\mho)$ and $l' = \delta^{n'}(\mho)$ for some $n$ and $n'$,
  (2) $l = \delta^n(\eta m)$ and $l' = \delta^{n'}(\eta m)$, or $l$ and $l'$ fail to terminate.
\end{lemma}