### 2024-01-24 paper-new/appendix.tex
work on proofs in appendix

\section{Gradual Typing Syntax}

\max{decide if we want to include $\times$ or not}

\begin{align*} % TODO is hole a term?
  &\text{Types } A := \dyn \alt \nat \alt A \times A \alt (A \ra A') \\
  &\text{Value Contexts } \Gamma := \cdot \alt (\Gamma, x : A) \\
  &\text{Terms } M, N := \upc c M \alt \dnc c M
  &\quad\quad \alt \ret {V} \alt \bind{x}{M}{N} \alt V_f\, V_x \alt \dn{A}{B} M 
\end{align*}

The type precision derivations $c : A \ltdyn A'$ is inductively defined by
reflexivity, transitivity, congruence for $\ra$ and $\times$, and generators
$\textsf{Inj}_\ra : (D \ra D) \ltdyn D$ and $\textsf{Inj}_{\text{nat}}
: \nat \ltdyn D$.
%
We define equivalence of type precision derivations to be inductively generated by congruence for all constructors, category laws for reflexivity and transitivity as well as functoriality laws for $\ra$ and $\times$ congruence
\[ (c_i \ra c_o)(c_i' \ra c_o) = (c_ic_i') \ra (c_oc_o') \]
\[ (c_1 \times c_2)(c_1' \times c_2') = (c_1c_1') \times (c_2c_2') \]

\begin{theorem}
  \begin{enumerate}
  \item For every $A$, there is a derivation $c : A \ltdyn D$
  \item Any two derivations $c,c' : A \ltdyn A'$ of the same precision
    are equivalent.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \begin{enumerate}
  \item See prior work \cite{prior-work}
  \item We show this by showing that derivations have a canonical
    form.

    The following presentation of precision derivations has unique derivations
    \begin{mathpar}
      \inferrule{}{\textrm{refl}(D) : D \ltdyn D}
      \inferrule{}{\textsf{Inj}_{\text{nat}} : \nat \ltdyn D}
      \inferrule{}{\textrm{refl}(\nat) : \nat \ltdyn \nat}
      \inferrule{c : A_i \ra A_o \ltdyn D\ra D}{c(\textsf{Inj}_{\text{arr}}) : A_i \ra A_o \ltdyn \nat}
      \inferrule{c : A_i \ltdyn A_i' \and d : A_o \ltdyn A_o'}{c \ra d : A_i \ra A_o \ltdyn A_i'\ra A_o'}
      %% \inferrule{A_1 \times A_2 \ltdyn D\times D}{A_1 \times A_2 \ltdyn \nat}
      %% \inferrule{A_1 \ltdyn A_1' \and A_2 \ltdyn A_2'}{A_1 \times A_2 \ltdyn A_1'\times A_2'}
    \end{mathpar}
    Since it satisfies reflexivity, cut-elimination and congruence, it
    is a model of the original theory. Since it is a sub-theory of the
    original theory, it is equivalent.
  \end{enumerate}
\end{proof}

The gradually typed lambda calculus we consider is call-by-value
gradual lambda calculus with $\ra$ and $\nat$ as the only base types.
Casts are generated by the rule
\begin{mathpar}
  \inferrule
  {\Gamma \vdash M : A \and c : A \ltdyn A'}
  {\Gamma \vdash \upc c M : A'}

  \inferrule
  {\Gamma \vdash M : A' \and c : A \ltdyn A'}
  {\Gamma \vdash \dnc c M : A}
\end{mathpar}

Type precision is a binary relation on typed terms. The original
gradual guarantee rules are as follows:
\begin{mathpar}
  \inferrule
  {\Gamma^\ltdyn \vdash M \ltdyn M' : c \and
    c : A \ltdyn A'\and
    c' : A \ltdyn A_2'
  }
  {\Gamma^\ltdyn \vdash M \ltdyn (M :: A_2') : c'}

  \inferrule
  {\Gamma^\ltdyn \vdash M \ltdyn M' : c \and
    c : A \ltdyn A'\and
    c' : A_2 \ltdyn A'
  }
  {\Gamma^\ltdyn \vdash (M :: A_2) \ltdyn M' : c'}
\end{mathpar}
These two rules are admissible from the following principles:

%% \begin{mathpar}
%%   \inferrule
%%   {}
%%   {\dnc {\injarr{}} \upc {\injarr{}} M \equidyn M}

%%   %% \inferrule
%%   %% {}
%%   %% {}
%% \end{mathpar}

\section{Call-by-push-value}

In CBPV models, all the type constructors are interpreted as functors:
\begin{enumerate}
\item $\to : \op\calV \times \calE \to \calE$
\item $\times : \calV \times \calV \to \calV$
\item $F : \calV \to \calE$
\item $U : \calE \to \calV$
\end{enumerate}
That is, they all have functorial actions on \emph{pure} morphisms of
value types and \emph{linear} morphisms of computation types.
%
We use these functorial actions extensively in the construction of
casts and their corresponding perturbations. But when defining
downcasts of value types and upcasts of computation types, we
additionally need a second functorial action of these categories:
functoriality in \emph{impure} morphisms of value types and
\emph{non-linear} morphisms of computation types. These notions of
morphism are given by the \emph{Kleisli} categories $\calVk$ and
$\calEk$ which have value types and computation types as objects but
morphisms are defined as
\[ \calVk(A,A') = \calE(F A, FA')\]
\[ \calEk(B,B') = \calV(U B, U B')\]
with composition given by composition in $\calE/\calV$.  That is we
need to define a second functorial action, that agrees with the above
on objects for these Kleisli categories:
\begin{enumerate}
\item $\tok : \op\calVk \square \calEk \to \calEk$
\item $\timesk : \calVk \square \calVk \to \calVk$
\item $\Fk : \calVk \to \calEk$
\item $\Uk : \calEk \to \calVk$
\end{enumerate}
Note that rather than the product of categories we use the ``funny
tensor product'' $\square$. This is because the action on
impure/non-linear morphisms for $\tok/\timesk$ do not satisfy ``joint
functoriality'' but instead only ``separate functoriality'', meaning
we give rather than an action on morphisms in both categories
simultaneously instead an action on each argument categories morphisms
with the object in the other category fixed. The existence of these
functorial actions for $\tok$ and $\timesk$ is reliant on the
\emph{strength} of the adjunction. We describe them using the internal
language of CBPV in order to more easily verify their
existence/functoriality:
\begin{enumerate}
\item For $\tok$ we define for $\phi : \calE(F A,F A')$ and $B \in \calE$ the morphism $\phi \tok B : \calV(U(A' \to B),U(A\to B))$ as
  \[ t:U(A'\to B) \vdash \phi \tok B = \{ \lambda x. x' \leftarrow \phi\,[\ret x]; ! t x'\} : U(A \to B) \]
  and for $A \in \calV$ and $f : \calV(UB,UB')$ we define $A \tok f : \calV(U(A \to B),U(A\to B'))$ as
  \[ t : U(A \to B) \vdash A \tok f = \{ \lambda x. !f[\{ ! t x \}]\} \]
\item For $\timesk$ we define for $\phi : \calE(F A_1,FA_2)$ and $A' \in \calV$ the morphism $\phi \timesk A_2$ as
  \[ \bullet : F(A_1\times A_2) \vdash \phi \timesk A_2 = (x_1,x_2) \leftarrow \bullet; x_1' \leftarrow \phi[\ret x_1]; \ret (x_1',x_2) : F(A_1'\times A_2)\]
  and $A_1 \timesk \phi$ is defined symmetrically.
\item For $\Uk$ we need to define for $f : \calV(UB,UB')$ a morphism $\Uk f : \calE(FUB,FUB')$. This is simply given by the functorial action of $F$: $\Uk f = F(f)$
\item Similarly $\Fk \phi = U\phi$
\end{enumerate}

Functoriality in each argument is easily established, meaning for
example for the function type is functorial in each argument:
\begin{enumerate}
\item $(\phi \circ \phi') \tok B = (\phi' \tok B) \circ (\phi \tok B)$
\item $\id \tok B = \id$
\item $A \tok (f \circ f') = (A \tok f) \circ (A \tok f)$
\item $A \tok \id = \id$
\end{enumerate}

Finally, note that all of these constructions lift to squares in a
double CBPV model since the squares themselves form a CBPV model and
the projection functions preserve CBPV structure. For instance, given a square
$\alpha : \phi \ltdyn_{F c_o}^{F c_i} \phi'$ and a horizontal morphism $d : B \rel B'$ of appropriate type, we get a square
\[ \alpha \tok d : \phi \tok B \ltdyn_{U(c_o \to d)}^{U(c_i \to d)} \phi' \tok B' \]

\section{Details of the Construction of an Extensional Model}

In Section \ref{sec:extensional-model-construction}, we outline the construction
of an extensional model of gradual typing starting from a step-1 intensional model.
In this section, we provide the details for each of the constructions mentioned there.

\subsection{Constructing a Model with Perturbations}

The goal of this section is to prove the following lemma:

\begin{lemma}\label{lem:step-1-model-to-step-2-model}
Let $\mathcal M$ be a \hyperref[def:step-1-model]{step-1 intensional model}.

Then we can construct a \hyperref[def:step-2-model]{step-2 intensional model}.
\end{lemma}

We begin with a definition and some lemmas that will be useful in the construction
of the model.

\begin{definition}
  Let $c : A \rel A'$ be a value relation of $\mathcal M$.
  Let $P_A$ be a monoid of perturbations on $A$ and $P_{A'}$ a monoid of perturbations on $A'$.
  A \emph{push-pull structure} $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$ consists of:
  \begin{itemize}
    \item A function $\push : P_A \to P_{A'}$ 
    such that for all $\delta^l \in P_A$ we have $\delta^l \ltdyn_c^c \push(\delta^l)$.
    \item A function $\pull : P_{A'} \to P_A$
    such that for all $\delta_r \in P_{A'}$ we have $\pull(\delta^r) \ltdyn_{c}^c \delta^r$.
  \end{itemize}

  We define a push-pull structure $\pie_d$ for $d : B \rel B'$ with respect to perturbation monoids
  $P_B$ and $P_{B'}$ in an analogous manner.

  % \eric{do we need that $\delta^*$ is preserved by push and pull?}
  % We stipulate the additional requirement that if $\delta_A^*$ is in $P_B$ 
  % (where $\delta_A^*$ is the distinguished perturbation required to be in every $\ef(FA, FA)$),
  % then $\push(\delta_A^*) = \delta_{A'}^*$ and likewise $\pull(\delta_{A'}^*) = \delta_A^*$.
\end{definition}

%%%%%%%%%%%%%%%
% Composition %
%%%%%%%%%%%%%%%
\begin{lemma}
  Let $c : A \rel A'$ and $c' : A' \rel A''$ be value relations,
  and let $P_A, P_{A'}$ and $P_{A''}$ be monoids of perturbations.
  Given a push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$, and
  and a push-pull structure $\piv_{c'}$ for $c'$ with respect to $P_{A'}$ and $P_{A''}$,
  we can define a push-pull structure $\piv_{c \comp c'}$ for $c \comp c'$.

  Likewise, we can define a push-pull structure for the composition of computation relations.
\end{lemma}
\begin{proof}
  We define $\piv_{c \comp c'}$ as the following push-pull structure:

  \begin{itemize}
    \item $\push_{c \comp c'} = \push_{c'} \circ \push_{c}$
    \item $\pull_{c \comp c'} = \pull_{c} \circ \pull_{c'}$
  \end{itemize}

  We observe that the required squares exist for both push and pull.
  In particular, for push we have that $\delta^l \ltdyn_c^c \push_c(\delta^l)$
  using the push property for $c$, and then using the push property for $c'$ we have
  $\push_c(\delta^l) \ltdyn_{c'}^{c'} \push_{c'}(\push_c(\delta^l))$.
  We can then compose these squares horizontally to obtain the desired square.
  The pull property follows similarly.

  The push-pull structure for the composition of computation relations is defined analogously.
\end{proof}

%%%%%%%%%%%%%
% UF and FU %
%%%%%%%%%%%%%
\begin{lemma}
  Let $c : A \rel A'$ and let $P_A$ and $P_{A'}$ be monoids of perturbations.
  Given a push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$,
  we can define a push-pull structure $\pie_{UF(c)}$ for $UF(c)$ with respect to $UF(P_{A})$
  and $UF(P_{A'})$.
  % Eric: Are we assuming here that F and U preserve perturbations?
\end{lemma}
\begin{proof}
  Follows by the functorial action of $UF$ and $FU$ on perturations and squares.
\end{proof}

%%%%%%%%%%%%
% Products %
%%%%%%%%%%%%
\begin{lemma}
  Let $c_1 : A_1 \rel A_1'$ and $c_2 : A_2 \rel A_2'$, and let $P_{A_1}$,
  $P_{A_2}$, $P_{A_1'}$ and $P_{A_2'}$ be monoids of perturbations.
  Given push-pull structures 
  $\piv_{c_1}$ with respect to $P_{A_1}$ and $P_{A_1'}$, and 
  $\piv_{c_2}$ with respect to $P_{A_2}$ and $P_{A_2'}$,
  we can define a push-pull structure $\piv_{c_1 \times c_2}$
  with respect to $P_{A_1} \times P_{A_2}$ and $P_{A_1'} \times P_{A_2'}$.

  Likewise, given push-pull structures 
  $\pie_{Fc_1}$ with respect to $F(P_{A_1})$ and $F(P_{A_1'})$, and
  $\pie_{Fc_2}$ with respect to $F(P_{A_2})$ and $F(P_{A_2'})$,
  we can define a push-pull structure $\pie_{F(c_1 \times c_2)}$
  with respect to $F(P_{A_1} \times P_{A_2})$ and $F(P_{A_1'} \times P_{A_2'})$.
\end{lemma}
\begin{proof}
We define the push function for $\piv_{c_1 \times c_2}$ by 
$\push(\delta_1, \delta_2) = (\push_{c_1}(\delta_1), \push_{c_2}(\delta_2))$ and likewise for $\pull$.
The push-pull property holds because it holds for each component.
\end{proof}

%%%%%%%%%
% Arrow %
%%%%%%%%%
\begin{lemma}
  Let $c : A \rel A'$ and $d : B \rel B'$.
  Given push-pull structures $\piv_c$ for $c$ and $\pie_d$ for $d$,
  we can define a push-pull structure $\pie_{c \arr d}$ for $c \arr d$.

  Given push-pull structures $\pie_{Fc}$ for $Fc$ and $\piv_{Ud}$ for $Ud$,
  we can define a push-pull structure $\piv_{U(c \arr d)}$ for $U(c \arr d)$
  with respect to the monoids $(P_{FA})^{op} \oplus P_B$
\end{lemma}
\begin{proof}
  The former follows from the functorial action of $\arr$ on perturbations and squares,
  and the latter uses the functorial action of $\tok$ on perturbations and squares.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%
% Model Construction %
%%%%%%%%%%%%%%%%%%%%%%
We now proceed with the construction of the model:
    % Write 
    % %
    % \[ \mathcal M = (\vf, \vsq, \ef, \esq, \Ff, \Fsq, \Uf, \Usq, \arrf, \arrsq). \] 
    % %

    Define a step-2 model $\mathcal M'$ as follows:
    \begin{itemize}
      \item Value objects are triples consisting of:
      \begin{itemize}
        \item A value object $A$ in $\vf$ 
        \item A monoid $P_A$ of ``pure'' perturbations that is a submonoid of 
        $\{ f \in \vf(A, A) \mid f \bisim \id_A \}$
        \item A monoid $P^K_A$ of ``impure'' or ``Kleisli'' perturbations that is a submonoid of
        $\{ \phi \in \ef(FA, FA) \mid \phi \bisim \id_{FA} \}$ such that $P^K_A$ contains
        $\delta_A^*$
      \end{itemize}

      \item Computation objects are triples consisting of:
      \begin{itemize}
        \item A computation object $B$ in $\ef$
        \item A monoid $P_B$ of ``pure'' perturbations that is a submonoid of
        $\{ \phi \in \ef(B, B) \mid \phi \bisim \id_B \}$
        \item A monoid $P^K_B$ of ``impure'' or ``Kleisli'' perturbations that is a submonoid
        of $\{ g \in \vf(UB, UB) \mid g \bisim \id_{UB} \}$
      \end{itemize}

      \item Morphisms are given by morphisms of the underlying objects in $\vf$ and $\ef$, respectively.
      %, i.e.,
      % \[ \vf'((A, P_A, \ptb_A, P^K_A, \ptbk_A), (A', P_{A'}, \ptb_{A'}, P^K_{A'}, \ptbk_{A'})) = \vf(A, A') \]
      %
      % and likewise for computations.

    \end{itemize}

    % Before introducing the relations, we make a definition.

    % \begin{definition}[push-pull structure]
    %   Let $c : A \rel A'$ be a value relation of $\mathcal M$. A \emph{value push-pull structure} $\piv_c$ for $c$ consists of:
    %   \begin{itemize}
    %     \item A function $\push : P_A \to P_{A'}$ 
    %           such that for all $\delta^l \in P_A$ we have $\delta^l \ltdyn_c^c \push(\delta^l)$.
    %     \item A function $\push^K : P^K_A \to P^K_{A'}$ 
    %           such that for all $\delta^K_l \in P^K_A$ we have $\delta^K_l \ltdyn_{Fc}^{Fc} \push(\delta^K_l)$.
    %     \item A function $\pull : P_{A'} \to P_A$
    %           such that for all $\delta_r \in P_{A'}$ we have $\pull(\delta^r) \ltdyn_{c}^c \delta^r$.
    %     \item A function $\pull^K : P^K_{A'} \to P^K_A$
    %           such that for all $\delta^K_r \in P^K_{A'}$ we have $\pull(\delta^K_r) \ltdyn_{Fc}^{Fc} \delta^K_r$.
    %   \end{itemize}

    %   For $d : B \rel B'$ a computation relation, we define a \emph{computation push-pull structure} $\pie_d$ for $d$
    %   in an analogous manner.
    % \end{definition}

    % Now we continue with the description of the construction:
    \begin{itemize}

      \item The objects of $\vsq'$ (i.e., the value relations) are triples consisting of:
      \begin{itemize}
        \item A value relation $c \in \vsq$.
        \item A push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$.
        \item A push-pull structure $\pie_{Fc}$ for $Fc$ with respect to $P^K_A$ and $P^K_{A'}$.
      \end{itemize}

      The objects of $\esq'$ are defined analogously, i.e.,
      \begin{itemize}
        \item A computation relation $d \in \esq$.
        \item A push-pull structure $\pie_d$ for $d$ with respect to $P_B$ and $P_{B'}$.
        \item A push-pull structure $\piv_{Ud}$ for $Ud$ with respect to $P^K_B$ and $P^K_{B'}$.
      \end{itemize}

      \item The morphisms of $\vsq'$ and $\esq'$ are given by the morphisms of $\vsq$ and $\esq$.

      \item We define composition of relations $(c, \piv_c, \pie_{Fc})$ and 
      $(c', \piv_{c'}, \pie_{Fc'})$ as $(c \comp c', \piv_{c \comp c'}, \pie_{Fc \comp Fc'})$
      and likewise for computation relations.

    \end{itemize}

      % Functors \times, +, F, U, arrow
      Now we define the actions of the functors:
      \begin{itemize}

      % \item We begin with the functor $m$ that composes relations and squares horizontally.
      % Given value relations $(c ,\piv_c, \pie_{Fc})$ and $(c', \piv_{c'}, \pie_{Fc'})$ where $c : A \rel A'$ and $c' : A' \rel A''$,
      % we define their composition to be the triple $(c \comp c', \piv_{c \comp c'})$ where we define
      % $\piv_{c \comp c'}$ as the following push-pull structure:

      % \begin{itemize}
      %   \item $\push_{c \comp c'} = \push_{c'} \circ \push_{c}$
      %   \item $\push^K_{c \comp c'} = \push^K_{c'} \circ \push^K_{c}$
      %   \item $\pull_{c \comp c'} = \pull_{c} \circ \pull_{c'}$
      %   \item $\pull^K_{c \comp c'} = \pull^K_{c} \circ \pull^K_{c'}$
      % \end{itemize}

      % We observe that each of the components satisfies the commuting necessary square.

      \item We define $\times$ on objects by

      \[ (A_1, P_{A_1}, P^K_{A_1}) \times (A_2, P_{A_2}, P^K_{A_2}) =
        (A_1 \times A_2, P_{A_1} \times P_{A_2}, P^K_{A_1} \oplus P^K{A_2}). \]

      where $\oplus$ is the coproduct in the category of monoids.

      We define $\times$ on relations by

      \[ (c_1, \piv_{c_1}, \pie_{Fc_1}), (c_2 \piv_{c_2}, \pie_{Fc_2}) = 
        (c_1 \times c_2, \piv_{c_1 \times c_2}, \pie_{F(c_1 \times c_2)}). \]

      \item We define $F$ on objects by

      \[ F(A, P_A, P^K_A) = (FA, P^K_A, UF(P_A)). \]

      We define $F$ on relations by

      \[ F(c, \piv_c, \pie_{Fc}) = (Fc, \pie_{Fc}, \piv_{UF(c)}). \]

      \item We define $U$ on objects by

      \[ U(B, P_B, P^K_B) = (UB, P^K_B, FU(P_B). )\]

      We define $U$ on relations by

      \[ U(d, \pie_d, \piv_{Ud}) = (Ud, \piv_{Ud}, \pie_{FU(d)})  \]

      We define $\arr$ on objects by

      \[ (A, P_A, P^K_A) \arr (B, P_B, P^K_B) = 
        (A \arr B, (P_A)^{op} \times P_B, (P^K_A)^{op} \oplus P^K_B). \]

      We define $\arr$ on relations by

      \[ (c, \piv_c, \pie_{Fc}) \arr (d, \pie_d, \piv_{Ud}) =
        (c \arr d, \pie_{c \arr d}, \piv_{U(c \arr d)}). \]

      % \item We define $F$ on objects by $F (A, \pv_A, \ptbv_A) = (FA, (1 + \pv_A), h_F)$
      % where $1$ is the trivial monoid, $+$ is the coproduct in the category of monoids, and $h_F$ is the homomorphism defined as follows:

      % \item We define $U$ on objects by $U (B, \pe_B, \ptbe_B) = (UB, \pe_B, h_U)$
      % where $h_U(p_B) = U(\ptbe_B(p_B))$.

      % \item We define $(A, \pv_A, \ptbv_A) \arr (B, \pe_B, \ptbe_B) = (A \arr B, \pv_A \times \pe_B, h_\arr)$
      % where $\times$ is the product in the category of monoids, and $h_\arr$ is defined by 
      % $h_\arr(p_A, p_B) = \ptbv_A(p_A) \arr \ptbe_B(p_B)$.
    \end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Constructing a Model with Quasi-Representable Relations}

The goal of this section is to prove the following lemma:

\begin{lemma}\label{lem:step-2-model-to-step-3-model}
  Let $\mathcal M$ be a \hyperref[def:step-2-model]{step-2 intensional model}.

  Then we can construct a \hyperref[def:step-3-model]{step-3 intensional model}.
\end{lemma}

Before proceeding with the proof, we begin with a definition.

\begin{definition}[representation structure]
Let $c : A \rel A'$ be a value relation. A \emph{left-representation structure}
$\rho^L_c$ for $c$ consists of a value morphism $e_c \in \vf(A, A')$ such that
$c$ is quasi-left-representable by $e_c$ (see Definition \ref{def:quasi-left-representable}).

Likewise, let $d : B \rel B'$ be a computation relation. A \emph{right-representation structure}
$\rho^R_d$ for $d$ consists of a computation morphism $p_d \in \ef(B', B)$
such that $d$ is quasi-right-representable by $p_d$ (see Definition
\ref{def:quasi-right-representable}).
\end{definition}

(Notice that the direction of the morphism is opposite in the definition of
right-representation structure.)

% TODO establish the notation

%%%%%%%%%%%%%%%
% Composition %
%%%%%%%%%%%%%%%
\begin{lemma}\label{lem:representation-comp}
  In the below, let $c : A \rel A'$ and $c' : A' \rel A''$ and $d : B \rel B'$ and $d' : B' \rel B''$.
  \begin{enumerate}
    \item Given left-representation structures $\rho^L_c$ for $c$ and $\rho^L_{c'}$ for $c'$,
    we can define a left-representation structure for the composition $c \comp c'$
    \item Given right-representation structures $\rho^R_d$ for $d$ and $\rho^R_{d'}$ for $d'$,
    we can define a right-representation structure for the composition $d \comp d'$
    \item Given right-representation structures $\rho^R_{Fc}$ for $Fc$ and $\rho^R_{Fc'}$ for $Fc'$,
    we can define a right-representation structure $\rho^R_{F(c \comp c')}$ for $F(c \comp c')$.
    \item Given left-representation structures $\rho^L_{Ud}$ for $Ud$ and $\rho^L_{Ud'}$ for $Ud'$,
    we can define a left-representation structure $\rho^L_{U(d \comp d')}$ for $U(d \comp d')$.
  \end{enumerate}
  \eric{Do we need parts 3 and 4?}

\end{lemma}
\begin{proof}
  \begin{enumerate}

    % 1.
    \item We define $\rho^L_{c \comp c'}$ as follows. In the definitions of the perturbations,
    we make use of the fact that $c$ and $c'$ satisfy the push-pull property.
    \begin{itemize}
      \item $e_{c \comp c'} = e_{c'} \circ e_c$
      \item $\delre_{c \comp c'} = \delre_{c'} \circ \push_{c'}(\delre_c)$
      \item $\delle_{c \comp c'} = \pull_c(\delle_{c'}) \circ \delle_c$
      \item $\upl$ is the following square: \input{squares/UpL-comp}

      The square $(*)$ exists by the push-pull property for $c'$, and the square $(**)$ exists
      because $r$ is a unit for horizontal composition, so $r(A') \comp c' = c'$, and so this
      is simply the identity square $\id_{c'} \in \vsq(c', c')$.

      \item $\upr$ is the following square: \input{squares/UpR-comp}
    \end{itemize}

    % 2.
    \item We define $\rho^R_{d \comp d'}$ as follows:
    \begin{itemize}
      \item $p_{d \comp d'} = p_d \circ p_{d'}$
      \item $\dellp_{d \comp d'} = \dellp_d \circ \pull_d(\dellp_{d'})$
      \item $\delrp_{d \comp d'} = \push_{d'}(\delrp_d) \circ \delrp_{d'}$
      \item $\dnr$ is the following square: \input{squares/DnR-comp}

      Here the square $(*)$ exists by the push-pull property for $d$.

      \item $\dnl$ is the following square: \input{squares/DnL-comp}
    \end{itemize}

    % 3.
    \item We define $\rho^R_{F(c \comp c')}$ as follows:
    \begin{itemize}
      \item $p_{F(c \comp c')} = p_{Fc} \circ p_{Fc'}$ % : FA'' \to FA
      \item $\dellp_{F(c \comp c')} = $
      \item $\delrp_{F(c \comp c')} = $
    \end{itemize}

    % 4.
    \item We define $\rho^L{(U(d \comp d'))}$ as follows:

  \end{enumerate}
\end{proof}

% Want to show: U(d \comp d') is weakly equivalent to U(d) \comp U(d').
% This holds because both are quasi-representable by the same projection
% \begin{lemma}
%   Let $\mathcal M$ be a step-3 intensional model, and let
%   $d : B \rel B'$ and $d' : B' \rel B''$.
%   Let $\rho_\text{comp}$ be an arbitrary right-representation for $d \comp d'$.
%   Then the projection $p_{d \circ d'}$ is weakly equivalent to
% \end{lemma}

%%%%%%%%%%%%%
% UF and FU %
%%%%%%%%%%%%%
\begin{lemma}\label{lem:representation-UF-FU}
  Let $c : A \rel A'$, and let $\rho^L_c$ be a left-representation structure for $c$.
  Then we can define a left-representation structure $\rho^L_{UF(c)}$ for $UF(c)$.

  Similarly, let $d : B \rel B'$ and let $\rho^R_d$ be a right-representation structure for $d$.
  Then we can define a right-representation structure $\rho^R_{FU(d)}$ for $FU(d)$.
\end{lemma}
\begin{proof}
  We define $\rho^L_{UF(c)}$ as follows:
  \begin{itemize}
    \item $e_{UF(c)} = UF(e_c)$
    \item $\delre_{UF(c)} = UF(\delre_c)$ (which is in the monoid of perturbations of $UF(A')$ because the
    perturbation monoids are closed under the actions of the functors $F$ and $U$)
    \item $\delle_{UF(c)} = UF(\delle_c)$
    \item We get the two commuting squares by the functorial action of $UF$ on the two squares for $c$,
    i.e., $\upr_{UF(c)} = UF(\upr_c)$
  \end{itemize}

  We define $\rho^R_{FU(d)}$ in a similar manner.
\end{proof}

%%%%%%%%%%%%
% Products %
%%%%%%%%%%%%
\begin{lemma}\label{lem:representation-product}
  Let $c_1 : A_1 \rel A_1'$ and $c_2 : A_2 \rel A_2'$.
  Let $\rho^L_{c_1}$ be a left-representation structure for $c_1$, and
  let $\rho^L_{c_2}$ be a left-representation structure for $c_2$.
  Then we can define a left-representation structure for $c_1 \times c_2$.

  Likewise, let $\rho^R_{Fc_1}$ and $\rho^R_{Fc_2}$ be right-representation
  structures for $Fc_1$ and $Fc_2$ respectively.
  Then we can deinfe a right-representation structure for $F(c_1 \times c_2)$.
\end{lemma}
\begin{proof}
  We define $\rho^L_{c_1 \times c_2}$ as follows:
  \begin{itemize}
    \item $e_{c_1 \times c_2} = e_{c_1} \times e_{c_2}$
    \item $\delre_{c_1 \times c_2} = \delre_{c_1} \times \delre_{c_2}$ and
    likewise for $\delle_{c_1 \times c_2}$
    \item We get the commuting squares via the functorial action of $\times$
    on the corresponding squares for $c_1$ and $c_2$.
  \end{itemize}

  % TODO check this
  We define $\rho^R_{F(c_1 \times c_2)}$ as follows:
  \begin{itemize}
    \item $p_{F(c_1 \times c_2)} = (p_{Fc_1} \timesk A_2) \circ (A_1' \timesk p_{Fc_2})$
    \item $\dellp_{F(c_1 \times c_2)} = (\dellp_{Fc_1} \timesk A_2) \circ (A_1 \timesk \dellp_{Fc_2})$
    \item $\delrp_{F(c_1 \times c_2)} = (\delrp_{Fc_1} \timesk A_2') \circ (A_1' \timesk \delrp_{Fc_2})$
    \item The commuting squares are obtained via the functorial action of $\timesk$ on the squares for $Fc_1$ and $Fc_2$.
    % TODO elaborate?
  \end{itemize}
\end{proof}

%%%%%%%%%
% Arrow %
%%%%%%%%%
\begin{lemma}\label{lem:representation-arrow}
  Let $c : A \rel A'$ and $d : B \rel B'$.
  Let $\rho^L_c$ be a left-representation structure for $c$,
  and let $\rho^R_d$ be a right-representation structure for $d$.
  Then we can define a right-representation structure for $c \arr d$.

  Likewise, let $\rho^R_{Fc}$ be a right-representation structure for $Fc$,
  and let $\rho^L_{Ud}$ be a left-representation structure for $Ud$.
  Then we can define a left-representation structure for $U(c \arr d)$.
\end{lemma}
\begin{proof}
  We define $\rho^R_{c \arr d}$ as follows:
  \begin{itemize}
    \item $p_{c \arr d} = e_c \arr p_d \in \ef(A' \arr B', A \arr B)$ 
    (using the functorial action of $\arr$ on morphisms).
    \item $\dellp_{c \arr d} = \delle_c \arr \dellp_d$
    \item $\delrp_{c \arr d} = \delre_c \arr \delrp_d$
    \item The squares $\dnr$ and $\dnl$ are obtained via the functorial action of $\arr$, i.e., we define

    \[ \dnr_{c \arr d} = \upr_{c} \arr \dnr_{d} : 
      (\delle_c \arr \dellp_d) \ltdyn_{r(A \arr B)}^{c \arr d} (e_c \arr p_d),  \]

    and

    \[ \dnl_{c \arr d} = \upl_{c} \arr \dnl_{d}. \]

  \end{itemize}

  % TODO check this
  We define $\rho^L_{U(c \arr d)}$ as follows:
  \begin{itemize}
    \item $e_{U(c \arr d)} = (p_{Fc} \tok B') \circ (A \tok e_{Ud})$
    \item $\delre_{U(c \arr d)} = (\delrp_{Fc} \tok B') \circ (A' \tok \delre_{Ud})$
    \item $\delle_{U(c \arr d)} = (\dellp_{Fc} \tok B) \circ (A \tok \delle_{Ud})$
    \item The squares $\upl$ and $\upr$ are obtained via the functorial action of $\tok$.
    For instance, $\upl$ is given by the following square:

    % https://q.uiver.app/#q=WzAsNixbMCwwLCJVKEEgXFx0byBCKSJdLFsyLDAsIlUoQScgXFx0byBCJykiXSxbMiwxLCJVKEEnIFxcdG8gQicpIl0sWzIsMiwiVShBJyBcXHRvIEInKSJdLFswLDEsIlUoQSBcXHRvIEInKSJdLFswLDIsIlUoQScgXFx0byBCJykiXSxbMCwxLCJVKGMgXFx0byBkKSJdLFs0LDIsIlUoYyBcXHRvIHIoQicpKSJdLFs1LDMsIlUocihBJykgXFx0byByKEInKSkiXSxbNCw1LCJwX3tGY30gXFx0b2sgQiciLDJdLFsxLDIsIkEnIFxcdG9rIFxcZGVscmVfe1VkfSJdLFsyLDMsIlxcZGVscnBfe0ZjfSBcXHRvayBCJyJdLFswLDQsIkEgXFx0b2sgZV97VWR9IiwyXSxbMTIsMTAsIlxcaWRfe0ZjfSBcXHRvayBcXHVwbF97VWR9IiwxLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzksMTEsIlxcZG5sX3tGY30gXFx0b2sgXFxpZF97cihCJyl9IiwxLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}[ampersand replacement=\&,row sep=large]
      {U(A \to B)} \&\& {U(A' \to B')} \\
      {U(A \to B')} \&\& {U(A' \to B')} \\
      {U(A' \to B')} \&\& {U(A' \to B')}
      \arrow["{U(c \to d)}", from=1-1, to=1-3]
      \arrow["{U(c \to r(B'))}", from=2-1, to=2-3]
      \arrow["{U(r(A') \to r(B'))}", from=3-1, to=3-3]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{p_{Fc} \tok B'}"', from=2-1, to=3-1]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{A' \tok \delre_{Ud}}", from=1-3, to=2-3]
      \arrow[""{name=2, anchor=center, inner sep=0}, "{\delrp_{Fc} \tok B'}", from=2-3, to=3-3]
      \arrow[""{name=3, anchor=center, inner sep=0}, "{A \tok e_{Ud}}"', from=1-1, to=2-1]
      \arrow["{\id_{Fc} \tok \upl_{Ud}}"{description}, draw=none, from=3, to=1]
      \arrow["{\dnl_{Fc} \tok \id_{r(B')}}"{description}, draw=none, from=0, to=2]
    \end{tikzcd}\]
  \end{itemize}

  The construction of $\upr$ is similar.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%
% Model Construction %
%%%%%%%%%%%%%%%%%%%%%%
Now we can give the proof of the main lemma:
  % Write 
  % %
  % \[ \mathcal M = (\vf, \vsq, \ef, \esq, \Ff, \Fsq, \Uf, \Usq, \arrf, \arrsq). \] 
  % %

  We define a step-3 model $\mathcal M'$ as follows:
  \begin{itemize}
    \item The objects of $\mathcal M'$ are defined to be the same as the objects of $\mathcal M$.
    \item The value and computation morphisms in $\mathcal M'$ are the same as those of $\mathcal M$.
    \item A value relation is defined to be a tuple $(c, \rho^L_c, \rho^R_{Fc})$ with
    \begin{itemize}
      \item $c$ a value relation in $\mathcal M$,
      \item $\rho^L_c$ a left-representation structure for $c$, and
      \item $\rho^R_{Fc}$ a right-representation structure for $Fc$
    \end{itemize}
    \item Likewise, a computation relation is defined to be a tuple $(d, \rho^R_d, \rho^L_{Ud})$ with
    \begin{itemize}
      \item $d$ a computation relation in $\mathcal M$,
      \item $\rho^R_d$ a right-representation structure for $d$, and
      \item $\rho^L_{Ud}$ a left-representation structure for $Ud$.
    \end{itemize}
    \item Morphisms of value relations (i.e., the value squares) are defined by simply
    ignoring the representation structures. That is, a morphism of value relations
    $\alpha \in \vsq'((c, \rho^L_c, \rho^R_{Fc}), (c' \rho^L_{c'}, \rho^R_{Fc'}))$ is simply a morphism of value
    relations in $\vsq(c, c')$. Likewise for computations.
  \end{itemize}

We define the functor $m$ which composes relations and squares horizontally as follows:
Let $c : A \rel A'$ and $c' : A' \rel A''$. We let

\begin{align*} 
  m ((c, \rho^L_c, \rho^R_{Fc}), (c', \rho^L_{c'}, \rho^R_{Fc'})) = \\
  (c \comp c', \rho^L_{c \comp c'}, \rho^R_{F(c \comp c')}). 
\end{align*}

Now we define the functors $F$, $U$, $\times$, and $\arr$.
On objects, the behavior is the same as the respective functors in $\mathcal M$.
For relations, we define

\[ \Fsq^{\mathcal M'} (c, \rho^L_c, \rho^R_{Fc}) = 
  (\Fsq^M c, \rho^R_{Fc}, \rho^L_{UF(c)}), \]
and
\[ \Usq^{\mathcal M'} (d, \rho^R_d, \rho^L_{Ud}) = 
  (\Usq^M d, \rho^L_{Ud}, \rho^R_{FU(d)}), \]

where $\rho^L_{UF(c)}$ and $\rho^R_{FU(d)}$ are as defined in the proof of Lemma
\ref{lem:representation-UF-FU}.

We define \[ (c_1, \rho^L_{c_1}, \rho^R_{Fc_1}) \times (c_2, \rho^L_{c_2}, \rho^R_{Fc_2}) = 
             (c_1 \times c_2, \rho^L_{c_1 \times c_2}, \rho^R_{F(c_1 \times c_2)}), \]

where $\rho^L_{c_1 \times c_2}$ and $\rho^R_{F(c_1 \times c_2)}$ are as defined in the
proof of Lemma \ref{lem:representation-product}.

Lastly, we define

\[ (c, \rho^L_c, \rho^R_{Fc}) \arr (d, \rho^R_d, \rho^L_{Ud}) =
    (c \arr d, \rho^R_{c \arr d}, \rho^L_{U(c \arr d)}), \]

where $\rho^R_{c \arr d}$ and $\rho^L_{U(c \arr d)}$ are as defined in the proof
of Lemma \ref{lem:representation-arrow}.

% We define $(c, \rho^L_c) \arr (d, \rho^R_d) = (c \arr d, \rho^R_{c \arr d})$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Defining an Extensional Model}\label{sec:extensional-construction-appendix}

We aim to prove the following lemma:

\begin{lemma}\label{lem:step-4-model-to-extensional-model}
  Let $\mathcal M$ be a \hyperref[def:step-4-model]{step-4 intensional model}.
  Then we can define an extensional model.
\end{lemma}
\begin{proof}

  Recall the extensional model defined in Section \ref{sec:extensional-model-definition}.

  We first establish the representability properties for this model.
  We show the left-representability squares; the right-representability squares are dual.

  \begin{itemize}
    \item We have the square

    % https://q.uiver.app/#q=WzAsNixbMCwwLCJBIl0sWzAsMSwiQSciXSxbNCwxLCJBX3IiXSxbNCwwLCJBX3IiXSxbMiwwLCJBJyJdLFsyLDEsIkEnIl0sWzAsNCwiYyIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzQsMywiY19yIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMCwxLCJlX2MiLDIseyJjdXJ2ZSI6MX1dLFszLDIsIlxccHVzaF97Y19yfShcXGRlbHJlX2MpIiwyLHsiY3VydmUiOjF9XSxbMSw1LCJyKEEnKSIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzQsNSwiXFxkZWxyZV9jIiwyXSxbNSwyLCJjX3IiLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFswLDEsImVfYyIsMCx7ImN1cnZlIjotMX1dLFszLDIsIlxcaWQiLDAseyJjdXJ2ZSI6LTF9XSxbOSwxNCwiXFxiaXNpbSIsMSx7InNob3J0ZW4iOnsic291cmNlIjoyMCwidGFyZ2V0IjoyMH0sInN0eWxlIjp7ImJvZHkiOnsibmFtZSI6Im5vbmUifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFs4LDEzLCJcXGJpc2ltIiwxLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzE2LDExLCIiLDEseyJzaG9ydGVuIjp7InNvdXJjZSI6MjAsInRhcmdldCI6MjB9LCJsZXZlbCI6Miwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}[ampersand replacement=\&]
      A \&\& {A'} \&\& {A_r} \\
      {A'} \&\& {A'} \&\& {A_r}
      \arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-3]
      \arrow["{c_r}", "\shortmid"{marking}, no head, from=1-3, to=1-5]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{e_c}"', curve={height=6pt}, from=1-1, to=2-1]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{\push_{c_r}(\delre_c)}"', curve={height=6pt}, from=1-5, to=2-5]
      \arrow["{r(A')}"', "\shortmid"{marking}, no head, from=2-1, to=2-3]
      \arrow[""{name=2, anchor=center, inner sep=0}, "{\delre_c}"', from=1-3, to=2-3]
      \arrow["{c_r}"', "\shortmid"{marking}, no head, from=2-3, to=2-5]
      \arrow[""{name=3, anchor=center, inner sep=0}, "{e_c}", curve={height=-6pt}, from=1-1, to=2-1]
      \arrow[""{name=4, anchor=center, inner sep=0}, "\id", curve={height=-6pt}, from=1-5, to=2-5]
      \arrow["\bisim"{description}, draw=none, from=1, to=4]
      \arrow[""{name=5, anchor=center, inner sep=0}, "\bisim"{description}, draw=none, from=0, to=3]
      \arrow[draw=none, from=5, to=2]
    \end{tikzcd}\]

    \item We have the square

    % https://q.uiver.app/#q=WzAsNixbMCwwLCJBX2wiXSxbMCwxLCJBX2wiXSxbMiwxLCJBIl0sWzQsMSwiQSciXSxbMiwwLCJBIl0sWzQsMCwiQSJdLFswLDQsImNfbCJdLFs0LDUsInIoQSkiXSxbMSwyLCJjX2wiLDJdLFsyLDMsImMiLDJdLFswLDEsIlxcaWQiLDIseyJjdXJ2ZSI6MX1dLFs1LDMsImVfYyIsMix7ImN1cnZlIjoxfV0sWzQsMiwiXFxkZWxsZV9jIl0sWzAsMSwiXFxwdWxsX3tjX2x9KFxcZGVsbGVfYykiLDAseyJjdXJ2ZSI6LTF9XSxbNSwzLCJlX2MiLDAseyJjdXJ2ZSI6LTF9XSxbMTAsMTMsIlxcYmlzaW0iLDEseyJzaG9ydGVuIjp7InNvdXJjZSI6MjAsInRhcmdldCI6MjB9LCJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJub25lIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMTEsMTQsIlxcYmlzaW0iLDEseyJzaG9ydGVuIjp7InNvdXJjZSI6MjAsInRhcmdldCI6MjB9LCJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJub25lIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XV0=
    \[\begin{tikzcd}[ampersand replacement=\&]
      {A_l} \&\& A \&\& A \\
      {A_l} \&\& A \&\& {A'}
      \arrow["{c_l}", from=1-1, to=1-3]
      \arrow["{r(A)}", from=1-3, to=1-5]
      \arrow["{c_l}"', from=2-1, to=2-3]
      \arrow["c"', from=2-3, to=2-5]
      \arrow[""{name=0, anchor=center, inner sep=0}, "\id"', curve={height=6pt}, from=1-1, to=2-1]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{e_c}"', curve={height=6pt}, from=1-5, to=2-5]
      \arrow["{\delle_c}", from=1-3, to=2-3]
      \arrow[""{name=2, anchor=center, inner sep=0}, "{\pull_{c_l}(\delle_c)}", curve={height=-6pt}, from=1-1, to=2-1]
      \arrow[""{name=3, anchor=center, inner sep=0}, "{e_c}", curve={height=-6pt}, from=1-5, to=2-5]
      \arrow["\bisim"{description}, draw=none, from=0, to=2]
      \arrow["\bisim"{description}, draw=none, from=1, to=3]
    \end{tikzcd}\]

\end{itemize}

\end{proof}

\section{Adequacy}\label{sec:appendix-adequacy}

In this section, we show an adequacy result for the extensional model of GTT we obtained by
applying the abstract construction introduced in Section
\ref{sec:extensional-model-construction} to the concrete model

First we establish some notation. Fix a morphism $f : 1 \to \li \mathbb{N} \cong \li \mathbb{N}$.
We write that $f \da n$ to mean that there exists $m$ such that $f = \delta^m(\eta n)$
and $f \da \mho$ to mean that there exists $m$ such that $f = \delta^m(\mho)$.

Recall that $\ltls$ denotes the relation on value morphisms defined as the bisimilarity-closure
of the intensional error-ordering on morphisms.
That is, we have $f \ltls g$ iff there exists $f'$ and $g'$ with

\[ f \bisim f' \le g' \bisim g. \]

The result we would like to show is as follows:
\begin{lemma}
If $f \ltls g : \li \mathbb{N}$, then:
\begin{itemize}
  \item If $f \da n$ then $g \da n$.
  \item If $g \da \mho$ then $f \da \mho$.
  \item If $g \da n$ then $f \da n$.
\end{itemize}
\end{lemma}

Unfortunately, this is actually not provable!
Roughly speaking, the issue is that this is a ``global'' result, and it is not possible
to prove such results inside of the guarded setting. 
In particular, if we tried to prove a result such as the above in the guarded
setting, we would run into a problem where we would have a natural number
``stuck'' under a $\later$, with no way to get out the underlying number.

Thus, to prove our adequacy result, we need to leave the guarded setting and pass back
to the more familiar, set-theoretic world with no internal notion of step-indexing.
As mentioned in the Technical Background section (Section \ref{sec:sgdt}), we can do this
using \emph{clock quantification}.

Recall that all of the constructions we have made in SGDT take place in the context of a clock $k$.
All of our uses of the later modality and guarded recursion have taken place with respect to this clock.
For example, consider the definition of the lift monad by guarded recursion in Section \ref{TODO}.
% We define the lift monad $\li^k X$ as the guarded fixpoint of $\lambda \tilde{T}. X + 1 + \later^k_t (\tilde{T}_t)$.
We can view this definition as being parameterized by a clock $k$: $\li^k : \type \to \type$.
Then for $X$ satisfying a certain technical requirement known as \emph{clock-irrelevance},
we can define the ``global lift'' monad as $\li^{gl} X := \forall k. \li^k X$.

Similarly, we can define a ``global'' version of the lock-step error ordering and the
weak bisimilarity relation on morphisms.

It can be shown that the global lift monad is isomorphic to the so-called Delay monad of Capretta \cite{TODO}.
Recall that the delay monad $\text{Delay}(X)$ is defined as the coinductive solution to the equation $\text{Delay}(X) \cong X + \text{Delay}(X)$.
It can be shown that $\li^{gl} X \cong \text{Delay}(X + 1)$ for $X$ a clock-irrelevant type.

Moreover, we claim that modulo this isomorphism, the global version of the extensional error ordering

% We have been writing the type as $\li X$, but it is perhaps more accurate to write it as $\li^k X$ to
% emphasize that the construction is parameterized by a clock $k$.