### 2024-01-26 paper-new/concrete-model.tex
work on concrete model

\section{Constructing a Concrete Model}

In this section, we build a concrete extensional model of gradual typing.
We begin by defining a step-1 intensional model, and then
apply the abstract constructions outlined in the previous section
to obtain an extensional model.

We begin with some definitions:

\begin{definition}
A \textbf{predomain} $A$ consists of a set $A$ along with two relations:
\begin{itemize}
    \item A partial order $\le_A$.
    \item A reflexive, symmetric ``bisimilarity'' relation $\bisim_A$.
\end{itemize}
\end{definition}

% TODO: later of a predomain
Given a predomain $A$, we can form the predomain $\later A$.
The underlying set is $\later |A|$ and the relation is defined in the obvious way,
i.e., $\tilde{x} \le_{\later A} \tilde{x'}$ iff $\later_t(\tilde{x}_t \le_A \tilde{x'}_t)$.
Likewise for bisimilarity.

We also give a predomain structure to the natural numbers $\mathbb{N}$, where both the
ordering and the bisimiarity relation are equality.

Morphisms of predomains are functions between the underlying sets that preserve the ordering
and the bisimilarity relation. More formally:
%
\begin{definition}
Let $A$ and $A'$ be predomains.
A morphism $f : A \to A'$ is a function between the underlying sets such that for all $x, x'$,
if $x \le_A x'$, then $f(x) \le f(x')$, and if $x \bisim_A x'$, then $f(x) \bisim_{A'} f(x')$.
\end{definition}

\begin{definition}
An \textbf{error domain} $B$ consists of a predomain $B$ along with the following data:
\begin{itemize}
    \item A distinguished ``error" element $\mho_B \in B$
    \item A morphism of predomains $\theta_B \colon \later B \to B$
\end{itemize}
\end{definition}

Morphisms of error domains are morphisms of the underlying predomains that preserve the
algebraic structure. More formally:
%
\begin{definition}
Let $B$ and $B'$ be error domains.
A morphism $\phi : B \wand B'$ is a morphism between the underlying predomains such that:
\begin{enumerate}
    \item $\phi(\mho_B) = \mho_{B'}$
    \item $\phi(\theta_B(\tilde{x})) = \theta_{B'}(\lambda t. \phi(\tilde{x}_t))$
\end{enumerate}
\end{definition}

We define a (monotone) relation on predomains $A$ and $A'$ to be a relation on the
underlying sets that is downward-closed under $\le_A$ and upward-closed under $\le_{A'}$.
More formally:

\begin{definition}
Let $A$ and $A'$ be predomains. A \emph{predomain relation} between $A$ and $A'$
is a relation $R$ between the underlying sets such that:
\begin{enumerate}
    \item (Downward closure): For all $x_1, x_2 \in A$ and $y \in A'$,
    if $x_1 \le_A x_2$ and $x_2 \mathbin{R} y$, then $x_1 \mathbin{R} y$.
    \item (Upward closure): For all $x \in A$ and $y_1, y_2 \in A'$,
    if $x \mathbin{R} y_1$ and $y_1 \le_{A'} y_2$, then $x \mathbin{R} y_2$.
\end{enumerate}
\end{definition}

Composition of relations on predomains is the usual relational composition.

Similarly, we define a (monotone) relation on error domains to be a relation on the
underlying predomains that respects error and preserves $\theta$.
%
\begin{definition}
    Let $B$ and $B'$ be error domains. An \emph{error domain relation} between
    $B$ and $B'$ is a relation $R$ between the underlying predomains such that
    \begin{enumerate}
       \item (Respects error): For all $y \in B'$, we have $\mho_B \mathbin{R} y$.
       \item (Preserves $\theta$): For all $\tilde{x}$ in $\later B$ and $\tilde{y} \in \later B'$,
       if 
       \[ \later_t( \tilde{x}_t \mathbin{R} \tilde{y}_t ), \]
       then
       \[ \theta_B(\tilde{x}) \mathbin{R} \theta_{B'}(\tilde{y}). \]
    \end{enumerate}
\end{definition}

We define composition of error domain relations $R$ on $B_1$ and $B_2$ and $S$
on $B_2$ and $B_3$ to be the least relation containing $R$ and $S$ that respects
error and preserves $\theta$.
Specifically, it is defined inductively by the following rules:

\begin{mathpar}
    \inferrule*[right = Comp]
    {b_1 \mathbin{R} b_2 \and b_2 \mathbin{S} b_3}
    {b_1 \mathbin{R \relcomp S} b_3}

    \inferrule*[right = PresErr]
    { }
    {\mho_{B_1} \mathbin{R \relcomp S} b_3}

    \inferrule*[right = PresTheta]
    {\later_t( \tilde{b_1} \mathbin{R \relcomp S} \tilde{b_3} ) }
    {\theta_{B_1}(\tilde{b_1}) \mathbin{R \relcomp S} \theta_{B_3}(\tilde{b_3}) }
\end{mathpar}

% We note that this composition has the following universal property.

We now describe the ``commuting squares".
Suppose we are given predomains $A_i, A_o, A_i'$, and $A_o'$,
relations $R_i$ and $R_o$, and morphisms $f, f'$ as shown below.

% https://q.uiver.app/#q=WzAsNCxbMCwwLCJBX2kiXSxbMSwwLCJBX2knIl0sWzAsMSwiQV9vIl0sWzEsMSwiQV9vJyJdLFswLDIsImYiLDJdLFsxLDMsImYnIl0sWzAsMSwiUl9pIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMiwzLCJSX28iLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFs0LDUsIlxcbHRkeW4iLDEseyJzaG9ydGVuIjp7InNvdXJjZSI6MjAsInRhcmdldCI6MjB9LCJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJub25lIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XV0=
\[\begin{tikzcd}[ampersand replacement=\&]
	{A_i} \& {A_i'} \\
	{A_o} \& {A_o'}
	\arrow[""{name=0, anchor=center, inner sep=0}, "f"', from=1-1, to=2-1]
	\arrow[""{name=1, anchor=center, inner sep=0}, "{f'}", from=1-2, to=2-2]
	\arrow["{R_i}", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow["{R_o}"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
	\arrow["\ltdyn"{description}, draw=none, from=0, to=1]
\end{tikzcd}\]

We say that the above square commutes, written $f \le f'$, if for all
$x \in A_i$ and $x' \in A_i'$ with $x \mathbin{R_i} x'$, we have
$f(x) \mathbin{R_o} f'(x')$.

We make the analogous definition for error domains.

\subsection{Guarded Lift Monad}\label{sec:guarded-lift-monad}

% Lift monad
The guarded error-lift monad $\li$ takes a predomain $A$ to the error domain $\li A$.
It is defined as follows:

\begin{align*}
  \li A &:= \\
  &\eta \colon A \to \li A \\
  &\mho \colon \li A \\
  &\theta \colon \later (\li A) \to \li A
\end{align*}

Formally, the lift monad $\li A$ is defined as the solution to the guarded recursive type equation

\[ \li A \cong A + 1 + \later \li A. \]

This captures the intuition that a program may either return a value,
fail at run-time, or take one or more observable steps of computation.
Previous work has studied such a similar construct, called the guarded lift
monad \cite{mogelberg-paviotti2016}; our version here our version augments it with the notion of error.

Since we claimed that $\li A$ is a monad, we need to define the monadic operations
and show that they respect the monadic laws. The return is just $\eta$, and the monadic extend
is defined via guarded recursion by cases on the input.
Verifying that the monadic laws hold uses \lob-induction and is straightforward.

% \eric{Check}
% We can also show that $\li A$ is the free error- and -later algebra on $A$, in that
% for any morphism of predomains $f : A \to UB$, there is a unique morphism of error domains
% $f^* : \li A \to \li A$ extending $f$.

There is a functor $U$ from error domains to predomains that on objects simply returns the
underlying predomain, and on morphisms returns the underlying morphism of predomains.
%
% TODO check this

It is easily verified that $\li A$ is the free error- and later-algebra on the predomain $A$,
so we have that $\li$ is left-adjoint to $U$.

% i.e., error domain morphisms from $\li A$ to $B$ are in one-to-one correspondence with
% predomain morphisms from $A$ to $UB$.

% We define $\delta : U(\li A) \To U(\li A)$ by $\delta(x) = \theta(\nxt x)$.
% We define $\hat{\delta} : \li A \arr \li A$ by $\hat{\delta} = \ext{(\delta \circ \eta)}{}$.
% Note that by definition of $\text{ext}$, we have that $\hat{\delta}$ is a morphism of error domains.

%\subsubsection{Lock-Step Error Ordering}\label{sec:lock-step}

The partial order $\le_{\li A}$ is the lock-step error ordering defined by guarded recursion as follows:

\begin{itemize}
    \item 	$\eta\, x \ltls \eta\, y$ if $x \le_A y$.
    \item 	$\mho \ltls l$ for all $l$ 
    \item   $\theta\, \tilde{r} \ltls \theta\, \tilde{r'}$ if
            $\later_t (\tilde{r}_t \ltls \tilde{r'}_t)$
  \end{itemize}

The idea is that two computations $l$ and $l'$ are related if they are in
lock-step with regard to their intensional behavior, up to $l$ erroring.

Given a relation $R : A \rel A'$, we define in an analogous manner a heterogeneous
version of the lock-step error ordering between $\li R : \li A \rel \li A'$.
% We define the action of $\li$ on a relation $R$ between $A$ and $A'$ to be the
% ``heterogeneous" version of the lock-step error ordering.

% TODO action of \li on commuting squares

%\subsubsection{Weak Bisimilarity}\label{sec:weak-bisimilarity}

For a predomain $A$, we define a relation on $\li A$, called ``weak bisimilarity",
written $l \bisim l'$. Intuitively, we say $l \bisim l'$ if they are equivalent ``up to delay''.
The weak bisimilarity relation is defined by guarded recursion as follows:

\begin{align*}
  &\mho \bisim \mho \\
%
  &\eta\, x \bisim \eta\, y \text{ if } 
    x \bisim_A y \\
%		
  &\theta\, \tilde{x} \bisim \theta\, \tilde{y} \text{ if } 
    \later_t (\tilde{x}_t \bisim \tilde{y}_t) \\
%	
  &\theta\, \tilde{x} \bisim \mho \text{ if } 
    \theta\, \tilde{x} = \delta^n(\mho) \text { for some $n$ } \\
%	
  &\theta\, \tilde{x} \bisim \eta\, y \text{ if }
    (\theta\, \tilde{x} = \delta^n(\eta\, x))
  \text { for some $n$ and $x : \ty{A}$ such that $x \bisim_A y$ } \\
%
  &\mho \bisim \theta\, \tilde{y} \text { if } 
    \theta\, \tilde{y} = \delta^n(\mho) \text { for some $n$ } \\
%	
  &\eta\, x \bisim \theta\, \tilde{y} \text { if }
    (\theta\, \tilde{y} = \delta^n (\eta\, y))
  \text { for some $n$ and $y : \ty{A}$ such that $x \bisim_A y$ }
\end{align*}

When both sides are $\eta$, then we ensure that the underlying values are related
by the bisimilarity relation on $A$.
When one side is a $\theta$ and the other is $\eta x$ (i.e., one side steps),
we stipulate that the $\theta$-term runs to $\eta y$ where $x$ is bisimilar to $y$.
Similarly when one side is $\theta$ and the other $\mho$.
If both sides step, then we allow one time step to pass and compare the resulting terms.
In this way, the definition captures the intuition of terms being equivalent up to
delays.

It can be shown (by \lob-induction) that the step-sensitive relation is symmetric.
However, it can also be shown that this relation is \emph{not} transitive:
The argument is the same as that used to show that the step-insensitive error
ordering $\semltbad$ described above is not transitive. Namely, we show that
if it were transitive, then it would have to be trivial in that $l \bisim l'$ for all $l, l'$.
that if this relation were transitive, then it would relate all values of type $\li A$.

% internal hom for predomains and error domains
Given predomains $A$ and $A'$, we can form the predomain of
predomain morphisms from $A$ to $A'$, denoted $A \To A'$.
\begin{itemize}
    % Should we give the definition involving x and x'?
    \item The ordering is defined by $f \le_{A \To A'} f'$ iff for all
    $x \in A$, we have $f(x) \le_{A'} f'(x)$.
    \item The bisimilarity relation is defined by $f \bisim_{A \To A'} f'$ iff
    for all $x, x' \in A$ with $x \bisim_{A} x'$, we have $f(x) \bisim_{A'} f'(x')$. 
\end{itemize}

Given $f : A_1' \to A_1$ and $g : A_2 \to A_2'$ we define the predomain morphism
$f \To g : (A_1 \To A_2) \to (A_1' \To A_2')$ by $\lambda h. \lambda x'. g(h(f(x')))$.

% TODO: include this?
% The monadic extension operation $\ext{\cdot}{} : (A_1 \To U (\li A_2)) \To (\li A_1 \To U(\li A_2))$
% is a morphism of predomains from $A_1 \To U(\li A_2)$ to $U(\li A_1) \To U(\li A_2)$, i.e.,
% it preserves the ordering and bisimilarity relations.

% Given a predomain $A$ and error domain $B$, we define $A \arr B := A \To UB$.
We note that $A \To UB$ carries a natural error domain structure
(in the below, the lambda is a meta-theoretic notation):
\begin{itemize}
    \item The error is given by $\lambda x . \mho_B$
    \item The $\theta$ operation is defined by
      \[ \theta_{A \To UB}(\tilde{f}) = \lambda x . \theta_B(\lambda t . \tilde{f}_t(x)). \]
\end{itemize}

Given a predomain $A$ and error domain $B$, we define
$A \arr B$ to be the error domain such that $U(A \arr B) = A \To UB$,
and whose error and $\theta$ operations are as defined above.
We can define the functorial action of $\arr$ on morphisms
$f \arr \phi$ in the obvious way.

It is easily verified that $A \arr B$ is an exponential of $UB$ by $A$
in the category of predomains and their morphisms.

Lastly, given a relation of predomains $R$ between $A$ and $A'$, and a relation
of error domains $S$ between $B$ and $B'$, we define the relation $R \arr S$
between $A \arr B$ and $A' \arr B'$ in the obvious way, i.e., $f \in A \arr B$
is related to $g \in A' \arr B'$ iff for all $x \in A$ and $x' \in A'$ with
$x \mathrel{R} x'$, we have $f(x) \mathrel{S} g(x')$.
%
One can verify that this relation is indeed a relation of error domains
in that it respects error and preserves $\theta$.

With all of the above data, we can form a step-1 intensional model of gradual typing
(See Definition \ref{def:step-1-model}).

\subsection{The Dynamic Type}

The predomain representing the dynamic type will be defined using guarded recursion
as the solution to the equation

\[ D \cong \mathbb{N}\, + (D \times D)\, + \laterhs U(D \arr FD). \]

% Note that the operators in the above equation are all combinators for predomains, so
% this also defines the ordering and the bisimilarity relation for $D$.

For the sake of clarity, we name the ``constructors" $\text{nat}$, $\text{times}$,
and $\text{fun}$, respectively.

We define $e_\mathbb{N} : \mathbb{N} \to D$ to be the injection into the first
component of the sum, and $e_\times : D \times D \to D$ to be the injection into
the second component of the sum, and $e_\to : U(D \arr F D)$ to be the morphism
$\nxt$ followed by the injection into the third component of the sum.

Explicitly, the ordering on $D$ is given by:

\begin{align*}
    \tnat(n) \le \tnat(n') 
        &\iff n = n' \\
    \ttimes (d_1, d_2) \le \ttimes (d_1', d_2')
        &\iff d_1 \le d_2 \text{ and } d_1' \le d_2'\\
    \tfun(\tilde{f}) \le \tfun(\tilde{f'}) 
        &\iff \later_t(\tilde{f}_t \le \tilde{f'}_t)
\end{align*}

We define a relation $\inat : \mathbb{N} \rel D$ by
$(n, d) \in \inat$ iff $e_\mathbb{N} \le_D d$.
We similarly define $\itimes : D \times D \rel D$ by
$((d_1, d_2), d) \in \itimes$ iff $e_\times(d_1, d_2) \le_D d$,
and we define $\text{inj}_\to : U(D \arr F D) \rel D$ by
$(f, d) \in \iarr$ iff $e_\to(f) \le_D d$.

Now we define the perturbations for $D$.
Recall from our construction of a model with perturbations
(Section \ref{sec:constructing-perturbations}) that for each value
type $A$ we associate a commutative monoid $P_A$ of perturbations
and a homomorphism into the monoid of endomorphisms bisimilar to the identity.

We define the perturbations for $D$ via least-fixpoint in the category of
commutative monoids as

\[ P_D \cong (P_D \times P_D) \times (P_D \times (\mathbb{N} \times P_D)), \]

where $\oplus$ is the product (also the coproduct) of commutative monoids.

We now explain how to interpret these perturbations as endomorphisms.
We define $\ptb_D : P_D \to \{ f : D \to D \mid f \bisim \id \}$ as in the below.
Note the use of the functorial action of $\arr$ on morphisms.
% via the universal property of the coproduct of monoids, giving a case for each of the generators.
% In the below, note the use of the functorial action of $\arr$ on morphisms.

%
% \[ P_D \cong (P_D \times P_D) \times (P_D \times (\mathbb{N} \times P_D)), \]
%

$\ptb_D(p_1, p_2, p_3, n, p_4) = \lambda d.\text{case $d$ of}$
\begin{align*}
    &\alt \tnat(m) \mapsto \tnat(m) \\
    &\alt \ttimes(d_1, d_2) \mapsto {\ttimes(\ptb_D(p_1)(d_1), \ptb_D(p_2)(d_2))} \\
    &\alt \tfun(\tilde{f}) \mapsto {\tfun(\lambda t. (\ptb_D(p_3) \arr (\delta^n \circ F(\ptb_D(p_4))))(\tilde{f}_t))} \\
\end{align*}

One can verify that this forms a homomorphism from $P_D \to \{ f : D \to D : f \bisim \id \}$.
For this it is easiest to use the fact that the product coindices with the coproduct
in the category of commutative monoids, and thus it suffices to check the condition for each summand separately.

We claim that the three relations $\inat$, $\itimes$, and $\iarr$
%and their lifted versions 
satisfy the push-pull property.
As an illustrative case, we establish the pull property for the relation $\iarr$.

We define $\pull_{\iarr} : P_D \to P_{U(D \arr FD)}$ by cases as

\[ \pull_{\iarr}(p_1, p_2, p_3, n, p_4) = (0, p_3, n, p_4) \]

(recall that $P_{UB} = \mathbb{N} \times P_B$, and $P_{A \arr B} = P_A \times P_B$,
so $P_{U(D \arr FD)}) = \mathbb{N} \times P_D \times \mathbb{N} \times P_D$.

Now let $(f, d) \in \iarr$. This means that $d$ must be of the form $\tfun{\tilde{f}}$ where
Then we have $\ptb_{U(D \arr FD)}(0, p_3, n, p_4) = $

\eric{Details}

\eric{Quasi-representability}

    % \item $\ptb_D(1)$
    % \item $\ptb_D(\delta^K_D)$ is defined similarly to the previous but has
    % \[ \id \arr i^K(\delta^K_D) \] instead.
    % \item $\ptb_D(\delta^K_D) = \lambda d.\text{case $d$ of}$
    %   \begin{align*} 
    %     &\alt \ttimes(d_1, d_2) \mapsto {\ttimes(i^K(\delta^K_D)(d_1), d_2)} \\
    %     &\alt d' \to d'
    %   \end{align*}
    % \item $\ptb_D(\delta^K_D)$ is defined similarly to the previous but has 
    % \[ (d_1, \ptb_D(\delta^K_D)(d_2)) \] instead.

\begin{comment}
Before defining the perturbations, recall from our concrete construction
(Section \ref{TODO}) that to each value type $A$ we associate a monoid $P_A$
of pure perturbations and a monoid $P^K_{A}$ of Kleisli perturbations.

We define simultaneously the pure and Kleisli perturbation monoids for $D$,
denoted $P_D$ and $P^K_D$ respectively.
We define $P_D$ to be the monoid such that

\[ P_D \cong ((P^K_D)^{op} \times P_D) \oplus (P_D \times P_D), \]

and we define $P^K_D$ to be the monoid such that

\[ P^K_D \cong \mathbb{N} \oplus ((P_D)^{op} \oplus P^K_D) \oplus (P^K_D \oplus P^K_D), \]

where we use $\mathbb{N}$ because it is the free monoid on one generator.

Now we explain how to interpret these monoids as submonoids of endomorphisms.
We define $i : P_D \to \{ f : D \to D \mid f \bisim \id \}$
and $i^K : P^K_D \to \{ \phi : FD \to FD \mid \phi \bisim \id \}$ mutually
via the universal property of the coproduct of monoids, giving a case for each of the generators.
In the below, note the use of the functorial action of $\arr$ on morphisms.

\begin{itemize}
    \item $i(\delta^K_D, \delta_D) = \lambda d.\text{case $d$ of}$
    \begin{align*}
        &\alt \tfun(\tilde{f}) \mapsto 
          \tfun(\lambda t. (i^K(\delta^K_D) \to i(\delta_D))(\tilde{f}_t)) \\
        &\alt d' \to d'
    \end{align*}

    \item $i(\delta_D^1, \delta_D^2) = \lambda d.\text{case $d$ of}$ 
    \begin{align*}
        &\alt \ttimes(d_1, d_2) \mapsto
           \ttimes(i(\delta_D^1)(d_1), i(\delta_D^2)(d_2)) \\
        &\alt d' \to d'
    \end{align*}
\end{itemize}

We define $i^K(\delta^D)$ by
%
%  P^K_D \cong \mathbb{N} \oplus ((P_D)^{op} \oplus P^K_D) \oplus (P^K_D \oplus P^K_D)
%
\begin{itemize}
    \item $i^K(1) = \hat{\delta}$
    \item $i^K(\delta_D) = F(\lambda d.\text{case $d$ of}$
      \begin{align*} 
        &\alt \tfun(\tilde{f}) \mapsto {\tfun(\lambda t. (i(\delta_D) \arr \id)(\tilde{f}_t))} \\
        &\alt d' \to d')
      \end{align*}
    \item $i^K(\delta^K_D)$ is defined similarly to the previous but has
    \[ \id \arr i^K(\delta^K_D) \] instead.
    \item $i^K(\delta^K_D) = F(\lambda d.\text{case $d$ of}$
      \begin{align*} 
        &\alt \ttimes(d_1, d_2) \mapsto {\ttimes(i^K(\delta^K_D)(d_1), d_2)} \\
        &\alt d' \to d')
      \end{align*}
    \item $i^K(\delta^K_D)$ is defined similarly to the previous but has 
    \[ (d_1, i^K(\delta^K_D)(d_2)) \] instead.
\end{itemize}

% perturbations and quasi-representability
We claim that the three relations $\inat$, $\itimes$, and $\iarr$ and their lifted
versions satisfy the push-pull property.
As an illustrative case, we establish the pull property for the $\li \iarr$.
We define $\pull : P^K_D \to P^K_{U(D \arr FD)}$ by cases as

\begin{align*}
 \pull(1) = \hat{\delta}
\end{align*}

We need to check that the relevant squares commute for all $\delta \in P^K_D$.
It suffices by the universal property of the coproduct of monoids to ensure that
this holds for the generators.

We claim that the relations $\inat$, $\itimes$, and $\iarr$ are quasi-left-representable,
and that their lifts are quasi-right-representable.
Indeed, since the relations are functional, it is easy to see that they are left-representable
where the perturbations are simply the identity.

The most interesting case is the right-quasi-representability of $\li(\iarr)$.
We define the projection as follows:

\begin{align*}
\end{align*}

And for the perturbations, we take

% The perturbation monoid for $D$ is defined inductively as the free monoid $M$ equipped
% with an operation $\To : M \times M \to M$, i.e., we take the set defined inductively by

% \begin{mathpar}
%     \inferrule*[]
%     {}
%     {e \in M}

%     \inferrule*[]
%     {p \in M \and p' \in M}
%     {p \odot p' \in M}

%     \inferrule*[]
%     {p \in M \and p' \in M}
%     {p \To p' \in M}
% \end{mathpar}

% and quotient by the necessary monoid equations.

\end{comment}

% \subsection{Obtaining an Extensional Model}

Now that we have defined an intensional model with an interpetation for the dynamic type, we can apply
the abstract construction introduced in Section \ref{sec:extensional-model-construction}.
Doing so, we obtain an extensional model of gradual typing, where the squares are given by the
``bisimilarity closure'' of the intensional error ordering.