### 2024-07-07 paper-new/concrete-term-model.tex
[paper] move technical background on SGDT; edit section on concrete term model

\section{A Concrete Term Model}\label{sec:concrete-term-model}

\eric{Wording?}
%
In the present section, we use the techniques of guarded type theory to give a
set-theoretic denotational semantics to the terms of the gradually-typed lambda
calculus introduced in Section \ref{sec:GTLC}. This serves two important
purposes: First, it acts as a warm-up for employing guarded type theory to
define a denotational semantics. This acts as a stepping stone into the next
section, where we extend the techniques employed here to accommodate the type
and term precision orderings of the gradually-typed lambda calculus. Second,
constructing this semantic model establishes the validity of the $\beta$ and
$\eta$ principles for the gradually-typed lambda calculus. 
%
% In Section \ref{sec:towards-relational-model}, we will discuss how to extend the
% denotational semantics to accommodate the type and term precision orderings. 
%
%necessitate the abstract components that are found in the definition of the model.

\subsection{Guarded Type Theory}

We begin with a brief overview of synthetic guarded domain theory, which we will
use to construct the model. \emph{Synthetic guarded domain theory}, or SGDT for
short, is an axiomatic framework in which we can reason about non-well-founded
recursive constructions while abstracting away the specific details of
step-indexing that we would need to track if we were working analytically. This
allows us to avoid the tedious reasoning associated with traditional
step-indexing techniques. We provide a brief overview here; more details can be
found in \cite{birkedal-mogelberg-schwinghammer-stovring2011}.

SGDT offers a synthetic approach to domain theory that allows for guarded
recursion to be expressed syntactically via a type constructor $\later \colon
\type \to \type$ (pronounced ``later''). The use of a modality to express
guarded recursion was introduced by Nakano \cite{Nakano2000}.
%
Given a type $A$, the type $\later A$ represents an element of type $A$ that is
available one time step later. There is an operator $\nxt : A \to\, \later A$
that ``delays'' an element available now to make it available later. We will use
a tilde to denote a term of type $\later A$, e.g., $\tilde{M}$.

% TODO later is an applicative functor, but not a monad

There is a \emph{guarded fixpoint} operator
%
\[
  \fix : \forall T, (\later T \to T) \to T.
\]
%
That is, to construct a term of type $T$, it suffices to assume that we have
access to such a term ``later'' and use that to help us build a term ``now''.
This operator satisfies the axiom that $\fix f = f (\nxt (\fix f))$. In
particular, this axiom applies to propositions $P : \texttt{Prop}$; proving a
statement in this manner is known as $\lob$-induction.

% Clocked Cubical Type Theory
\subsubsection{Ticked Cubical Type Theory}
% TODO motivation for Clocked Cubical Type Theory, e.g., delayed substitutions?

Ticked Cubical Type Theory \cite{mogelberg-veltri2019} is an extension of
Cubical Type Theory \cite{CohenCoquandHuberMortberg2017} that has an additional
sort called \emph{ticks}. Ticks were originally introduced in
\cite{bahr-grathwohl-bugge-mogelberg2017}. A tick $t : \tick$ serves as evidence
that one unit of time has passed. In Ticked Cubical Type Theory, the type
$\later A$ is the type of dependent functions from ticks to $A$. The type $A$ is
allowed to depend on $t$, in which case we write $\later_t A$ to emphasize the
dependence.

% TODO next as a function that ignores its input tick argument?

% TODO include a figure with some of the rules for ticks

The rules for tick abstraction and application are similar to those of ordinary
$\Pi$ types. A context now consists of ordinary variables $x : A$ as well as
tick variables $t : \tick$. The presence of the tick variable $t$ in context
$\Gamma, (t : \tick), \Gamma'$ intuitively means that the values of the
variables in $\Gamma$ arrive ``first'', then one time step occurs, and then the
values of the variables in $\Gamma'$ arrive.
%
The abstraction rule for ticks states that if in context $\Gamma, t : \tick$
the term $M$ has type $A$, then in context $\Gamma$ the term $\lambda t.M$ has
type $\later_t A$.
%
Conversely, if we have a term $M$ of type $\later A$, and we have available in
the context a tick $t' : \tick$, then we can apply the tick to $M$ to get a
term $M[t'] : A[t'/t]$. However, there is an important restriction on when we
are allowed to apply ticks. In order to apply $M$ to tick $t$, $M$ must be
well-typed in the prefix of the context occurring before the tick $t$. That is,
all variables mentioned in $M$ must be available before $t$. This ensures that
we cannot, for example, define a term of type $\later \laterhs A \to\, \laterhs
A$ via repeated tick application.
% TODO restriction on tick application
%
For the sake of brevity, we will also write tick application as $M_t$.

\subsection{A Call-by-push-value Model}

% TODO there aren't actually computation types in the syntax. They
% arise because we are embedding call-by-value into CBPV.

For the sake of modularity and ease of description, we will formulate the model
for the gradually-typed lambda calculus using the \emph{call-by-push-value}
paradigm \cite{levy99}. In a call-by-push-value model, there is a
category $\calV$ of \emph{value types} and a category $\calE$ of
\emph{computation types}. Value types represent static values, while computation
types represent effectful terms. We denote value types by $A, A', A_1, A_i,
A_o$, etc, and computation types by $B, B', B_1, B_i, B_o$, etc. There is a pair
of adjoint functors $F : \calV \to \calE$ and $U : \calE \to \calV$ with $F
\dashv U$, and a functor $\arr : \calV^{op} \times \calE \to \calE$ such that
$U(A \arr B) \cong A \To UB$ where $\To$ denotes the internal hom in $\calV$.

We embed our call-by-value gradually-typed cast calculus into a
call-by-push-value model as follows. The value types will be sets, and
computation types will be sets with additional algebraic structure corresponding
to the possible behaviors of a gradually-typed program. One such behavior is
\emph{failure}: a program may fail at run-time because of a type error. In
addition to this, a program may fail to terminate. This is possible because the
dynamic type includes function types and so is recursive. We will discuss the
dynamic type further in a subsequent section, but for now it suffices to know
that we can express a diverging computation in our syntax. Therefore, it will be
necessary to track the ``steps'' taken by a term; we can regard this as a way of
intensionally modelling \emph{partiality}.

More concretely, the computation types are defined to be sets $B$ that are
equipped with a distinguished error element $\mho_B$ and a map $\theta_B \colon
\laterhs B \to B$. We call such sets \emph{error domains}. A \emph{morphism of
error domains} from $B_1$ to $B_2$ is a function $\phi$ on the underlying sets
that respects the error and $\theta$ maps, i.e., for which 
%
(1) $\phi(\mho_{B_1}) = \mho_{B_2}$ and
%
(2) $\phi(\theta_{B_1}(\tilde{x})) = \theta_{B_2}(\later (\phi(\tilde{x})))$
(using the functorial action of $\later$).

The functor $U : \errdom \to \Set$ simply takes the underlying set of the error
domain. The functor $F : \Set \to \errdom$ is defined to be the left adjoint to
$U$ and constructs the \emph{free error domain} on a set $A$.
%
The functor $\arr : \Set^{op} \times \errdom \to \errdom$ is defined on objects
as follows. Given a set $A$ and error domain $B$, $A \arr B$ is the error domain
whose underlying set is the set of functions $A \to UB$. The error element is
the constant error function, i.e., $\lambda x. \mho_B$. The map $\theta :
(\laterhs (A \to UB)) \to (A \to UB)$ is defined by 
$\theta(\tilde{f}) = \lambda x . \theta_B (\tilde{f}_t)$.
%
Given a morphism $f : A_o \to A_i$ and $\phi : B_i \to B_o$, the morphism $f
\arr \phi : (A_i \arr B_i) \to (A_o \arr B_o)$ is given by pre-composing by $f$
and post-composing by $\phi$.

\subsection{The Lift + Error Monad}\label{sec:lift-monad}

We describe a monad that decomposes into the free-forgetful adjunction
between sets and error domains. We base the construction on the \emph{guarded
lift monad} described in \cite{mogelberg-paviotti2016}. Here, we augment the
guarded lift monad to accommodate the additional effect of failure. For a type
$A$, we define the \emph{guarded lift monad with failure} $\li A$, which we will
just call the \emph{guarded lift monad} or simply the \emph{lift monad}.
%
\begin{align*}
  \li A :=\quad
  &\alt \eta \colon A \to \li A \\
  &\alt \mho \colon \li A \\
  &\alt \theta \colon \laterhs (\li A) \to \li A
\end{align*}
%
Unless otherwise mentioned, all constructs involving $\later$ or $\fix$ are
understood to be with respect to a fixed clock $k$. So for the above, we really
have for each clock $k$ a type $\li^k A$ with respect to that clock.
%
Formally, the lift monad $\li A$ is defined as the solution to the guarded
recursive type equation
%
\[ \li A \cong A + 1\, + \laterhs \li A. \]
%
An element of $\li A$ should be viewed as a computation that can either (1)
return a value (via $\eta$), (2) raise an error and stop (via $\mho$), or (3)
take a step (via $\theta$).
%
Notice there is a computation $\fix\, \theta$ of type $\li A$. This represents a
computation that runs forever and never returns a value.
%
Since we claimed that $\li A$ is a monad, we need to define the monadic
operations and show that they respect the monadic laws. The return is just
$\eta$, and given $f : A \to \li A'$, the monadic extension $\text{ext}(f)
\colon (\li A \to \li A')$ is defined via guarded recursion by cases on the
input value of type $\li A$.
% It is instructive to give at least one example of a use of guarded recursion, so
% we show below how to define extend:
% TODO
%
%
Verifying that the monadic laws hold uses \lob-induction and is straightforward.
%
% TODO mention that error domains are algebras of the lift monad?
We observe that the guarded lift monad applied to a set $A$ is the underlying
set of the free error domain on $A$ $A$, i.e., we have $\li = UF$ where $F :
\Set \to \errdom$ and $U : \errdom \to \Set$ and $F \dashv U$. Given a set $A$
and an error domain $B$, the set of functions $A \to UB$ is naturally isomorphic
to the set of error domain morphisms $FA \to B$.

% The function type A \ra A' will be modeled as \sem{A} \to \li \sem{A'}

\subsection{Modeling the Dynamic Type}\label{sec:dynamic-type}

We now discuss how to model the dynamic type $\dyn$.
The goal is to define a set $D$ that satisfies the isomorphism
%
\[ D \cong \Nat + (D \times D) + (D \to \li D). \]
%
where the use of $\li D$ in the codomain corresponds to the fact that when a
value of type $\dyn$ is a function, the function when called may return an error
or take a step of computation. This equation does not have inductive or
coinductive solutions. The usual way of dealing with such equations is via
domain theory, by which we can obtain an exact solution. However, by using
guarded recursion we can solve the equation and obtain a guarded solution in a
process analogous to giving a normal set-theoretic denotational semantics.
Consider the following similar-looking equation:
%
\[ D \cong \Nat + (D \times D)\, + \laterhs (D \to \li D). \]
%
Since the negative occurrence of $D$ is guarded under a later, this equation has
a solution constructed via a combination of least fixed-point and guarded
recursion. Specifically, consider the parameterized inductive type
%
\[ D'\, X := \mu T. \Nat + (T \times T) + X. \]
%
Now consider the function $f$ defined by
%
\[ \lambda (\tilde{D} \colon \laterhs \type) . D' (\laterhs_t (\tilde{D}_t \to (\li (\tilde{D}_t)))). \]
%
Recall that the tick $t : \tick$ is evidence that time has passed, and since
$\tilde{D}$ has type $\later \type$, i.e. $\tick \to \type$, then $\tilde{D}_t$
has type $\type$.
%
Finally, define 
\[ D := \fix f. \]
%
Observe that the unfolding equality characterizing the guarded fixpoint implies
that $D = D' (\later D \to \li D)$, and by the definition of $D'$ as a least
fixpoint we see that this is equal to $\Nat + (D \times D) + \later (D \to \li
D)$.

\subsection{Term Semantics}\label{sec:term-interpretation}

We can now give a semantics to the types and terms of the gradually-typed lambda
calculus. 
%
Much of the semantics is similar to a normal call-by-value denotational
semantics, so we focus only on the cast semantics. We interpret types as sets;
the interpretation of types is given in Figure \ref{fig:type-interpretation}.
Contexts $\Gamma = x_1 \colon A_1, \dots, x_n \colon A_n$ are interpreted as the
product $\sem{A_1} \times \cdots \times \sem{A_n}$. The semantics of the dynamic
type $\dyn$ is the set $D$ introduced in Section \ref{sec:dynamic-type}. The
product type $A \times A'$ is interpreted as the Cartesian product of the
denotations of $A$ and $A'$. The function type $A \ra A'$ is interpreted as the
set of functions from $\sem{A}$ to $\li (\sem {A'})$.
%
The interpretation of a value $\hasty {\Gamma} V A$ is a function from
$\sem{\Gamma}$ to $\sem{A}$. Likewise, a term $\hasty {\Gamma} M {{A}}$ is
interpreted as a function from $\sem{\Gamma}$ to $\li \sem{A}$.

We now discuss the semantics of the cast terms, shown in Figure
\ref{fig:term-semantics}. Upcasts are pure, so the upcast ${\upc c}$, where $c :
A \ltdyn A'$, denotes an ordinary function from $\sem{A}$ to $\sem{A'}$.
Downcasts are effectful, so a downcast ${\dnc c}$ is interpreted as a morphism
of error domains between $F\sem{A'}$ and $F\sem{A}$. By definition of the
adjunction this is equivalent to an ordinary function between $\sem{A'}$ and
$UF\sem{A}$, i.e., $\sem{A'} \to \li \sem{A}$. We use $\text{ext}$ to go from
the former to the latter in the definitions of the downcasts.

% The upcast for $c_i \ra c_o$ and the downcast for $c_1 \times c_2$ make use of
% \emph{Kleisli functors} $\tok$ and $\timesk$.

\eric{Introduce Kleisli categories and functors}

Recall the definition of type precision derivations given in Figure
\ref{fig:typrec}. The semantics of the up- and downcasts for a type precision
derivation $c$ are defined by induction on $c$.
%
For $r(A)$ the up- and downcasts are simply the identity function. 
%
For the composition $c \comp c'$ we compose the cast for $c$ and the cast for
$c'$ as functions.
%
For $c_i \ra c_o$ we apply the casts for $c_i$ and $c_o$ in the domain and
codomain respectively. More concretely, for the upcast we are given a function
$V_f : A_i \to \li A_o$ and we must define a function $A_i' \to \li A_o'$. The
function first downcasts its argument according to $c_i$, resulting in an
element of $\li A_i$. It then applies $\ext{V_f}{}$ to this value to obtain an
element of $\li A_o$. Finally, it applies the upcast of $c_o$ via the functorial
action of the lift monad, obtaining an element of $\li A_o'$ as needed. For the
downcast from $A_i' \to \li A_o'$ to $A_i \to \li A_o$, the resulting function
first applies the upcast by $c_i$ to its argument, then applies the original
function, and finally applies the downcast by $c_o$ to the result.
%
Likewise, in the upcast for $c_1 \times c_2$, we apply the cast for $c_1$ on the
left and the cast for $c_2$ on the right.

The ``base cases'' for the casts are $\inat$, $\itimes$, and $\iarr$. Recall
that $D$ is isomorphic to $\Nat\, + (D \times D)\, + \laterhs (D \to \li D)$
(say that the sum is left-associative).
%
For $\inat$, the upcast is simply $\inl \circ \inl$. The downcast has type $D
\to \li D$ and performs a case inspection on the sum type. If it is a natural
number $n$, then we return $\eta\, n$. Otherwise, we return $\mho$, modelling
the fact that a run-time error occurs.
%
For $\itimes$, the upcast is $\inl \circ \inr$, and the downcast performs a case
inspection analogously to the natural number downcast.
%
% TODO explain this further
For $\iarr$, the upcast is given by $\inr \circ \nxt$. The downcast performs a
case inspection, and in the $\inr$ case, it uses a $\theta$ in order to gain
access to the function under the $\later$.

%
% TODO write this out explicitly?

% TODO should we include function casts?

\begin{figure*}
  \begin{align*}
    \sem{\nat} &= \Nat \\
    \sem{\dyn} &= D \\
    \sem{A \times A'} &= \sem{A} \times \sem{A'} \\ 
    \sem{A \ra A'} &= \sem{A} \To \li \sem{A'} \\
  \end{align*}
  \caption{Interpretation of types}
  \label{fig:type-interpretation}
\end{figure*}

\begin{figure*}
  \eric{Check this}

  % TODO check these
  \begin{align*}
    % \sem{\zro}         &= \lambda \gamma . 0 \\
    % \sem{\suc\, V}     &= \lambda \gamma . (\sem{V}\, \gamma) + 1 \\
    % \sem{x \in \Gamma} &= \lambda \gamma . \gamma(x) \\
    % \sem{\lda{x}{M}}   &= \lambda \gamma . \lambda a . \sem{M}\, (*,\, (\gamma , a))  \\
    % \sem{V_f\, V_x}    &= \lambda \gamma . {({(\sem{V_f}\, \gamma)} \, {(\sem{V_x}\, \gamma)})} \\
    % \sem{\err_B}       &= \lambda \gamma . \mho \\
    %
    % upcasts
    \sem{\upc{r(A)}} &= \id \\
    \sem{\upc{(c \comp c')}} &= \sem{\upc{c'}} \circ \sem{\upc{c}} \\
    \sem{\upc{(c_i \ra c_o)}} &=
      \lambda V_f . \text{map } \upc{c_o} \circ \ext{V_f}{} \circ \dnc{c_i} \\
    \sem{\upc{(c_1 \times c_2)}} &= 
      \lambda (V_1, V_2) . ((\sem{\upc{c_1} V_1}) , 
                            (\sem{\upc{c_2} V_2})) \\
    \sem{\upc{\inat}} &= 
      \lambda V_n . \inl\, \inl\, V_n \\
    \sem{\upc{\itimes}} &= 
      \lambda (V_1, V_2) . \inl\, \inr\, (V_1 , V_2) \\
    \sem{\upc{\iarr}} &= 
      \lambda V_f . \inr\, \nxt V_f \\[2ex]
    %
    % downcasts
    \sem{\dnc{r(A)}} &= \id \\
    \sem{\dnc{(c \comp c')}} &= \sem{\dnc{c}} \circ \sem{\dnc{c'}} \\
    \sem{\dnc{(c_i \ra c_o)}} &=
     \text{ext}(\eta(\lambda V_f . \dnc{c_o} \circ V_f \circ \upc{c_i})) \\
    \sem{\dnc{(c_1 \times c_2)}} &= (\dnc{c_1} \timesk A_2) \circ (A_1' \timesk \dnc{c_2}) \\
    \sem{\dnc{\inat}} &= \text{ext}(
      \lambda V_d . \text{case $({V_d})$ of}
      \{ \inl\, (\inl\, n) \to \eta\, n
         \alt \text{otherwise} \to \mho \}) \\
    \sem{\dnc{\itimes}} &= \text{ext}(
      \lambda V_d . \text{case $({V_d})$ of}
      \{ \inl\, (\inr\, (d_1, d_2)) \to \eta\, (d_1, d_2)
         \alt \text{otherwise} \to \mho \}) \\
    \sem{\dnc{\iarr}} &= \text{ext}(
      \lambda V_d . \text{case $({V_d})$ of}
      \{ \inr\, \tilde{f} \to \theta\, (\lambda t. \eta (\tilde{f}_t))
         \alt \text{otherwise} \to \mho \}) \\
    % \sem{\ret\, V}       &= \lambda \gamma . \eta\, \sem{V} \\
    % \sem{\bind{x}{M}{N}} &= \lambda \delta . \ext {(\lambda x . \sem{N}\, (\delta, x))} {\sem{M}\, \delta} \\
  \end{align*}

  \caption{Semantics of casts.}
  \label{fig:term-semantics}
\end{figure*}