### 2024-07-09 paper-new/towards-relational-model.tex
rename some files, progress on the relational model motivation section

\section{Towards a Full Relational Model}\label{sec:towards-relational-model}

% In the previous section, we gave a denotational semantics to the terms of the
% gradually-typed lambda calculus using the tools of SGDT. We defined a big-step
% semantics for closed terms of type $\nat$, as functions from $\mathbb{N} \to
% ((\mathbb{N} + 1) + 1)$.

Now that we have seen how to model the terms without regards to
graduality, we next turn to how to enhance this model to provide a
compositional semantics that additionally satisfies graduality. By a
compositional semantics, we mean that we want to provide a
compositional interpretation of \emph{type} and \emph{term precision}
such that we can extract the graduality relation from our model of
term precision.

\subsection{A first attempt: Modeling Term Precision with Posets}

In prior work, New and Licata model the term precision relation by
equipping $\omega$-CPOs with a second ordering $\ltdyn$ they call the
``error ordering'' such that a term precision relation $M \ltdyn M' :
r(A)$ is modeled by $\sem{M} \ltdyn \sem{M'}$ in the semantics. Then
the error ordering on the free error domain $\li A$ is based on the
graduality property: essentially either $\sem{M}$ errors or both terms
diverge or both terms terminate with values related by the ordering
relation on $A$.
%
The obvious first try at a guarded semantics that models graduality
would be to similarly enhance our value types and computation types to
additionally carry a poset structure. For this value types can simply
be posets. For computation types, we want our error domains to be
posets as well. To consider what additional properties we want the
error domains to have, let's consider the particular case of the free
error domain $\li A$.

The ordering on $\li A$ should model the graduality property, but how
to interpret this is not so straightforward. The graduality property
as stated mentions divergence, but we only work indirectly with
diverging computations in the form of the abstract steps/think
maps. The closest direct encoding of the graduality property is an
ordering we call the \emph{step-insensitive error ordering}, defined
by guarded recursion in
Figure~\ref{fig:step-insensitive-error-ordering}.
%
Firstly, to model graduality, $\mho$ is the least element. Next, two
returned values are related just when they are in the ordering $\leq$
on $A$.
%
Two thinking elements are related just when they are related later.
%
The interesting cases are then those where one side is thinking and
the other has completed to either an error or a value.
%
Since the graduality property is \emph{extensional}, i.e., oblivious
to the number of steps taken, it is sensible to say that if one side
has terminated and the other side is thinking, that we must require
the thinking side to eventually terminate with a related behavior,
which is the content of the final $4$ cases of the definition.

\begin{figure*}
    \begin{align*}
        &\mho \semltbad l \text{ iff } \top \\
        %
        &\eta\, x \semltbad \eta\, y \text{ if } 
            x \semlt y \\
        %		
        &\theta\, \tilde{l} \semltbad \theta\, \tilde{l'} \text{ iff } 
            \later_t (\tilde{l}_t \semltbad \tilde{l'}_t) \\
        %	
        &\theta\, \tilde{l} \semltbad \mho \text{ iff } \exists n. \theta\, \tilde{l} = \delta^n(\mho) \\
        %	
        &\theta\, \tilde{l} \semltbad \eta\, y \text{ iff } \exists n. \exists x \leq y.
            (\theta\, \tilde{l} = \delta^n(\eta\, x)) \\
        %
        &\theta\, \tilde{l} \semltbad \mho \text { iff } 
            \exists n. \theta\, \tilde{l} = \delta^n(\mho) \\
        %	
        &\eta\, x \semltbad \theta\, \tilde{l'} \text { iff }
            \exists n. \exists y \geq x. (\theta\, \tilde{l'} = \delta^n (\eta\, y))
    \end{align*}
    \caption{The step-insensitive error ordering.}
    \label{fig:step-insensitive-error-ordering}
\end{figure*}

This definition of $\semltbad$ is an entirely reasonable attempt to
model graduality: for instance it is true that if $l \semltbad l'$ for
$\li \mathbb{N}$ where $\mathbb{N}$ is the flat poset whose ordering
is equality, then the bigstep semantics we derive for $l, l'$ do in
fact satisfy the intended graduality relation.
%
However, there is a major issue with this definition: it's not a poset
at all: it is not anti-symmetric, but more importantly it is
\emph{not} transitive!
%
To see why, we observe the following undesirable property of any
relation that is transitive and ``step-insensitive'' on one side:
\begin{theorem}\label{thm:no-go}
  Let $R$ be a binary relation on the free error domain $U(\li A)$. If
  $R$ satisfies the following properties
  \begin{enumerate}
  \item Transitivity
  \item $\theta$-congruence: If $\later_t (\tilde{x}_t \binrel{R} \tilde{y}_t)$, then $\theta(\tilde{x}) \binrel{R} \theta(\tilde{y})$.
  \item Right step-insensitivity: If $x \binrel{R} y$ then $x \binrel{R} \delta y$.
  \end{enumerate}
  Then for any $l : U(\li A)$, we have $l R \Omega$. If $R$ is left
  step-insensitive instead then $\Omega \binrel{R} x$.
\end{theorem}
\begin{proof}
  By L\"ob induction, in the appendix.
\end{proof}
\begin{corollary}
  Let $R$ be a binary relation on $U(\li A)$. If $R$ satisfies
  transitivity, $\theta$-congruence and left and right
  step-insensitivity, then $R$ is the total relation: $\forall x, y. x
  \binrel{R} y$.
\end{corollary}
\begin{proof}
  $x \binrel{R} \Omega$ and $\Omega \binrel{R} y$ by the previous
  theorem. Then by transitivity $x \binrel R y$.
\end{proof}

This in turn implies $\semltbad$ is not transitive, since it is a
$\theta$-congruence and step-insensitive on both sides, but not
generally trivial: e.g., for the flat poset on $\mathbb N$, $\eta 0 \semltdbad
\eta 1$ is false.
%
This shows definitively that we cannot provide a compositional
semantic graduality proof based on types as posets if we use this
step-insensitive ordering on $\li A$.

\subsection{Why transitivity is essential}

At this point we might try to weaken from our initial attempt at
providing a poset semantics to a semantics where types are equipped
with merely a reflexive relation, and continue using the
step-insensitive ordering. However some level of transitivity is
absolutely essential to providing compositional reasoning, and is used
pervasively in the New-Licata development.
%
The reason that it is essential has to do with the details of the
graduality proof, so first we expand on how we prove graduality
compositionally. First of all, we need to extend our compositional
semantics of types and terms to also a compositional semantics of
\emph{type precision derivations} and \emph{term precision}.
%
Sticking to a poset-based semantics, if our value types are posets, we
then want our precision derivations $c : A \ltdyn A'$ to be
\emph{poset relations} on $A,A'$, that is relations between the
underlying sets of $A,A'$ that are \emph{downward-closed} in $A$ and
\emph{upper-closed} in $A'$. The obvious example of such a relation is
the poset ordering itself $\leq_A : A \rel A'$...\max{TODO: Eric pick
  up from here. Go into the squares for casts and how representability
  gives us a compositional approach to proving them}

\eric{Read this}
The lack of transitivity presents a major barrier towards giving a compositional
semantics to gradual typing: if our relations are not transitive, then we cannot
compose squares horizontally. But we have argued above that horizontal
composition is essential for modelling the axioms of gradual typing in a
syntax-independent manner.

% In particular, to model the UpL/UpR/DnL/DnR rules
% for casts in the presence of transitivity, it is sufficient to establish the
% validity of simpler rules that do not ``build in'' composition. We can then use
% transitivity to derive the original versions of the rules. On the other hand,
% without transitivity we must instead validate the cast rules in the model ``from
% scratch''. In fact, we cannot even define the semantic interpretation of type
% precision in a syntax-independent manner. The issue is that in the definition of
% relation, the requirement that it be representable now involves quantifying over
% all other relations, which is circular.

% So, although this approach would suffice for proving graduality, it lacks the
% compositionality that we seek in a reusable framework for the semantics of
% gradual typing.

%Because it is convenient to make use of transitive reasoning in proving graduality, ...
We therefore seek to formulate an ordering relation on $\li A \times \li A'$
that \emph{is} transitive. The theorem above tells us, however, that any such
relation must either not be a congruence with respect to $\theta$, or must not
be closed under delay.
\footnote{Technically, we only need to give up closure under delay on \emph{one}
side, as the above theorem requires $R$ to be closed on both sides. We could
define two separate ``step-semi-sensitive'' relations on $\li A \times \li A'$,
each closed under delays on the left and right respectively. The denotation of
term precision would then involve the conjunction of those relations. We do not
pursue this approach in this paper, because (1) it is not clear if the
step-semi-sensitive relations would be transitive, and (2) we would need to
reason about the two ordering relations separately for every pair of terms.}
%
We cannot forego the property of being a $\theta$-congruence, as without this we
would not be able to prove properties of the relation using \lob-induction,
e.g., that $\ext{f}{}$ is a \emph{monotone} function with respect to the
ordering.
%and moreover this would mean that $\theta$ would not be a monotone function.

Thus, we choose to sacrifice closure under delays. That is, we will define an
ordering relation that requires terms to be in ``lock-step''. In order for two
computations to be related in this ordering, they must have the same stepping
behavior (or the left-hand side must be an error). To deal with terms that take
differing numbers of steps, we then define a separate relation called \emph{weak
bisimilarity} that relates terms that are extensionally equal and may differ by
a finite number of delays. Then, the semantics of the error ordering for the
guarded lift monad will involve a combination of these two relations: the
``closure'' of the lock-step error ordering under weak bisimilarity on both
sides.
% Although the combined relation will not be transitive (for the same reason that
% $\semltbad$ is not transitive), this...
This decomposition has the advantage that we can recover some transitive
reasoning and push much of the reasoning about stepping to the margins of the
development.

% Doing so, we define a \emph{lock-step} error ordering, where roughly speaking,
% in order for computations to be related, they must have the same stepping
% behavior. We then formulate a separate relation, \emph{weak bisimilarity}, that
% relates computations that are extensionally equal and may differ only in their
% stepping behavior. % up to a finite number of \delta's
% Finally, the semantics of term precision will involve a combination of these two
% relations, a sort of closure of the lock-step ordering under weak bisimilarity
% on both sides. This decomposition has the advantage that we can recover some
% transitive reasoning and push the parts involving stepping to the margins of the
% development.

\subsection{Resolution: Splitting Error Ordering and Bisimilarity}\label{sec:lock-step-and-weak-bisim}

As mentioned, to recover transitive reasoning we will decompose the error
ordering on the guarded lift monad into two separate relations. The first is the
\emph{lock-step error-ordering} (also called the ``step-sensitive error
ordering''), the idea being that two computations $l$ and $l'$ are related if
they are in lock-step with regard to their intensional behavior, up to $l$
erroring. This will be the partial ordering with which we equip the guarded lift monad.
Figure \ref{fig:lock-step-error-ordering} gives the definition.

\begin{figure}
    \fbox{$l_1 \ltls l_2$}
    \begin{align*}
        &\eta\, x \ltls \eta\, y \text{ if } 
            x \mathbin{R} y \\
        %		
        &\mho \ltls l' \\
        %
        &\theta\, \tilde{l} \ltls \theta\, \tilde{l'} \text{ if } 
            \later_t (\tilde{l}_t \ltls \tilde{l'}_t)
    \end{align*}
    \caption{The lock-step error ordering.}
    \label{fig:lock-step-error-ordering}
\end{figure}

When both sides are $\eta$, then we check that the returned values are related
in $R$. The error term $\mho$ is below everything. Lastly, if both sides step
(i.e., are a $\theta$) then we compare the resulting computations one time step
later.
%
It is straightforward to prove using \lob-induction that this relation is
reflexive, transitive and anti-symmetric given that the underyling relation $R$
has those properties. The lock-step ordering is therefore the partial ordering
we will associate with $\li X$.
%
We similarly define a heterogeneous version of this ordering between the lifts
of two different types $X$ and $Y$, parameterized by a relation $R$ between $X$
and $Y$.

The second relation we define on the guarded lift monad is \emph{weak bisimilarity}.
% For a type $X$, we define a relation on $\li X$, called ``weak bisimilarity",
% written $l \bisim l'$. 
Intuitively, we say $l \bisim l'$ if they are the same ``up to delays''.
%
Weak bisimilarity on $\li X$ is parameterized by an underlying relation $R$ on $X$
and is defined by guarded fixpoint as shown in Figure \ref{fig:weak-bisimilarity}.

\begin{figure}
    \fbox{$l_1 \bisim l_2$}
    \begin{align*}
        &\mho \bisim \mho \\
      %
        &\eta\, x \bisim \eta\, y \text{ if } 
          x \mathbin{R} y \\
      %		
        &\theta\, \tilde{x} \bisim \theta\, \tilde{y} \text{ if } 
          \later_t (\tilde{x}_t \bisim \tilde{y}_t) \\
      %	
        &\theta\, \tilde{x} \bisim \mho \text{ if } 
          \theta\, \tilde{x} = \delta^n(\mho) \text { for some $n$ } \\
      %	
        &\theta\, \tilde{x} \bisim \eta\, y \text{ if }
          (\theta\, \tilde{x} = \delta^n(\eta\, x))
        \text { for some $n$ and $x : X$ such that $x \mathbin{R} y$ } \\
      %
        &\mho \bisim \theta\, \tilde{y} \text { if } 
          \theta\, \tilde{y} = \delta^n(\mho) \text { for some $n$ } \\
      %	
        &\eta\, x \bisim \theta\, \tilde{y} \text { if }
          (\theta\, \tilde{y} = \delta^n (\eta\, y))
        \text { for some $n$ and $y : X$ such that $x \mathbin{R} y$ }
      \end{align*}
    \caption{The weak bisimilarity relation.}
    \label{fig:weak-bisimilarity}
\end{figure}

When both sides are $\eta$, then we ensure that the underlying values are
bisimilar in the underlying bisimilarity relation on $A$. When one side is a
$\theta$ and the other is $\eta x$ (i.e., one side steps), we stipulate that the
$\theta$-term runs to $\eta y$ where $x$ is related to $y$. When one side is
$\theta$ and the other $\mho$, we require that the $\theta$-term runs to $\mho$.
If both sides step, then we allow one time step to pass and compare the
resulting terms. This definition captures the intuition of two terms being
related up to delays.

It can be shown (by \lob-induction) that weak bisimilarity is reflexive and
symmetric. We can also show that this relation is \emph{not} transitive. The
argument is the same as that used to show that the step-insensitive error
ordering $\semltbad$ described above is not transitive. In particular, we
observe that weak bisimilarity is a congruence with respect to $\theta$, and we
can show by $\lob$-induction that it is closed under delay on both sides. Thus,
we may appeal to the no-go theorem as before.
% We appeal to the no-go theorem given above

% This issue of non-transitivity will be resolved when we consider the relation's \emph{globalization}.

Now we can define the denotation of term precision for closed terms of type
$\nat$ as follows:
%
\[ l_1 \leq^{\bisim} l_2 := 
  \Sigma_{l_1', l_2' \in \li \mathbb{N}}. l_1 \bisim l_1' \ltls l_2' \bisim l_2. \]

Importantly, we note that as with the original ordering $\semltbad$ defined at
the beginning of this section, the relation just defined is not transitive
(again as a result of the above no-go theorem). It may therefore seem that we
have not solved the original issue we faced. In a sense, this is true. The model
of gradual typing that we end up with will not support \emph{extensional}
compositional reasoning. However, by decomposing the denotation of term
precision in the above manner, we can employ \emph{intensional} compositional
reasoning by working with the lock-step error ordering. Thus, the tentative plan
going forward will be to carry out the proofs compositionally using the
lock-step ordering. We will then apply the closure under weak bisimilarity in
the denotation of term precision given above to handle the aspects involving
stepping.

% ...to carry out the proofs using the lock-step ordering and handle the stepping
% behavior separately via weak bisimilarity. In the end we combine everything
% together using the above denotation for term precision

% Indeed, we will need to weaken our abstract notion of categorical model to no
% longer require transitivity of term precision. On the other hand, we now have
% the ability to apply transitive reasoning when working with the lock-step
% ordering. This allows us to carry out complex proofs compositionally, a key goal
% of our framework.

\subsection{Modeling Type and Term Precision}
\max{TODO: figure out where this stuff goes}
% We will model type precision $c : A \ltdyn A'$ as a relation between the sets
% $\sem{A}$ and $\sem{A'}$.

To model term precision, we begin by equipping the denotation of every type with
an ordering relation. Since term precision is reflexive and transitive, and
since we identify terms that are equi-precise, we model value types as
partially-ordered sets and values $\Gamma \vdash V : A$ as \emph{monotone}
functions. Analogously, every error domain is now equipped with a partial
ordering for which the error element is the bottom element, and the map
$\theta_B : \laterhs B \to B$ is now required to be monotone. 
%
Morphisms of error domains are morphisms of the underlying partially-ordered
sets that preserve the error element and $\theta$ map, as was the case in the
previous section.

We model a type precision relation $c : A \ltdyn A'$ as a \emph{monotone relation}, i.e., a
relation $c$ that is upward-closed under the relation on $A'$ and
downward-closed under the relation on $A$. We denote such a relation between $A$
and $A'$ by $c : A \rel A'$. The relation on the poset $A$ is denoted $r(A)$.
% This extends to products $c_1 \times c_2$ in the obvious way.
Composition of relations on predomains is the usual relational composition
(which is truncated to be propositional).

We likewise define a notion of \emph{error domain relation} between error
domains $B$ and $B'$. An error domain relation $d : B \rel B'$ is a monotone
relation on the underlying sets that preserves error and respects $\theta$. More
formally:

\begin{definition}
  Let $B$ and $B'$ be error domains. An \emph{error domain relation} between
  $B$ and $B'$ is a monotone relation $d$ on the underlying partially-ordered sets such that:
  \begin{enumerate}
     \item (Respects error): For all $y \in B'$, we have $(\mho_B, y) \in d$.
     \item (Preserves $\theta$): For all $\tilde{x}$ in $\laterhs B$ and
     $\tilde{y} \in \laterhs B'$, if $\laterhs_t ((\tilde{x}_t, \tilde{y}_t) \in
     d)$, then $(\theta_B (\tilde{x}), \theta_{B'} (\tilde{y})) \in d$. 
  \end{enumerate}
\end{definition}

Because of a technical restriction involving the interaction between
propositional truncation and the later modality, we cannot define the
composition of error domain relations as simply the composition of their
underlying predomain relations. Instead, we define composition of error domain
relations $d$ on $B_1$ and $B_2$ and $d'$ on $B_2$ and $B_3$ to be the least
relation containing $d$ and $d'$ that is downward closed under $\le_{B_1}$,
upward-closed under $\le_{B_3}$, respects error, and preserves $\theta$.
Specifically, it is defined inductively by the following rules:
%
\begin{mathpar}
    \inferrule*[right = Comp]
    {b_1 \mathbin{d} b_2 \and b_2 \mathbin{d'} b_3}
    {b_1 \mathbin{d \relcomp d'} b_3}

    \inferrule*[right = DnClosed]
    {b_1' \le_{B_1} b_1 \and b_1 \mathbin{d \relcomp d'} b_3}
    {b_1' \mathbin{d \relcomp d'} b_3}

    \inferrule*[right = UpClosed]
    {b_1 \mathbin{d \relcomp d'} b_3 \and b_3 \le_{B_3} b_3'}
    {b_1 \mathbin{d \relcomp d'} b_3'}

    \inferrule*[right = PresErr]
    { }
    {\mho_{B_1} \mathbin{d \relcomp d'} b_3}

    \inferrule*[right = PresTheta]
    {\later_t( \tilde{b_1} \mathbin{d \relcomp d'} \tilde{b_3} ) }
    {\theta_{B_1}(\tilde{b_1}) \mathbin{d \relcomp d'} \theta_{B_3}(\tilde{b_3}) }
\end{mathpar}

% That is,
% $(\mho_B, y) \in d$ for all $y : B'$ and if $\laterhs_t ((\tilde{x}_t,
% \tilde{y}_t) \in d)$, then $(\theta_B (\tilde{x}), \theta_{B'} (\tilde{y})) \in
% d$.

The functors $U$, $F$, $\times$, and $\arr$ introduced in the previous
section extend to relations. The functor $U$ simply takes the underlying
monotone relation of an error domain relation. Given a relation $c_1 : A_1 \rel
A_1'$ and $c_2 : A_2 \rel A_2'$, the relation $c_1 \times c_2 : (A_1 \times A_2)
\rel (A_1' \times A_2')$ is defined by $((x, y), (x', y')) \in (c_1 \times c_2)$
if and only if $(x, x') \in c_1$ and $(y, y') \in c_2$. Given a relation $c : A
\rel A'$ and an error domain relation $d : B \rel B'$, the relation $c \arr d :
(A \arr B) \rel (A' \arr B')$ is the error domain relation defined by $(f, g)
\in (c \arr d)$ if and only if for all $x : A$ and $x' : A'$ with $(x , x') \in
c$ we have $(f(x), g(x')) \in Ud$. The fact that this relation preserves error
and respects $\theta$ follows from the corresponding properties for $d$.
%
We we will postpone the description of the action of $F$ on relations to later
in this section.

\subsubsection{Modelling Term Precision}\label{sec:modeling-term-precision}
To model term precision, we need the notion of \emph{square}, which specifies a
relation between two monotone functions. The definition is as follows. Let $A_i,
A_o, A_i', A_o'$ be partially-ordered sets. Let $c_i : A_i \rel A_i'$ and $c_o :
A_o \rel A_o'$ be monotone relations, and let $f : A_i \to A_o$ and $g : A_i'
\to A_o'$ be monotone functions. We say that $f \ltsq{c_i}{c_o} g$ if for all $x
: A_i$ and $y : A_i'$ with $(x, y) \in c_i$, we have $(f(x), g(y)) \in c_o$. We
visualize this situation as the following square:

% https://q.uiver.app/#q=WzAsNCxbMCwwLCJBX2kiXSxbMCwxLCJBX28iXSxbMSwwLCJBX2knIl0sWzEsMSwiQV9vJyJdLFsyLDMsImciXSxbMCwxLCJmIiwyXSxbMCwyLCJjX2kiLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFsxLDMsImNfbyIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
\[\begin{tikzcd}[ampersand replacement=\&]
	{A_i} \& {A_i'} \\
	{A_o} \& {A_o'}
	\arrow["{c_i}", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow["f"', from=1-1, to=2-1]
	\arrow["g", from=1-2, to=2-2]
	\arrow["{c_o}"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
\end{tikzcd}\]

We note that for any $c$, there is a ``vertical identity'' square $\id
\ltsq{c}{c} \id$, as can be seen by unfolding the definition. Similarly, for any
montone function $f : A_i \to A_o$, there is a ``horizontal identity'' square $f
\ltsq{r(A_i)}{r(A_o)}$ by the definition of monotone function.
We likewise make the analogous definition of square for error domains.

Having defined squares, we now discuss the interpretation of the cast rules
UpL/UpR/DnL/DnR. These rules specify a relationship between the semantics of
type precision and the corresponding casts (see Section \ref{sec:GTLC}). In
particular, they specify that the relations are \emph{representable} by the
casts. The UpL and UpR rules refer to upcasts, and since upcasts are pure, these
rules apply to \emph{value} relations $c$. The DnL and DnR rules refer to
downcasts, and since downcasts are effectful, these rules apply to
\emph{computation} relations $d$.

As an example, consider the UpL rule. It states that if $c : A_1 \ltdyn A_2$ and
$c' : A_2 \ltdyn A_3$ and $x : A_1$ is related to $y : A_3$ via the composition
$c \comp c'$, then the upcast $\upc{c} x$ is related to $y$ via $c'$. This is
equivalent to the existence of the following square:

% https://q.uiver.app/#q=WzAsNSxbMCwwLCJBXzEiXSxbMSwwLCJBXzIiXSxbMiwwLCJBXzMiXSxbMCwxLCJBXzIiXSxbMiwxLCJBXzMiXSxbMyw0LCJjJyIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzAsMywiXFx1cGN7Y30iLDJdLFsyLDQsIlxcaWQiXSxbMCwxLCJjIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMSwyLCJjJyIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
\[\begin{tikzcd}[ampersand replacement=\&]
	{A_1} \& {A_2} \& {A_3} \\
	{A_2} \&\& {A_3}
	\arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow["{\upc{c}}"', from=1-1, to=2-1]
	\arrow["{c'}", "\shortmid"{marking}, no head, from=1-2, to=1-3]
	\arrow["\id", from=1-3, to=2-3]
	\arrow["{c'}"', "\shortmid"{marking}, no head, from=2-1, to=2-3]
\end{tikzcd}\]

The three squares corresponding to the other rules are analogous. Observe that
although this is really a fact about the relationship between $c$ and $\upc{c}$,
we must universally quantify over all relations $c'$. This means we cannot
require the existence of this square as part of the definition of a monotone
relation, as it is self-referential. This would seem to imply that we cannot
give a syntax-independent semantic model for gradual typing. However, the key
observation is that through compositionality, we can \emph{derive} the above
squares from simpler ones that do not involve composition of relations. For
example, below is the simpler square corresponding to UpL:

% https://q.uiver.app/#q=WzAsNCxbMCwwLCJBXzEiXSxbMCwxLCJBXzIiXSxbMSwwLCJBXzIiXSxbMSwxLCJBXzIiXSxbMCwyLCJjIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMSwzLCJyKEFfMikiLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFswLDEsIlxcdXBje2N9IiwyXSxbMiwzLCJcXGlkIl1d
\[\begin{tikzcd}[ampersand replacement=\&]
	{A_1} \& {A_2} \\
	{A_2} \& {A_2}
	\arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow["{\upc{c}}"', from=1-1, to=2-1]
	\arrow["\id", from=1-2, to=2-2]
	\arrow["{r(A_2)}"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
\end{tikzcd}\]

We can use this to derive the original UpL square by composing this square with
the identity square for $c'$:

% https://q.uiver.app/#q=WzAsNixbMCwwLCJBXzEiXSxbMSwwLCJBXzIiXSxbMiwwLCJBXzMiXSxbMCwxLCJBXzIiXSxbMiwxLCJBXzMiXSxbMSwxLCJBXzIiXSxbMCwxLCJjIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMSwyLCJjJyIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzMsNSwicihBXzIpIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbNSw0LCJjJyIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzAsMywiXFx1cGN7Y30iLDJdLFsyLDQsIlxcaWQiXSxbMSw1LCJcXGlkIiwyXV0=
\[\begin{tikzcd}[ampersand replacement=\&]
	{A_1} \& {A_2} \& {A_3} \\
	{A_2} \& {A_2} \& {A_3}
	\arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow["{\upc{c}}"', from=1-1, to=2-1]
	\arrow["{c'}", "\shortmid"{marking}, no head, from=1-2, to=1-3]
	\arrow["\id"', from=1-2, to=2-2]
	\arrow["\id", from=1-3, to=2-3]
	\arrow["{r(A_2)}"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
	\arrow["{c'}"', "\shortmid"{marking}, no head, from=2-2, to=2-3]
\end{tikzcd}\]

The existence of these simpler squares for a relation can be stated without
reference to other relations. Thus, we can require that these squares exist in
the definition of relation without running into the circularity issue we faced
with the original rules.

Finally, we use squares to model the term precision ordering $\Gamma \ltdyn
\Gamma' \vdash M \ltdyn N : A \ltdyn A'$ as an ordering relation between the
monotone functions denoted by $M$ and $N$:

% https://q.uiver.app/#q=WzAsNCxbMCwwLCJcXEdhbW1hIl0sWzAsMSwiVUYgQSJdLFsxLDAsIlxcR2FtbWEnIl0sWzEsMSwiVUZBJyJdLFsyLDMsIlxcc2Vte059Il0sWzAsMSwiXFxzZW17TX0iLDJdLFswLDIsIlxcR2FtbWFee1xcbHRkeW59IiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMSwzLCJVRmMiLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dXQ==
\[\begin{tikzcd}[ampersand replacement=\&]
	\Gamma \& {\Gamma'} \\
	{UF A} \& {UFA'}
	\arrow["{\Gamma^{\ltdyn}}", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow["{\sem{M}}"', from=1-1, to=2-1]
	\arrow["{\sem{N}}", from=1-2, to=2-2]
	\arrow["UFc"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
\end{tikzcd}\]

where $\Gamma = \{A_1, A_2, \dots\}$ and $\Gamma' = \{A_1' , A_2', \dots\}$ and for
each $j$, $c_j : A_j \ltdyn A_j'$, and the relation $\Gamma^{\ltdyn}$ is simply
the product of the $c_i$.

It remains to define $Fc$, i.e., the action of $F$ on relations.
%
% However, lifting a relation between $A$ and $A'$ to a relation between $\li A$
% and $\li A'$ proves problematic if we allow computations that take different
% numbers of steps to be related. To illustrate the issue, let us define an
% ordering $\semltbad$ between $\li A$ and $\li A'$; we call this the
% \emph{step-insensitive error ordering}. 
%
We note that the graduality property and the axioms of the inequational theory
are independent of the intensional stepping behavior of terms, so our ultimate
notion that interprets term precision will need to be oblivious to stepping as
well.
%
To that end, let us define a \emph{step-insensitive error ordering} $\semltbad$
between $\li A$ and $\li A'$; The ordering is parameterized by an ordering
relation $\le$ between $A$ and $A'$. The definition is by guarded recursion and
is shown in Figure \ref{fig:step-insensitive-error-ordering}. Recall that
$\delta : \li A \to \li A$ is defined by $\delta = \theta_A \circ \nxt$.

Two computations that immediately return $(\eta)$ are related if the returned
values are related in the underlying ordering. The computation that errors
$(\mho)$ is the least term in the ordering. If both sides step (i.e., both sides
are $\theta$), then we allow one time step to pass and compare the resulting
terms (this is where use the relation defined ``later'' and is why we employ
guarded recursion to define the relation).
%
Lastly, if one side steps and the other immediately returns a value, then in
order for these terms to be related, the side that steps must terminate with a
value in some finite number of steps $n$, and that value must be related to the
value returned by the other side. Likewise, if the LHS steps and the RHS
immediately errors, then in order to be related, the LHS must eventually
terminate with error.

\subsection{Adjusting the Cast Rules}\label{sec:adjusting-cast-rules}

We have seen how we can recover transitive reasoning by working with the
lock-step error ordering. However, it turns out that we cannot completely
isolate the reasoning about steps from the reasoning about the ordering.
% Unfortunately, by defining the partial order on the guarded lift monad to be
% the lock-step error ordering, we introduce a new issue.
In particular, the cast rules required for proving graduality no longer hold if
interpreted using the lock-step ordering. The source of the problem is that the
downcast from the dynamic type to a function takes a step, i.e., involves a
$\theta$. Consider the DnL rule shown below:
%
\begin{mathpar}
  \inferrule*{c : A \ltdyn A' \and M \ltdyn N : A'}
             {\dnc{c}{M} \ltdyn N : c}
\end{mathpar}
%
Suppose that $c$ is $\iarr \colon \dyntodyn \ltdyn\, \dyn$, and recall the
definition of the semantics of a downcast given in Section
\ref{sec:term-interpretation}. The DnL rule corresponds to the following square:

% https://q.uiver.app/#q=WzAsNCxbMCwwLCJGRCJdLFswLDEsIkZVKEQgXFxhcnIgRkQpIl0sWzEsMCwiRkQiXSxbMSwxLCJGRCJdLFswLDIsInIoRkQpIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMSwzLCJGXFxpYXJyIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMCwxLCJcXGRuY3tcXGlhcnJ9IiwyXSxbMiwzLCJcXGlkIl1d
\[\begin{tikzcd}[ampersand replacement=\&]
	FD \& FD \\
	{FU(D \arr FD)} \& FD
	\arrow["{r(FD)}", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow["{\dnc{\iarr}}"', from=1-1, to=2-1]
	\arrow["\id", from=1-2, to=2-2]
	\arrow["F\iarr"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
\end{tikzcd}\]

The downcast on the left will insert a $\theta$ in the case where the value of
type $D$ is a later-function $\tilde{f}$.
%
Thus, in order for the downcast to be related to the RHS in the lock-step error
ordering on $\li (D \to \li D)$, the RHS must be of the form $\theta(\dots)$,
and moreover, after one time step, the argument to $\theta$ on the LHS must be
related to the argument of $\theta$ on the RHS. As it stands now, this need not
be the case. Thus, we conclude that the DnL rule does not in general hold under
the lock-step error ordering. However, observe that we can remedy this
particular situation by inserting a ``delay'' on the right, i.e., wrapping the
RHS in $\delta = \theta \circ \nxt$. Notice that this delay function is weakly
bisimlar to the identity, and thus it will disappear when we apply the
bisimilarity-closure construction described above. That is, this delay makes no
difference in the extensional setting, but its presence is crucial in the
intensional setting.

% TODO Maybe show the correct form of the rule in the special case $c = inj-arr$.

\subsection{Extending the Model to Higher-Order Types}

% At this point, we could carry out the full construction of a concrete relational
% model with these additional features.

The above techniques give a semantic interpretation of term precision for closed
terms of base type. We now consider how to extend these constructions to
potentially-open terms of \emph{all} types. In particular, in order to model
higher-order data types (i.e., functions) we need to equip \emph{every} semantic
object with not only a partial ordering relation but also a ``bisimilarity''
relation that is reflexive and symmetric. We must similarly equip every object
with a structured set of delays that can be inserted and manipulated to ensure
the appropriate cast rules hold at all higher-order types. For example, the
upcast for a derivation $c_i \ra c_o$ involves the downcast corresponding to
$c_i$ and the upcast corresponding to $c_o$. We must be able to insert delays in
a functorial manner to mimic the structure of the casts. In the next section, we
will make the necessary definitions and describe the relevant constructions.

% For the sake of reusability and modularity, rather than carry out this
% construction in the concrete setting developed here, we will instead return to
% the abstract setting and adjust our definition of model to account for these
% requirements. We will break the construction into smaller steps and isolate the
% pieces that require the techniques of SGDT from those that do not. Then with
% this framework at our disposal, we will return to the construction of a concrete
% model in Section \ref{sec:concrete-model}.
% , taking as a starting point the definitions introduced in the current section.

% Thus in the next section we define revised notions of a model of
% gradual typing based on the lessons learned in the present section.