### 2024-07-12 paper-new/concrete-relational-model.tex
updates to relational model section: discuss interpretation of syntax and revise section on adequacy

\section{A Guarded Model of Graduality}\label{sec:concrete-relational-model}

Next we describe in detail our compositional model of GTLC that
satisfies graduality. To manage the complexity, we carry out the
construction in two phases, beginning with a relational model
consisting of partially-ordered sets with a notion of bisimilarity and
then adding requirements for a structured set of syntactic
perturbations on each type as well as the notion of
quasi-representability for relations that models the cast rules. We
then discuss the construction of the dynamic type and the refined
denotation of term precision that relates terms that can be
synchronized to be in lock-step. We conclude by demonstrating the
adequacy of our model with respect to the graduality property.

\subsection{Phase One: Predomains and Error Domains}

In the previous section, we introduced the notion of weak bisimilarity as a
relation on the guarded lift monad. Recall that weak bisimilarity on $\li A$ is
parameterized by a relation on $A$. To define a compositional model for all
types, we equip the posets and error domains with a reflexive, symmetric
bisimilarity relation. This leads to our first new definition:

\begin{definition}
A \textbf{predomain} $A$ consists of a set $A$ along with two relations:
\begin{itemize}
    \item A partial order $\le_A$.
    \item A reflexive, symmetric ``bisimilarity'' relation $\bisim_A$.
\end{itemize}\max{shouldn't this be $\ltdyn$?}
\end{definition}

We will generally write $A$ for both a predomain and its underlying set; if we
need to emphasize the difference we will write $|A|$ for the underlying set of
$A$.

Given a predomain $A$, we can form the predomain $\later A$. The underlying set
is $\later |A|$ and the relation is defined in the obvious way, i.e., $\tilde{x}
\le_{\later A} \tilde{x'}$ iff $\later_t(\tilde{x}_t \le_A \tilde{x'}_t)$.
Likewise for bisimilarity.
%
We also give a predomain structure to the natural numbers $\mathbb{N}$, where
both the ordering and the bisimilarity relation are equality.
%
Morphisms of predomains are functions between the underlying sets that preserve
the ordering and the bisimilarity relation. More formally:
%
\begin{definition}
Let $A$ and $A'$ be predomains. A \textbf{morphism of predomains} $f : A \to A'$
is a function between the underlying sets such that for all $x, x'$, if $x \le_A
x'$, then $f(x) \le f(x')$, and if $x \bisim_A x'$, then $f(x) \bisim_{A'}
f(x')$.
\end{definition}

We also enhance our error domains with ordering and bisimilarity
relation, and we impose a condition relating the $\theta$ morphism to
the bisimilarity relation. More concretely:
%
\begin{definition}
An \textbf{error domain} $B$ consists of a predomain $B$ along with the following data:
\begin{itemize}
    \item A distinguished ``error'' element $\mho_B \in B$ such that $\mho \le_B x$ for all $x$
    \item A morphism of predomains $\theta_B \colon \later B \to B$
    \item For all $x : B$ we have $\theta_B(\nxt\, x) \bisim_B x$
\end{itemize}\max{Don't we also need $\forall x. \mho \le x$?}
\end{definition}
%
For an error domain $B$, we define the predomain morphism $\delta_B := \theta_B
\circ \nxt$.
%
A \textbf{morphism of error domains} is simply a morphism of the underlying
predomains that preserves the error element and the $\theta$ map.

%
The definition of relations between predomains is the same as that between
posets introduced in the previous section, as the bisimilarity relations of the
predomains are not involved in the definition of monotone relation.

Next we define an analogous notion of \emph{relation} between error domains $B$
and $B'$. An error domain relation $d : B \rel B'$ is a relation on the
underlying posets that preserves error and respects $\theta$. More formally:
%
\begin{definition}
  Let $B$ and $B'$ be error domains. An \textbf{error domain relation} between
  $B$ and $B'$ is a monotone relation $d$ on the underlying partially-ordered sets such that:
  \begin{enumerate}
     \item (Respects error): For all $y \in B'$, we have $\mho_B \binrel{d} y$.\max{can we please use infix notation consistently for relations?}
     \item (Preserves $\theta$): For all $\tilde{x}$ in $\laterhs B$ and
     $\tilde{y} \in \laterhs B'$, if $\laterhs_t (\tilde{x}_t \binrel{d} \tilde{y}_t)$, then 
     $(\theta_B (\tilde{x}) \binrel{d} \theta_{B'} (\tilde{y}))$. 
  \end{enumerate}
\end{definition}
%
As with predomain relations, we can compose error domain relations. However, the
composition of the underlying predomain relations is not an error domain
relation.
% Can explain this further in the appendix
Instead, we define composition of error domain relations $d$ on $B_1$ and $B_2$
and $d'$ on $B_2$ and $B_3$ inductively to be the least relation containing $d$
and $d'$ that is downward closed under $\le_{B_1}$, upward-closed under
$\le_{B_3}$, respects error, and preserves $\theta$. See Definition
\ref{def:free-comp-ed-rel} in the appendix for the precise definition.

The definitions of squares for predomains are also unchanged from the previous
section, except that the morphisms on the left and right are now morphisms of
predomains. We likewise define a notion of \textbf{error domain square}: Let
$d_i : B_i \rel B_i'$ and $d_o : B_o \rel B_o'$ be error domain relations, and
let $\phi : B_i \multimap B_o$ and $\phi' : B_i' \multimap B_o'$ be morphisms of
error domains. We say that there is an error domain square $\phi \ltsq{d_i}{d_o}
\phi'$ exactly when there is a predomain square involving the underlying
predomain morphisms and relations.

\subsubsection{Functors}\label{sec:free-error-domain}

The functors $U$, $\li$, $\times$, and $\arr$ introduced in the previous section
act on predomains and error domains and also have actions on their morphisms,
relations, and squares. The functor $U$ simply takes the underlying predomain
relation of an error domain relation. Given a relation $c_1 : A_1 \rel A_1'$ and
$c_2 : A_2 \rel A_2'$, the relation $c_1 \times c_2 : (A_1 \times A_2) \rel
(A_1' \times A_2')$ is defined by $((x, y), (x', y')) \in (c_1 \times c_2)$ if
and only if $(x, x') \in c_1$ and $(y, y') \in c_2$. Given a relation $c : A
\rel A'$ and an error domain relation $d : B \rel B'$, the relation $c \arr d :
(A \arr B) \rel (A' \arr B')$ is the error domain relation defined by $(f, g)
\in (c \arr d)$ if and only if for all $x : A$ and $x' : A'$ with $(x , x') \in
c$ we have $(f(x), g(x')) \in Ud$. The fact that this relation respects error\max{preserves error?}
and preserves $\theta$ follows from the corresponding properties for $d$.

% Free error domain
The functor $\li$ takes a predomain $A$ to the free error domain on
$A$. We first define the predomain $U\li A$. The underlying set of
$U\li A$ is defined to be $\li |A|$ (the guarded lift monad applied to
$|A|$). The ordering relation is the lock-step error ordering
introduced in the previous section , and the bisimilarity relation is
the weak bisimilarity relation on $\li |A|$ also defined in the
previous section. With these relations, the constructors $\eta$ and
$\theta$ of the guarded lift monad are in fact morphisms of
predomains, i.e., they are monotone and preserve bisimilarity.
%
We observe that this predomain structure extends to an error domain structure by
noting that the required error element is given by the constructor $\mho$ and
the required $\theta$ map is the constructor $\theta$. Lastly, it can be shown
that the delay morphism $\delta = \theta \circ \nxt : UB \to UB$ satisfies $x \bisim
\delta\, x$ for all $x$ as is required by the definition of error domain.

The monadic bind operation $-^\dagger$ takes a predomain morphism $f : A \to UB$
to an error domain morphism $\li A \multimap B$; with this we define the action of $\li$ on
morphisms: Given $f : A \to A'$ we define $\li f = (\eta \circ f)^\dagger : \li A \to \li A'$.\max{what is the point of this sentence?}

Finally, we define the action of $\li$ on a predomain relation $c$ to be the
heterogeneous lock-step error ordering between $U\li A$ and $U\li A'$ and note
that it by definition satisfies the additional requirements of being an error
domain relation between the error domains $\li A$ and $\li A'$.

\subsection{Phase Two: Perturbations and Quasi-Representable Relations}

We now introduce the additional definitions needed to complete the construction
of our model. Recall from Section \ref{sec:lock-step-and-weak-bisim} that in order
to establish the DnL rule for the downcast of $\iarr$, we needed to adjust the
rule by inserting a ``delay'' on the right-hand side. A similar adjustment is
needed for the DnR rule for $\iarr$. Moreover, the need to insert delays impacts
the semantics of the cast rules for \emph{all} relations, because of the
functorial nature of casts. That is, the upcast at a function type $c_i \ra c_o$
involves a downcast in the domain and an upcast in the codomain. This has two
consequences: first, the squares corresponding to the rules UpL and UpR may also
require the insertion of a delay. Second, we need to be able to insert
``higher-order'' delays in a way that follows the structures of the casts.

To accomplish this, we equip every predomain $A$ with a monoid $M_A$ of
\emph{syntactic perturbations}, as well as a means of \emph{interpreting} these
syntactic perturbations as \emph{semantic} perturbations on $A$, i.e., as
endomorphisms that are bisimilar to the identity.
% Moreover, we want the
% resulting endomorphisms to be extensionally equivalent to the identity function,
% a condition we express by saying that they are weakly bisimilar to the identity
% morphism on $A$. 
More formally:
\begin{definition}[value and computation objects + morphisms]
  A \textbf{value object} consists of a predomain $A$, a monoid $M_A$ and a
  homomorphism of monoids $M_A \to \morbisimid{A} := \{ f : A \to A \mid f
  \bisim \id_A \}$.

  Likewise, a \textbf{computation object} consists of an error domain $B$, a
  monoid $M_B$ and a homomorphism of monoids $M_B \to \morbisimid{B}$.

  We define a \textbf{value morphism} to be a morphism of the underlying
  predomains, and a \textbf{computation morphism} to be a morphism of the
  underlying error domains.
\end{definition}

% every error domain $B$ with a monoid $M_B$ and interpretation homomorphism $i_B
% : M_B \to \morbisimid{B}$. This grouping of data deserves its own definition: We
% call the triple $(A, M_A, iA)$ a \textbf{value object} as it will be the
% denotation of value types in our final model. Likewise, the triple $(B, M_B,
% i_B)$ will be called a \textbf{computation object}.\max{make this a proper \textit{Definition} and define morphisms as well}

Having defined syntactic perturbations and value and computation objects, we can
now make formal the final aspect of the model: the notion of
quasi-represntability of relations.
%
\begin{definition}[quasi-left-representable relations]\label{def:quasi-left-representable}
Let $(A, M_A, i_A)$ and $(A', M_{A'}, i_{A'})$ be value objects, and let $c : A
\rel A'$ be a predomain relation. We say that $c$ is
\emph{quasi-left-representable} by a predomain morphism $e : A \to A'$ if there
are perturbations $\delle_c \in M_A$ and $\delre_c \in M_{A'}$ such that the
following two squares exist: (1) $\upl$: $e \ltsq{c}{\le_{A'}} i_{A'}(\delre_c)$, and
(2) $\upr$: $i_A(\delle_c) \ltsq{\le_{A}}{c} e$.
\end{definition}
%
Observe that this weakens the previous notion of representability, since under
that definition the perturbations were required to be the identity.
%
We make the analogous definition of quasi-left-representability for error domain
relations $d$.

Likewise, we define quasi-right-representability as follows:
%
\begin{definition}[quasi-right-represntable relations]\label{def:quasi-right-representable}
Let $(B, M_B, i_B)$ and $(B', M_{B'}, i_{B'})$ be computation objects, and let
$d : B \rel B'$ be an error domain relation. We say that $d$ is
\emph{quasi-right-representable} by an error domain morphism $p : B' \multimap B$ if
there are perturbations $\dellp_d \in M_B$ and $\delrp_d \in M_{B'}$ such that
the following two squares exist: 
(1) $\dnl$: $p \ltsq{\le_{B'}}{d} i_{B'}(\delrp_d)$ and
(2) $\dnr$: $i_B(\dellp_d) \ltsq{d}{\le_{B}} p$.
Quasi-right-representability for predomain relations is defined similarly.
\end{definition}

% \max{idea: move this after the quasi-representability part and say that this structure is what's needed to make them compose}
Given two quasi-representable relations $c$ and $c'$, it is not the case in
general that $cc'$ is quasi-representable. We need an additional condition that
specifies how relations interact with perturbations. In particular, we need to
be able to ``push'' and ``pull'' perturbations along relations $c$ and $d$.
%
The intuition for this requirement comes from the construction of the square
corresponding to the original UpL rule from the square for the simplified
version as was shown in Section \ref{sec:towards-relational-model}. Recall that
we horizontally composed the simplified UpL square for $c$ with the identity
square for $c'$ on the right. Now that the square for UpL involves a
perturbation on the right instead of the identity, this construction will not
work unless we can ``push'' the perturbation on $A_2$ to one on $A_3$. That is,
given a relation $c : A \rel A'$ and a syntactic perturbation $m_A \in M_A$, we
need to be able to turn it into a syntactic perturbation $m_{A'} \in M_{A'}$
such that the resulting semantic perturbations obtained by applying the
respective homomorphisms $i_A$ and $i_{A'}$ form a square. Dually, we will need
to ``pull'' a perturbation from right to left. This notion is made formal by the
following definition:
%
\begin{definition}
    Let $(A, M_A, i_A)$ and $(A', M_{A'}, i_{A'})$ be value objects,
    and let $c : A \rel A'$ be a predomain relation. A
    \textbf{push-pull structure} for $c$ (with respect to $M_A$ and
    $M_{A'}$) consists of monoid homomorphisms $\push : M_A \to M_A'$
    and $\pull : M_A' \to M_A$ such that for all $m_A \in M_A$ there
    is a square $i_A(m_A) \ltsq{c}{c} i_{A'}(\push\, m_A)$ and for all
    $m_{A'} \in M_{A'}$ there is a square $i_A(\pull\, m_{A'})
    \ltsq{c}{c} i_{A'}(m_{A'})$. A push-pull structure for error
    domain relations is defined similarly.
\end{definition}

We can now give the final definition of value and computation relations for our
model:
%
\begin{definition}[value relations]
A \textbf{value relation} between value objects $(A, M_A, i_A)$ and $(A',
M_{A'}, i_{A'})$ consists of a predomain relation $c : A \rel A'$ that has a
push-pull structure, is quasi-left-representable by a morphism $e_c : A \to A'$,
and is such that $\li c$ is quasi-right-representable by an error domain morphism
$p_c : \li A' \to \li A$.
\end{definition}

\begin{definition}[computation relations]
A \textbf{computation relation} between computation objects $(B, M_B, i_B)$ and
$(B', M_{B'}, i_{B'})$ consists of an error domain relation $d$ that has a
push-pull structure, is quasi-right-representable by a morphism $p_d : B' \to
B$, and is such that $Ud$ is quasi-left-representable by a morphism $e_d : UB
\to UB'$
\end{definition}

The reason for requiring $\li c$ to be quasi-right-representable in the
definition of value relations is so that we can define the action of the functor
$\li$ on relations, taking a value relation to a computation relation. Likewise,
we require $Ud$ to be quasi-left-representable in the definition of computation
relations so that we can define the action of $U$ on relations.

% Because we have augmented the definitions of object and relation for our model,
% we need to specify the actions of the functors $F$, $U$, $\times$, and $\arr$ on
% these objects and relations. 
% For instance, the functor $F$ now acts on \emph{value objects} $(A, M_A, i_A)$,
% which means we need to specify the monoid $M_{FA}$ and interpretation $i_{FA}$
% corresponding to the error domain $FA$.

\subsubsection{Constructions involving Perturbations}\label{sec:perturbation-constructions}
\max{this section needs prose}
In this section, we discuss the actions of the functors on value and computation
objects. Specifically, we define the actions of the functors on the syntactic
perturbations and their interpretation as semantic perturbations.

We first discuss the action of $\li$ on objects. Given a value object $(A, M_A,
i_A)$ we define the syntactic perturbations $M_{\li A}$ to be $\mathbb{N} \oplus
M_A$, where $\oplus$ denotes the free product of monoids (the coproduct in the
category of monoids). The intuition behind needing to include $\mathbb{N}$ comes
from the example of the downcast for $\iarr$. Recall that we adjusted the
squares corresponding to the DnL and DnR rules by adding $(\delta \circ
\eta)^\dagger$ on the side opposite the downcast. Thus, we need a syntactic
perturbation in $M_{\li A}$ that will be interpreted as $(\delta \circ
\eta)^\dagger$. We accomplish this by taking the coproduct with $\mathbb{N}$.
%
Then to interpret the perturbations as endomorphisms, by the universal property
of the coproduct of monoids it suffices to define two homomorphisms, one from
$\mathbb{N} \to \morbisimid{\li A}$ and one from $M_A \to \morbisimid{\li A}$.
The first homomorphism sends the generator $1 \in \mathbb{N}$ to the error
domain endomorpism $(\delta \circ \eta)^\dagger$, and the second sends $m_A \in
M_A$ to $\li(i_A(m_A))$ and we observe in both cases that the resulting
endomorphisms are bisimilar to the identity since the action of $\li$ on
morphisms preserves bisimilarity.

Now we discuss the action of $U$ on objects. Given a computation object $(B,
M_B, i_B)$, we define $M_{UB}$ to be $\mathbb{N} \oplus M_B$. The reason for
requiring $\mathbb{N}$ is related to the Kleisli arrow functor: to establish
quasi-representability of $U(c \arr d)$ given quasi-representability of $c$ and
$d$, we will need to turn a perturbation on $\li A$ into a perturbation on $U(A
\arr B)$. Since the perturbations for $\li A$ involve $\mathbb{N}$, so must the
perturbations for $UB$. The interpretation $i_{UB}$ of the perturbations on $UB$
works in the same manner as that of $\li A$.

The action of $\times$ on objects is as follows. Given $(A_1, M_{A_1}, i_{A_1})$
and $(A_2, M_{A_2}, i_{A_2})$, we define the monoid $M_{A_1 \times A_2} =
M_{A_1} \oplus M_{A_2}$. To construct the homomorphism $i_{A_1 \times A_2}$, we
appeal to the universal property and give homomorphisms $M_{A_1} \to
\morbisimid{A_1 \times A_2}$ and $M_{A_2} \to \morbisimid{A_1 \times A_2}$. The
former applies $i_{A_1}$ in the first component and the identity in the second
component, while the latter does the reverse with $i_{A_2}$.
% Could state this using the functorial action of \times

The action of $\arr$ on objects is as follows. Given $(A, M_A, iA)$ and $(B,
M_B, i_B)$ we define the monoid $M_{A \arr B} = M_A^{op} \oplus M_B$. Then to
construct the interpretation as endomorphisms it suffices to define a
homomorphism $M_A^{op} \to \morbisimid{A \arr B}$ and $M_B \to \morbisimid{A
\arr B}$. The former takes an element $m_A$ and a predomain morphism $f : A \to
UB$ and returns the composition $f \circ i_A(m_A)$, while the latter is defined
similarly by post-composition with $i_B$.
% Could state this using the functorial action of \arr

\subsubsection{Constructions involving Value and Computation Relations}

\eric{What, if anything, should go here? Should we discuss composition of
value/computation relations? Actions of the functors on relations?}

The fact that value and computation relations compose is proved in the appendix
(Definition \ref{def:value-computation-rel-comp}). The actions of the functors
$\li$, $U$, $\times$, and $\arr$ are also defined in the appendix (Definition
\ref{def:functorial-actions-on-relations}).

\subsection{The Dynamic Type}
We now describe the denotation of the dynamic type. Analogously to how we
defined the dynamic type in Section \ref{sec:concrete-term-model}, the predomain
representing the dynamic type is defined using a combination of least fixpoint
and guarded recursion as the solution to the equation of \emph{predomains}
%
\[ D \cong \mathbb{N}\, + (D \times D)\, + \laterhs U(D \arr \li D). \]
%
As before, we name the constructors $\inat$, $\itimes$, and $\iarr$
respectively, and the upcasts $e_{\nat}$, $e_\times$, and $e_\to$ are the same
as the constructors except for $e_{\iarr}$ which includes a $\nxt$.
%
%
% We define $e_\mathbb{N} : \mathbb{N} \to D$ to be the $\text{nat}$ constructor,
% $e_\times : D \times D \to D$ to be $\text{times}$, and $e_\to : U(D \arr F D)$
% to be the morphism $\nxt$ followed by $\text{fun}$.
%
The definition of the ordering $\le_D$ is provided in the appendix for completeness.

% given by:\max{this is kind of obvious IMO maybe it can go in the appendix}
%

%
%
We define three relations involving $D$ as follows. We define $\inat :
\mathbb{N} \rel D$ by $(n, d) \in \inat$ iff $e_\nat(n) \le_D d$. We similarly
define $\itimes : D \times D \rel D$ by $((d_1, d_2), d) \in \itimes$ iff
$e_\times(d_1, d_2) \le_D d$, and we define $\text{inj}_\to : U(D \arr \li D)
\rel D$ by $(f, d) \in \iarr$ iff $e_\to(f) \le_D d$.

Now we define the perturbations for $D$. Recall that to each predomain $A$ we
associate a monoid $M_A$ of syntactic perturbations and a homomorphism into the
monoid of endomorphisms bisimilar to the identity, and likewise for error
domains. It may seem as though we need to define the perturbations for $D$ by
guarded recursion, but in fact we define them via least-fixpoint in the category
of monoids:
%
\( M_D \cong (M_{D \times D}) \oplus M_{U(D \to \li D)}. \)
%
We now explain how to interpret these perturbations as endomorphisms.

We define $i_D : M_D \to \{ f : D \to D \mid f \bisim \id \}$ by induction as follows:
%
\begin{align*}
  i_D(m_\times) &= \lambda d. \text{case $d$ of }
    \{ \itimes (d_1, d_2) \To \itimes (i_{D \times D}(m_\times)(d_1, d_2)) \alt x \To x \} \\
  i_D(m_\to) &= \lambda d. \text{case $d$ of }
    \{ \iarr (\tilde{f}) \To \iarr (\lambda t. i_{U(D \to \li D)}(m_\to)(\tilde{f}_t)) \alt x \To x \} \\
\end{align*}
%
In the case of a perturbation on $D \times D$, we use the interpretation for the
perturbations of $D \times D$ as defined earlier in this section, which in turn
will use $i_D$ inductively. Similarly, for a perturbation on $U(D \to \li D)$ we
use the interpretation for perturbations on $U(D \to \li D)$.

% \begin{align*}
%  i_D(m_{\text{times}}, m_{\text{fun}}) &= \lambda d.\text{case $d$ of}  \\
%  &\alt \tnat(m) \mapsto \tnat(m) \\
%     &\alt \ttimes(d_1, d_2) \mapsto {\ttimes(i_{D \times D}(p_\text{times})(d_1, d_2))} \\
%     &\alt \tfun(\tilde{f}) \mapsto {\tfun(\lambda t. i_{U(D \to \li D)}(\tilde{f}_t))}
% \end{align*}

One can verify that this defines a homomorphism from $M_D \to \{ f : D \to D : f
\bisim \id \}$. Then it is straightforward to check that the three relations $\inat$, $\itimes$, and
$\iarr$ satisfy the push-pull property. As an illustrative case, we establish
the push-pull property for the relation $\iarr$. We define $\pull_{\iarr} : M_D
\to M_{U(D \arr \li D)}$ by giving a homomorphism from $M_{D \times D} \to
M_{U(D \arr \li D)}$ and from $M_{U(D \to \li D)} \to M_{U(D \arr \li D)}$. The
former is the trivial homomorphism sending everything to the identity element,
while the latter is the identity homomorphism.
%
Conversely, we define $\push_{\iarr} : M_{U(D \arr \li D)} \to M_D$ as the
inclusion into the coproduct.
%
Showing that the relevant squares exist involving push and pull is
straightforward.

It is easy to see that the relations $\inat$, $\itimes$, and $\iarr$
are quasi-left-representable, where all perturbations are taken to be the
identity elements of the respective monoids.
%
For quasi-right-representability, the only nontrivial case is $\li(\iarr)$.
Recall the defintion of the downcast for $\iarr$ given in Section
\ref{sec:term-interpretation}. We know that in order for the DnR and DnL squares
for $\li \iarr$ to exist, we must insert a delay on the side opposite the
downcast. In terms of syntactic perturbations, this means that we take $\dellp$
and $\delrp$ in the definition of quasi-right-representability to both be
$\inl\, 1$, where $\inl$ is the left injection into the coproduct of monoids. We
recall that the definition of syntactic perturbations for $\li$ involves a
coproduct with $\mathbb{N}$, and that the interpretation homomorphism maps the
generator $1$ to the endomorphism of error domains $(\delta \circ
\eta)^\dagger$.
%
With this choice for the perturbations, it is straightforward to show that the
DnL and DnR squares exist using the definition of the downcast and the
definition of the relation $\iarr$.

\subsection{Interpreting Term Precision: Extensional Squares}

As introduced in the previous section, to obtain a model of term precision that
is oblivious to the stepping behavior of terms, we weaken our notion of square
to allow for the morphisms on either side to be synchronized to be in lock-step.
More formally, let $c_i : A_i \rel A_i'$ and $c_o : A_o \rel A_o'$.
%
% We define an extensional square between morphisms $f : A_i \to A_o$ and $g :
% A_i' \to A_o'$ to consist of:
%
% \begin{itemize}
%   \item A morphism $f' : A_i \to A_o$ with $f \bisim f'$.
%   \item A morphism $g' : A_i' \to A_o'$ with $g \bisim g'$.
%   \item A square $f' \ltsq{c_i}{c_o} g'$.
% \end{itemize}
%
We say that $f \ltsqbisim{c^i}{c_o} g$ if there exist $f'$ and $g'$ such that
%
\[ f \bisim f' \ltsq{c_i}{c_o} g' \bisim g. \]
%
We call a square $f \ltsqbisim{c^i}{c_o} g$ an \emph{extensional square}, and we
will call a square $f \ltsq{c^i}{c_o} g$ an \emph{intensional square}. If the
relations $c_i$ and $c_o$ are obvious from the context, we will simply write $f
\ltbisim g$. We make the analogous construction for the computation squares.
%
We now verify the existence of the squares for the cast rules. For UpL, the
relevant extensional square is obtained as follows:
%
% https://q.uiver.app/#q=WzAsNixbMCwwLCJBXzEiXSxbMSwwLCJBXzIiXSxbMiwwLCJBXzMiXSxbMCwxLCJBXzIiXSxbMSwxLCJBXzIiXSxbMiwxLCJBXzMiXSxbMCwzLCJlX2MiLDIseyJjdXJ2ZSI6MX1dLFsxLDQsIlxcZGVscmVfYyIsMl0sWzIsNSwiXFxwdXNoX3tjJ30oXFxkZWxyZV9jKSIsMix7ImN1cnZlIjoxfV0sWzAsMSwiYyIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzEsMiwiYyciLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFszLDQsIlxcbGVfe0FfMn0iLDIseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFs0LDUsImMnIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMiw1LCJcXGlkIiwwLHsiY3VydmUiOi0xfV0sWzAsMywiZV9jIiwwLHsiY3VydmUiOi0xfV0sWzgsMTMsIlxcYmlzaW0iLDEseyJzaG9ydGVuIjp7InNvdXJjZSI6MjAsInRhcmdldCI6MjB9LCJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJub25lIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbNiwxNCwiXFxiaXNpbSIsMSx7InNob3J0ZW4iOnsic291cmNlIjoyMCwidGFyZ2V0IjoyMH0sInN0eWxlIjp7ImJvZHkiOnsibmFtZSI6Im5vbmUifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dXQ==
\[\begin{tikzcd}[ampersand replacement=\&,column sep=large]
	{A_1} \& {A_2} \& {A_3} \\
	{A_2} \& {A_2} \& {A_3}
	\arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
	\arrow[""{name=0, anchor=center, inner sep=0}, "{e_c}"', curve={height=6pt}, from=1-1, to=2-1]
	\arrow[""{name=1, anchor=center, inner sep=0}, "{e_c}", curve={height=-6pt}, from=1-1, to=2-1]
	\arrow["{c'}", "\shortmid"{marking}, no head, from=1-2, to=1-3]
	\arrow["{\delre_c}"', from=1-2, to=2-2]
	\arrow[""{name=2, anchor=center, inner sep=0}, "{\push_{c'}(\delre_c)}"', curve={height=6pt}, from=1-3, to=2-3]
	\arrow[""{name=3, anchor=center, inner sep=0}, "\id", curve={height=-6pt}, from=1-3, to=2-3]
	\arrow["{\le_{A_2}}"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
	\arrow["{c'}"', "\shortmid"{marking}, no head, from=2-2, to=2-3]
	\arrow["\bisim"{description}, draw=none, from=0, to=1]
	\arrow["\bisim"{description}, draw=none, from=2, to=3]
\end{tikzcd}\]
%
% Note that the horizontal composition of the squares is defined, as they are
% intensional squares.
We have used the push-pull structure on $c'$ to turn the perturbation on $A_2$
into a perturbation on $A_3$. Lastly we note that the bottom edge of the square
is equal to $c'$ by the fact that $c'$ is downward closed. The squares for UpR,
DnL, and DnR are obtained similarly.

% Now that we have defined an intensional model with an interpretation for the
% dynamic type, we can apply the abstract constructions introduced in Section
% \ref{sec:extensional-model-construction}. Doing so, we obtain an extensional
% model of gradual typing, where the squares are given by the ``bisimilarity
% closure'' of the intensional error ordering.

\max{Missing a subsection on how to interpret the syntax into the semantics, i.e., the actual graduality proof}
\subsection{Interpreting the Relational Axioms}

\eric{Check this}

We now turn to the interpretation of the type precision and term precision
axioms of the cast calculus (Section \ref{sec:GTLC}).

The congruence rules for term precision hold in the model, as all of
constructions are monotone and preserve weak bisimilarity.
%
The interpretation of the cast rules was discussed above.
%

It remains to verify the equivalence rules for type precision. For example,
consider the rule $(c_i \ra c_o)(c_i' \ra c_o')\equiv (c_ic_i' \ra c_oc_o')$. To
validate this rule in the model, we construct two squares $\id \ltsqbisim{(c_i
\ra c_o)(c_i' \ra c_o')}{(c_ic_i' \ra c_oc_o')} \id$ and $\id
\ltsqbisim{(c_ic_i' \ra c_oc_o')}{(c_i \ra c_o)(c_i' \ra c_o')} \id$. Since the
definition of extensional square allows us to specify intermediate morphisms
that are synchronized in the intensional ordering, it will be sufficient to
construct two intensional squares between the relations, with perturbations on
both sides. This motivates the definition of \emph{quasi-order-equivalence}.

\begin{definition}[quasi-order-equivalence]\label{def:quasi-order-equivalent}
  Let $A$ and $A'$ be value objects, and let $c, c' : A \rel A'$ be relations
  between the underlying predomains. We say that $c$ and $c'$ are
  \textbf{quasi-order-equivalent}, written $c \qordeq c'$, if there exist
  syntactic perturbations $\delta^l_1, \delta^l_2 \in M_A$ and $\delta^r_1,
  \delta^r_2 \in M_{A'}$ such that there is a square $i_A(\delta^l_1)
  \ltsq{c}{c'} i_{A'}(\delta^r_1)$ and a square $i_A(\delta^l_2) \ltsq{c'}{c}
  i_{A'}(\delta^r_2)$. Given computation objects $B$ and $B$, we make the
  analogous definition for relations $d, d' : B \rel B'$ between the underlying
  error domains.
\end{definition}

Then we have the following quasi-order-equivalences:
%
\begin{lemma}
The following hold:
\begin{itemize}
  \item $U(d \comp d') \qordeq U(d) \comp U(d')$
  \item $\li(c \comp c') \qordeq \li (c) \comp \li (c')$
  \item $(c \comp c') \to (d \comp d') \qordeq (c \to d) \comp (c' \to d')$
  \item $(c_1 \comp c_1') \times (c_2 \comp c_2') \qordeq (c_1 \times c_2) \comp (c_1'\times c_2')$
\end{itemize}
\end{lemma}
\begin{proof}
  See appendix (Lemma \ref{lem:quasi-order-equiv-functors}).
\end{proof}

\subsection{Relational Adequacy}\label{sec:adequacy}

\eric{Check this}

In this section, we prove the adequacy of the interpretation of term precision.

\begin{comment}
First we establish some notation. Fix a morphism $f : 1 \to \li \mathbb{N} \cong
\li \mathbb{N}$. We write $f \da n$ to mean that there exists $m$ such that $f =
\delta^m(\eta n)$ and $f \da \mho$ to mean that there exists $m$ such that $f =
\delta^m(\mho)$.

Recall that $\ltbisim$ denotes the relation on value morphisms defined as the
closure of the intensional error-ordering on morphisms under weak bisimilarity
on bboth sides. Specialized to the setting of morphisms from $1 \to U \li
\mathbb{N}$, we have $f \ltbisim g$ iff there exist $f'$ and $g'$ with
%
\[ f \bisim_{\li \mathbb{N}} f' \le_{\li \mathbb{N}} g' \bisim_{\li \mathbb{N}} g. \]\max{$\le$? I thought it was $\ltdyn$}
%
We first note that in this ordering, the semantics of error is not equivalent to
the semantics of the diverging term. The main result we would like to show is as
follows: If $f \ltbisim g : \li \mathbb{N}$, then:
\begin{itemize}
  \item If $f \da n$ then $g \da n$.
  \item If $g \da \mho$ then $f \da \mho$.
  \item If $g \da n$ then $f \da n$ or $f \da \mho$.
\end{itemize}
%
However, this is actually not provable! Intuitively, the issue is
that to prove this result we need to the above result is a ``global''\max{what does ``global'' mean?} result, and it is not possible to prove such results
in the guarded setting\max{isn't everything we do in a ``guarded setting''?}. In particular, if we tried to prove the above result in
the guarded setting we would run into a problem where we would have a natural
number ``stuck'' under a $\later$, with no way to get out the underlying number.
%
Thus, to prove our adequacy result, we need to leave the guarded
setting\max{``pass to leave the guarded setting''?}. We have seen in Section \ref{sec:big-step-term-semantics} that clock
quantification makes this possible.
\end{comment}

Recall the big-step term semantics $- \Downarrow -$ defined in Section
\ref{sec:big-step-term-semantics}. The result we seek to show is the following:
% If $f \bisim^{gl} f' \ltls^{gl} g' \bisim^{gl} g$, then

\begin{theorem}[adequacy]
  If $\cdot \vdash M \ltdyn N : \nat$, then:
  \begin{itemize}
    \item If $M \Downarrow n$ then $N \Downarrow n$.
    \item If $N \Downarrow \mho$ then $M \Downarrow \mho$.
    \item If $N \Downarrow n$ then $M \Downarrow n$ or $M \Downarrow \mho$.
  \end{itemize}
\end{theorem}

We have already established graduality above, which states that syntactic term
precision implies that the denoted terms are related in the semantic
interpretation of term precision.
%
% Recall as discussed in that section that the definitions of our model are all
% parameterized by a clock $k$. 
%
It therefore remains to show that the denotation of term precision in our model
$f \ltsqbisim{}{} g$, where $f, g : \li \mathbb{N}$, implies a relationship
between the partial elements $\mathbb{N} + {\mho}$ corresponding to $f$ and $g$. Recall
that to define the big-step semantics we used the equivalence of the
globalization of the free error domain and Capretta's coinductive delay monad:
$\li^{gl} \mathbb{N} \cong \delay(\mathbb{N} + {\mho})$. We now extend this
correspondence to a relationship between a global version of our denotation of
term precision and the corresponding relation on the delay monad.

% We saw in Section \ref{sec:big-step-term-semantics} that there is an isomorphism
% between the globalization of the free error domain and Capretta's coinductive
% delay monad: $\li^{gl} \mathbb{N} \cong \delay(\mathbb{N} + {\mho})$. We now
% seek to extend this to a relatinoship between a global version of our semantic
% notion of term precision and the corresponding relation for the delay monad.

To that end, we define a version of our denotation of term precision
that relates terms of type $\li^{gl} \mathbb{N}$. More concretely, consider the
predomain $\mathbb{N}$ with ordering and bisimilarity both the equality
relation. We define a global version of the lock-step error ordering and the
weak bisimilarity relation on elements of $\li^{gl} \mathbb{N}$; the former is
defined by
%
\( x \ltls^{gl} y := \forall k. x[k] \ltls y[k], \)
%
and the latter is defined by
%
\( x \bisim^{gl} y := \forall k. x[k] \bisim y[k]. \)
%

Correspondingly, we define a version of the lock-step error ordering and weak
bisimilarity relation on $\delay(\mathbb{N} + {\mho})$ 
(see Section \ref{sec:relations-on-delay-monad} in the appendix for the definitions).
% \max{do these intermediate details matter for the main body of the paper?}
%
%
By adapting Theorem 4.3 of \cite{kristensen-mogelberg-vezzosi2022} to the
setting of inductively-defined relations, we can show that both the global
lock-step error ordering and the global weak bisimilarity admit coinductive
definitions. In particular, modulo the above isomorphism between $\li^{gl} X$
and $\delay(\mathbb{N} + {\mho})$, the global version of the lock-step error
ordering is equivalent to the lock-step error ordering on $\delay(\mathbb{N} +
{\mho})$, and the global version of weak bisimilarity is equivalent to weak
bisimilarity on $\delay(\mathbb{N} + {\mho})$.
%
This implies that the global denotation of the term precision semantics for
$\li^{gl} \mathbb{N}$ agrees with the corresponding relation $\ltsqbisim{}{}^\text{Delay}$ for $\delay(\mathbb{N} + {\mho})$, i.e.,
the closure of its lock-step ordering under weak bisimilarity on both sides.
%
Then to obtain the desired result, it suffices to show being related in
$\ltsqbisim{}{}^\text{Delay}$ implies that the corresponding partial elements of
$\mathbb{N} + {\mho}$ are related in the expected manner. That is, if $d
\ltsqbisim{}{}^\text{Delay} d'$, then $p(d) \ltdyn^{Partial} p(d')$ where $p$ is
a partial function from delays into $\mathbb{N} + {\mho}$ defined by existential
quantification over the number of steps $i$ in which $d$ terminates, and the
relation $\ltdyn^\text{Partial}$ is defined such that $\mho$ is the least
element, $\bot \ltdyn^\text{Partial} \bot$, and $n \ltdyn^\text{Partial} n$.
This implication follows from the definition of $\ltsqbisim{}{}^\text{Delay}$.

% Then our result follows by proving the corresponding result for
% $\delay(\mathbb{N} + {\mho})$ which in turn follows from the definitions of the
% relations.

% \max{this doesn't actually relate to our notion of big step semantics}

% We have been writing the type as $\li X$, but it is perhaps more accurate to write it as $\li^k X$ to
% emphasize that the construction is parameterized by a clock $k$.

% Need : nat is clock irrelevant, as well as the inputs and outputs of effects
% Axioms about forcing clock
% Adapt prior argument to get that the defining of the global bisim
% and global lock-step error ordering are coinductive