### 2024-10-23 paper-new/intro.tex

\section{Introduction}

% gradual typing, graduality
\subsection{Gradual Typing and Graduality}
In programming language design, there is a tension between \emph{static} typing
and \emph{dynamic} typing disciplines. With static typing, the code is
type-checked at compile time, while in dynamic typing, the type checking is
deferred to run-time. Both approaches have benefits and excel in different
scenarios, with static typing offering compile-time assurance of a program's
type safety and type-based reasoning principles that justify program
optimizations, and dynamic typing allowing for rapid prototyping of a codebase
without committing to fixed type signatures.
%
Most languages choose between static or dynamic typing and as a result,
programmers that initially write their code in a dynamically typed language need
to rewrite some or all of their codebase in a static language if they would like
to receive the benefits of static typing once their codebase has matured.

\emph{Gradually-typed languages} \cite{siek-taha06, tobin-hochstadt06} seek to
resolve this tension by allowing for both static and dynamic typing disciplines
to be used in the same codebase. These languages support smooth interoperability
between statically-typed and dynamically-typed styles, allowing the programmer to
begin with fully dynamically-typed code and \emph{gradually} migrate portions of the
codebase to a statically typed style without needing to rewrite the project in a
completely different language.

%Gradually-typed languages should satisfy two intuitive properties.
% The following two properties have been identified as useful for gradually typed languages.

% \eric{This paragraph could be deleted}
%% In order for this to work as expected, gradually-typed languages should allow for
%% different parts of the codebase to be in different places along the spectrum from
%% dynamic to static, and allow for those different parts to interact with one another.
%% Moreover, gradually-typed languages should support the smooth migration from
%% dynamic typing to static typing, in that the programmer can initially leave off the
%% typing annotations and provide them later without altering the meaning of the program.
%% % Sound gradual typing
%% Furthermore, the parts of the program that are written in a dynamic
%% style should soundly interoperate with the parts that are written in a
%% static style.  That is, the interaction between the static and dynamic
%% components of the codebase should preserve, to the extent possible,
%% the guarantees made by the static types.  In particular, while
%% statically-typed code can error at runtime in a gradually-typed
%% language, such an error can always be traced back to a
%% dynamically-typed term that violated the typing contract imposed by
%% statically typed code. Further, static type assertions are sound in
%% the static portion, and should enable type-based reasoning and
%% optimization.

% Moreover, gradually-typed languages should allow for
% different parts of the codebase to be in different places along the spectrum from
% dynamic to static, and allow for those different parts to interact with one another.
% In a \emph{sound} gradually-typed language,
% this interaction should respect the guarantees made by the static types.

% Graduality property
One of the fundamental theorems for gradually typed languages is
\emph{graduality}, also known as the \emph{dynamic gradual guarantee},
originally defined by Siek, Vitousek, Cimini, and Boyland
\cite{siek_et_al:LIPIcs:2015:5031, new-ahmed2018}.
%
Informally, graduality says that migrating code from dynamic to
static typing should only allow for the introduction of static or
dynamic type errors, and not otherwise change the behavior of the
program.
%
This is a way to capture programmer intuition that increasing type
precision corresponds to a generalized form of runtime assertions in
that there are no observable behavioral changes up to the point of the
first dynamic type error\footnote{once a dynamic type error is raised,
in languages where the type error can be caught, program behavior may
then further diverge, but this is typically not modeled in gradual
calculi.}.
%
Fundamentally, this property comes down to the behavior of
\emph{runtime type casts}, which implement these generalized runtime
assertions.

Additionally, gradually typed languages should offer some of the
benefits of static typing. While classical type soundness, that
well-typed programs are free from runtime errors, is not compatible
with runtime type errors, we can instead focus on \emph{type-based
reasoning}. For instance, while dynamically typed $\lambda$ calculi
only satisfy $\beta$ equality for their type formers, statically typed
$\lambda$ calculi additionally satisfy type-dependent $\eta$
properties that ensure that functions are determined by their behavior
under application and that pattern matching on data types
is safe and exhaustive.

% moved the remaining paragraphs to a new section later in the intro

\subsection{Denotational Semantics in Guarded Domain Theory}

Our goal in this work is to provide an \emph{expressive},
\emph{reusable}, \emph{compositional} semantic framework for defining
such well-behaved semantics of gradually typed programs.
%
Our approach to achieving this goal is to provide a compositional
\emph{denotational semantics}, mapping types to a kind of semantic
domain, terms to functions and relations such as term precision to
proofs of semantic relations between the denoted functions.
%
Since the denotational constructions are all syntax-independent, the
constructions we provide may be reused for similar languages. Since it
is compositional, components can be mixed and matched depending on
what source language features are present.

Providing a semantics for gradual typing is inherently complicated in
that it involves: (1) recursion and recursive types through the
presence of dynamic types, (2) effects in the form of divergence and
errors (3) relational models in capturing the graduality
property. Recursion and recursive types must be handled using some
flavor of domain theory. Effects can be modeled using monads in the
style of Moggi, or adjunctions in the style of
Levy\cite{moggi,levy}. Relational properties and their verification
lead naturally to the use of reflexive graph categories or double
categories\cite{reflgraphcats,doublecats}.

The only prior denotational semantics for gradual typing was given by
New and Licata and is based on a classical Scott-style \emph{domain
theory} \cite{new-licata18}. The fundamental idea is to equip
$\omega$-CPOs with an additional ``error ordering'' $\ltdyn$ which
models the graduality ordering, and for casts to arise from
\emph{embedding-projection pairs}. Then the graduality property
follows as long as all language constructs can be interpreted using
constructions that are monotone with respect to the error ordering.
%
This framework has the benefit of being compositional, and was
expressive enough to be extended to model dependently typed gradual
typing \cite{gradualizing-cic}.
%
\max{we should make clear that we don't yet model these perverse recursive features}
However, an approach based on classical domain theory has fundamental
limitations: domain theory is incapable of modeling certain perversely
recursive features of programming languages such as dynamic type tag
generation and higher-order references, which are commonplace in
real-world gradually typed systems as well as gradual calculi.
%
Our long-term goal is to develop a semantics that will scale up to these
advanced features, and so we begin with this work by departing from
classical domain theory.

The main denotational alternative to classical domain theory that can
successfully model these advanced features is \emph{guarded} domain
theory, which we adopt in this work. While classical domain theory is
based on modeling types as ordered sets with certain joins, guarded
domain theory is based on an entirely different foundations, sometimes
(ultra)metric spaces but more commonly as ``step-indexed sets'', i.e.,
objects in the \emph{topos of trees}
\cite{birkedal-mogelberg-schwinghammer-stovring2011}.  Such an object
consists of a family $\{X_n\}_{n \in \mathbb{N}}$ of sets along with
restriction functions $r_n : X_{n+1} \to X_n$ for all $n$.  (in
category theoretic terminology, these are presheaves on the poset of
natural numbers.)  We think of a $\mathbb{N}$-indexed set as an infinite
sequence of increasingly precise approximations to the true type being
modeled.
%
Key to guarded domain theory is that there is an operator
$\triangleright$ on step-indexed sets called ``later''. In terms of
sequences of approximations, the later operator delays the
approximation by one step. Then the crucial axiom of guarded domain
theory is that any guarded domain equation $X \cong F(\triangleright
X)$ has a unique solution. This allows guarded domain theory to model
essentially \emph{any} recursive concept, with the caveat that the
recursion is \emph{guarded} by a use of the later operator.

%\max{it's confusing that this is the synthetic section but we switch back to external reasoning. Maybe move this to the previous section and revisit it here?}
Because the solutions we obtain working in guarded domain theory are constructed
as a sequence of increasingly better approximations, when we want to establish a
property that holds in the limit, we must reason about all finite
approximations. In the setting of the topos of trees, this manifests as the need
to consider \emph{global elements} of the presheaves, i.e., a family of elements
$x_i \in X_i$ compatible with the restriction maps. For example, consider the
set of programs that may take a computational step or return unit.\footnote{This example is adapted from \cite{mogelberg-paviotti2016}.} In the
topos of trees, we can model this set as the indexed family of sets $\{X_n\}_{n \in \mathbb{N}}$
where for each $n \ge 1$ we define $X_n = \{0,1,\dots,n-1, \bot\}$. Here, $i$ denotes a program that steps
$i$ times and then returns, and $\bot$ denotes a program that fails to terminate
in $n-1$ steps or fewer. For any fixed $n$, the set $X_n$ fails to distinguish a
program that takes $n$ steps and then terminates from a program that never
terminates. On the other hand, if we instead consider the denotation of a
program to be a global element of the presheaf $\{X_n\}_{n \in \mathbb{N}}$ -- a family of elements
$x_i \in X_i$ for all $i \in \mathbb{N}$ -- then the denotation of the diverging
program will be distinct from that of a program that terminates, regardless of
the number of steps it takes.

% \max{this next paragraph is a bit too mysterious}
% This has important ramifications for defining an adequate denotational semantics
% in guarded domain theory, as we seek to do in this paper: we do not want our
% denotational semantics to conflate a diverging program with a program that fails
% to terminate in sufficiently few steps. Thus, establishing adequacy of a
% semantics in guarded domain theory will require a means of globalizing,
% a point we will return to in Section \ref{sec:big-step-term-semantics}.

\subsection{Synthetic Guarded Domain Theory}

While guarded domain theory can be presented analytically using
ultrametric spaces or the topos of trees, in practice it is
considerably simpler to work \emph{synthetically} by working in a
non-standard foundational system such as \emph{guarded type
theory}. In guarded type theory later is taken as a primitive
operation on types, and we take as an axiom that guarded domain
equations have a (necessarily unique) solution. The benefit of this
synthetic approach is that when working in the non-standard
foundation, we don't need to model an object language type as a
step-indexed set, but instead simply as a set, and object-language
terms can be modeled in the Kleisli category of a simple monad defined
using guarded recursion. Not only does this make on-paper reasoning
about guarded domain theory easier, it also enables a simpler avenue
to verification in a proof assistant. Whereas formalizing analytic
guarded domain theory would require a significant theory of presheaves
and making sure that all constructions are functors on categories of
presheaves, formalizing synthetic guarded domain theory can be done by
directly adding the later modality and the guarded fixed point
property axiomatically.
%
% \max{TODO: more here, specifically this is where we should talk about adequacy I think}
%
In this paper, we work informally in a guarded type theory which is described in
more detail in Section \ref{sec:guarded-type-theory}.
%
As in the setting of analytic guarded domain theory, in the synthetic setting it
is also possible to construct global solutions. The specific approach we use in
this work is that of \emph{clocks} and \emph{clock-quantification}
\cite{atkey-mcbride2013}.

% \max{shouldn't we say we use clock quantification?}
% In the setting of synthetic guarded domain theory, there is an analogue of
% constructing global solutions. To construct a global solution we must use an
% additional construct, e.g., the $\square$ modality
% \cite{10.1007/978-3-319-08918-8_8}, whose semantics in the topos of trees model
% is to compute the global elements of a presheaf. A related approach involves
% objects known as \emph{clocks}, whereby \emph{clock-quantification}
% \cite{atkey-mcbride2013} provides a means to obtain a global solution.

\subsection{Adapting the New-Licata Model to the Guarded Setting}

Since guarded domain theory only provides solutions to guarded domain equations,
there is no systematic way to convert a classical domain-theoretic semantics to
a guarded one.  Classical domain theory has limitations in what it can model,
but it provides \emph{exact} solutions to domain equations when it applies. When
adapting the New-Licata approach to guarded domain theory, the presence of later
in the semantics makes it \emph{intensional}: unfolding the dynamic type
requires an observable computational step. For example, a function that pattern
matches on an element of the dynamic type and then returns it unchanged is
\emph{not} equal to the identity function, because it ``costs'' a step to
perform the pattern match.

% - The essential role of transitivity
% - We can't have transitivity + extensional reasoning
% - Solution: split the ordering into two, and introduce perturbations
%   to recover some amount of transitive reasoning

The intensional nature of guarded domain theory leads to the main departure of
our semantics from the New-Licata approach. New and Licata observed that
in order to even \emph{formulate} the model for gradual typing in a
syntax-independent manner, some access to transitive reasoning is essential.
%
However, in the setting of guarded domain theory, it is not possible to carry
out reasoning that is simultaneously transitive and \emph{extensional}, where
the latter means being insensitive to the number of steps the elements take. We
make this statement precise in a no-go theorem (Theorem \ref{thm:no-go}). This fundamental
limitation means that the natural adaptation of the New-Licata semantics to the
guarded setting, where types are interpreted as sets with a
\emph{step-insensitive} error ordering, cannot work, because that ordering is
necessarily \emph{not} transitive.

To recover some amount of compositional reasoning, we model the graduality
ordering using a combination of two relations: a \emph{lock-step error ordering}
in which elements must take the same number of steps, and a \emph{weak
bisimilarity relation} that relates elements differing only in the number of
computational steps that they take. The lock-step ordering \emph{is} transitive,
but it is not sufficient to model the New-Licata proof. The problem is that in
the guarded setting, the denotation of a downcast from the dynamic type involves
a computational step, and by the functorial nature of casts this implies that
the denotations of arbitrary casts may involve computational steps. As a result,
the embedding-projection characterization of casts in the New-Licata model does
not hold in the guarded setting, at least not up to equality.
%
However, while the casts may involve delays, they are otherwise extensionally
equivalent to embedding-projection pairs. Thus, one might try to use weak
bisimilarity to account for the delays, viewing casts as embedding-projection
pairs ``up to weak bisimilarity''. Unfortunately, this does not work: since weak
bisimilarity is \emph{not} transitive (which follows also from our no-go
result), we are back to the original issue of lack of transitive reasoning.

% However, since the casts involve delays but are otherwise extensionally
% equivalent to embedding-projection pairs, one might try to use weak bisimilarity
% to account for the delays. That is, one could try characterizing the casts as
% embedding-projection pairs ``up to weak bisimilarity''. Unfortunately, this does
% not work: since weak bisimilarity is \emph{not} transitive (which follows also
% from our no-go result), we are back to the original issue of lack of
% transitive reasoning.

It would seem that at this point, we are stuck: we need compositional reasoning,
but we also need to be able to account for the steps taken by casts. And now we
come to the key technical contribution of this work. The crucial observation is
that, although the denotations of casts involve delays, the specific ways in
which those delays arise follow a specific, type-directed pattern. Hence, we can
track the delay behavior in a type-directed manner, equipping each type with a
structured set of what we call \emph{syntactic perturbations}. These can then be
used to explicitly \emph{synchronize} elements to ensure that they are in lock-step.
%
In the end, we have come to a resolution of the issue: the lock-step error
ordering where we employ syntactic perturbations to synchronize elements is
still transitive, while simultaneously being able to relate terms that involve
an explicit, known pattern of computational steps.
%
We discuss these issues in more detail in Section
\ref{sec:towards-relational-model} and give a detailed description of the
resulting model in Section \ref{sec:concrete-relational-model}.

% Perturbations allow us to recover enough extensional reasoning to model the
% graduality property compositionally.

% weak bisimilarity: elements differing only in their stepping behavior are related.

% while the New-Licata semantics
% equips $\omega$-CPOs with an error ordering, we work with types equipped with
% not just an error ordering but also a \emph{bisimilarity} relation that reflects
% when elements differ only in the number of abstract computational steps that
% they take. The reason for requiring these two separate notions is that in the
% setting of guarded domain theory, it is not possible to carry out transitive
% reasoning while being oblivious to the number of steps the elements take. We
% make this statement precise in \ref{sec:towards-relational-model}. To overcome
% this limitation, we develop further changes to the core concepts of our
% semantics as discussed in that section.

% Because bisimilarity, unlike equality, is not transitive in the guarded
% setting, this necessitates further changes to the core
% concepts of our semantics. These challenges are discussed in Section \ref{sec:towards-relational-model}.

\subsection{The Graduality Proof}

Proving graduality by constructing a denotational model in guarded type theory
consists of two main steps. First, there is the construction of the model
itself. This gives us an interpretation of the terms as well as the syntactic
axioms that model the graduality property. However, because this model is
defined in the guarded domain theory, while graduality is a statement about a
set-theoretic semantics, it is not a priori clear that the guarded model
validates the graduality property.
%
What we need is to show that the guarded model induces a sensible set-theoretic
semantics in ordinary mathematics. We cannot hope to derive such a semantics for
\emph{all} types (e.g., the dynamic type), but for the subset of closed terms of
base type, we should be able to extract a well-behaved semantics for which the
graduality property holds.

% The next step is to ensure that the semantics actually validates the graduality
% property. That is, we can extract from this guarded model a well-behaved
% semantics in ordinary mathematics For the terms, this means that closed terms of
% base type should evaluate to themselves.

% Defining a model of gradual typing in guarded type theory is not enough to prove
% graduality. We must then extract a well-behaved ``big-step'' semantics for terms
% of base type and show that the denotation of term precision has the expected
% behavior. More specifically, we want a semantics in which the denotation of
% error is different from the denotation of a value, which is in turn different
% from the denotation of the diverging computation. Recall that by working in
% guarded domain theory, our definitions are constructed as a sequence of
% approximations. What we actually want is a statement that holds ``in the
% limit''.

More concretely, consider a gradually typed language whose only effects are
gradual type errors and divergence. If we fix a result type of natural numbers,
a whole program semantics is a partial function from closed programs to either
natural numbers or errors:
%
\[ -\Downarrow : \{M \,|\, \cdot \vdash M : \nat \} \rightharpoonup \mathbb{N}
\cup \{\mho\} \] 
%
where $\mho$ is notation for a runtime type error. We write $M \Downarrow n$ and
$M\Downarrow \mho$ to mean this semantics is defined as a number or error, and
$M\Uparrow$ to mean the semantics is undefined, representing divergence.
%
A well-behaved semantics should then satisfy several properties. First, it
should be \emph{adequate}: natural number constants should evaluate to
themselves: $n \Downarrow n$. Second, it should validate type based reasoning.
To formalize type based reasoning, we give a typed equational theory for terms
of the language $M \equiv N$ for when two terms should be considered equivalent.
Then we want to verify that the big step semantics respects this equational
theory: if closed programs $M \equiv N$ are equivalent in the equational theory
then they have the same semantics, $M \Downarrow n \iff N \Downarrow n$ and
$M\Uparrow \iff N \Uparrow$ and $M \Downarrow \mho \iff N \Downarrow \mho$.

Lastly, the semantics should be adequate for the graduality property. Graduality
is typically axiomatized by giving an \emph{inequational} theory called term
precision, where $M \ltdyn N$ roughly means that $M$ and $N$ have the same type
erasure and $M$ has at each point in the program a more precise/static type than
$N$. Then adequacy says that if $M$ and $N$ are whole programs returning type
$\nat$ and $M \ltdyn N$, then the denotations of $M$ and $N$ as big-step terms
should be related in the expected way: either $M$ errors, or $M$ and $N$ have
the same extensional behavior. That is, either $M\Downarrow \mho$ or $M
\Downarrow n $ and $N \Downarrow n$ or $M \Uparrow $ and $N
\Uparrow$\footnote{we use a slightly more complex definition of this relation in
our technical development below that is classically equivalent but
constructively weaker}.

To define the big-step semantics, we apply clock-quantification to the guarded
model to define a global semantics for closed terms of base type where a term
$M$ is mapped to an element of the coinductive delay monad of Capretta
\cite{lmcs:2265}. We compose this map with a partial function on elements of the
delay monad that returns the result with which given element terminates, or is
undefined if it does not terminate. This gives us the desired partial function
from closed programs to results. We describe this process in more detail in Section
\ref{sec:big-step-term-semantics}.
%
We then make the analogous construction for the relation that denotes the term
precision ordering, showing that a term precision ordering $M \ltdyn N$ implies
the corresponding ordering on the partial functions denoted by $M$ and $N$. This
construction uses as an intermediate step an analogue of the graduality relation
for the coinductive delay monad. The details of the proof are given in Section
\ref{sec:adequacy}.

% For example, when establishing the adequacy of a semantics defined using
% guarded type theory, we must reason about all finite approximations.

% In this paper, we develop an adequate denotational semantics that satisfies
% graduality and soundness of the equational theory of cast calculi using
% synthetic guarded domain theory.

\subsection{Contributions and Outline}

The main contribution of this work is a compositional denotational semantics in
guarded type theory for a simple gradually typed language that validates
$\beta\eta$ equality and satisfies a graduality theorem.
%
% In particular, the notion of syntactic perturbations stands out as a key
% technical contribution and is the most significant and novel feature of our
% model that allowed us to successfully adapt the denotational approach of New and
% Licata to the guarded setting.
% In particular, the notion of syntactic perturbations stands out as a key feature
% of our model, and is our most significant technical contribution allowing us to
% successfully adapt the denotational approach of New and Licata to the guarded
% setting.
The goal we have kept in mind while designing our model is to handle
gradually-typed languages with advanced features that classical domain theory in
incapable of modeling. We do not address these features in this work but leave
them instead to immediate future work, as handling even the simplest
gradually-typed language already poses interesting challenges.
%
The notion of syntactic perturbation stands out as a key aspect of our model
and our most significant technical contribution, allowing us to successfully
adapt the denotational approach of New and Licata to the guarded setting.
%
Most of our work has further been verified in Guarded Cubical Agda
\cite{veltri-vezzosi2020}, demonstrating that the semantics is readily
mechanizable. We provide an overview of the mechanization effort, and what
remains to be formalized, in Section \ref{sec:mechanization}.
%

% Syntactic perturbations 
%
% The notion of \emph{syntactic perturbations} allow us to encode the steps
% taken by a term in a form that we can manipulate. This allows us to recover
% enough extensional reasoning to model the graduality property compositionally.

% Syntactic perturbations give us a type-directed way of imposing delays on terms.

\begin{comment}
\begin{enumerate}
\item First, we give a simple concrete term semantics where we show
  how to model the dynamic type as a solution to a guarded domain equation.
\item Next, we identify where prior work on classical domain theoretic
  semantics of gradual typing breaks down when using guarded semantics
  of recursive types.
\item We develop a key new concept of \emph{syntactic perturbations},
  which allow us to recover enough extensional reasoning to model the
  graduality property compositionally.
\item We combine this insight together with an abstract categorical
  model of gradual typing using reflexive graph categories and
  call-by-push-value to give a compositional construction of our
  denotational model.
\item We prove that the resulting denotational model provides a
  well-behaved semantics as defined above by proving \emph{adequacy},
  respect for an equational theory and the graduality property.
\end{enumerate}
\end{comment}

The paper is laid out as follows:
% \max{I think there's too much detail in this outline, especially parts 4 and 5. Can we instead expand those into the Intro?}
\begin{enumerate}
\item In Section \ref{sec:GTLC} we fix our input language, a fairly
  typical gradually typed cast calculus.
\item In Section \ref{sec:concrete-term-model} we develop a
  denotational semantics in synthetic guarded domain theory for the
  \emph{terms} of the gradual lambda calculus.  The model is adeqauate
  and validates the equational theory, but it does not satisfy
  graduality. We use this to introduce some of our main technical
  tools: modeling recursive types in guarded type theory and modeling
  effects using call-by-push-value.
\item In Section \ref{sec:towards-relational-model} we show where the New-Licata
  classical domain theoretic approach fails to adapt cleanly to the guarded
  setting and explore the difficulties of proving graduality in an intensional
  model. We prove the no-go theorem about extensional, transitive relations,
  introduce the lock-step error ordering and weak bisimilarity relation, and
  motivate the need for perturbations.
\item In Section \ref{sec:concrete-relational-model} we describe the
  construction of the model in detail, and discuss the
  translation of the syntax and axioms of the gradually typed cast calculus into the model.
  Lastly, we prove that the model is adequate for the graduality property.
  % The resulting model validates the axioms for type and term precision specified
  % in Section \ref{sec:GTLC}. The final step is to prove that the model is
  % \emph{adequate} for the graduality property: a closed term precision $M \ltdyn
  % N : \nat$ has the expected semantics, i.e., that $M$ errors or $M$ and $N$
  % have the same extensional behavior. We do so by extending the globalization
  % techniques used in defining the big-step term model to account for the
  % interpretation of term precision.
\item In Section \ref{sec:discussion} we discuss prior work on proving
  graduality, the partial mechanization of our results in Agda, and
  future directions for denotational semantics of gradual typing.
\end{enumerate}