### 2024-11-21 paper-new/paper.tex
add POPL-specific metadata

\documentclass[acmsmall]{acmart}
\let\Bbbk\relax

\usepackage{quiver}
\usepackage{mathpartir}
% \usepackage{tikz-cd}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{comment}
\usepackage{array}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newenvironment{DIFnomarkup}{}{}
\pagestyle{empty}
\settopmatter{printfolios=false}

%% CCS Concepts
\begin{CCSXML}
  <ccs2012>
     <concept>
         <concept_id>10003752.10010124.10010131.10010133</concept_id>
         <concept_desc>Theory of computation~Denotational semantics</concept_desc>
         <concept_significance>500</concept_significance>
         </concept>
     <concept>
         <concept_id>10011007.10011006.10011039.10011311</concept_id>
         <concept_desc>Software and its engineering~Semantics</concept_desc>
         <concept_significance>300</concept_significance>
         </concept>
     <concept>
         <concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
         <concept_desc>Software and its engineering~Functional languages</concept_desc>
         <concept_significance>300</concept_significance>
         </concept>
   </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Denotational semantics}
\ccsdesc[300]{Software and its engineering~Semantics}
\ccsdesc[300]{Software and its engineering~Functional languages}

%% Additional Key Words and Phrases
\keywords{Gradual Typing, Denotational Semantics, Synthetic Guarded Domain Theory, Guarded Type Theory}

%%% The following is specific to POPL '25 and the paper
%%% 'Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory'
%%% by Eric Giovannini, Tingting Ding, and Max S. New.
%%%
\setcopyright{rightsretained}
\acmDOI{10.1145/3704863}
\acmYear{2025}
\acmJournal{PACMPL}
\acmVolume{9}
\acmNumber{POPL}
\acmArticle{27}
\acmMonth{1}
\received{2024-07-11}
\received[accepted]{2024-11-07}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.

% \setcopyright{acmcopyright}
% \copyrightyear{2018}
% \acmYear{2018}
% \acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
% \acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
%   Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
% \acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%   June 03--05, 2018, Woodstock, NY}
% \acmPrice{15.00}
% \acmISBN{978-1-4503-XXXX-X/18/06}

\input{defs}

\begin{document}

\title{Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory}
\author{Eric Giovannini}
\affiliation{
  \department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{ericgio@umich.edu}
\orcid{0009-0003-6871-1714}

\author{Tingting Ding}
\affiliation{
  \department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{tingtind@umich.edu}
\orcid{0009-0000-5676-1886}

\author{Max S. New}
\affiliation{
  \department{Electrical Engineering and Computer Science}
  \institution{University of Michigan}
  \country{USA}
}
\email{maxsnew@umich.edu}
\orcid{0000-0001-8141-195X}

\begin{abstract}
  Gradually typed programming languages, which allow for soundly
  mixing static and dynamically typed programming styles, present a
  strong challenge for metatheorists. Even the simplest sound
  gradually typed languages feature at least recursion and errors,
  with realistic languages featuring furthermore runtime allocation of
  memory locations and dynamic type tags. Further, the desired
  metatheoretic properties of gradually typed languages have become
  increasingly sophisticated: validity of type-based equational
  reasoning as well as the relational property known as
  graduality. Many recent works have tackled verifying these
  properties, but the resulting mathematical developments are highly
  repetitive and tedious, with few reusable theorems persisting across
  different developments.

  In this work, we present a new denotational semantics for gradual
  typing developed using guarded domain theory. Guarded domain theory
  combines the generality of step-indexed logical relations for
  modeling advanced programming features with the modularity and
  reusability of denotational semantics. We demonstrate the
  feasibility of this approach with a model of a simple gradually
  typed lambda calculus and prove the validity of beta-eta equality
  and the graduality theorem for the denotational model. This model
  should provide the basis for a reusable mathematical theory of
  gradually typed program semantics. Finally, we have mechanized most
  of the core theorems of our development in Guarded Cubical Agda, a
  recent extension of Agda with support for the guarded recursive
  constructions we use.
\end{abstract}

\maketitle

% Outline

% 1. Intro: What do we want out of gradually typed languages and why
%    is it hard to prove? Explanation: Gradual Typing inherently
%    involves recursive types, multiple effects, relational properties.
%    We argue that the increasing complexity of the metatheory of
%    gradual typing makes it a good candidate for

% 2. Extensional Dream Semantics: double categorical

% 3. The Problem with Step-indexing

% 4. A Compromise

% 5. Formalization in Guarded Cubical Agda

\newif\ifdraft
\drafttrue
\renewcommand{\max}[1]{\ifdraft{\color{blue}[{\bf Max}: #1]}\fi}
\newcommand{\eric}[1]{\ifdraft{\color{orange}[{\bf Eric}: #1]}\fi}
\newcommand{\tingting}[1]{\ifdraft{\color{red}[{\bf Tingting}: #1]}\fi}

\input{intro}

%\input{technical-background}

\input{syntax}

% \input{extensional-model}

\input{concrete-term-model}

\input{towards-relational-model}

% \input{semantics}

%% \input{ordering}

%% \input{graduality}

% \input{categorical-model}

%\input{denotational-model}
%\input{terms}
\input{concrete-relational-model}

\input{discussion}

\bibliographystyle{ACM-Reference-Format}
\begin{DIFnomarkup}
\bibliography{references}
\end{DIFnomarkup}

\newpage
\appendix
\input{appendix}

\end{document}